### 2015-10-29 figures/Figure2-flowDiagram.tex
Adding non-large files.

% Flowchart
% Author: Qrrbrbirlbel
\documentclass[tikz,border=10pt]{standalone}
%%%<
\usepackage{verbatim}
%%%>
\begin{comment}
:Title: Flowchart
:Tags: Diagrams;Flowcharts;Node positioning;Nodes and shapes
:Author: Qrrbrbirlbel
:Slug: flowchart

This example flowchart uses the positioning-plus library and the
node-families library by the same author. The fit library is implicitly loaded
by positioning-plus, the backgrounds library is used to draw stuff behind
other stuff, the calc library for some coordinate calculations and the
shapes.geometric library for the ellipse shape.

This example was written by Qrrbrbirlbel answering a question on TeX.SE.
\end{comment}
\usetikzlibrary{shapes.geometric,backgrounds,
  positioning-plus,node-families,calc}
\tikzset{
  basic box/.style = {
    shape = rectangle,
    align = center,
    draw  = #1,
    fill  = #1!25,
    rounded corners},
  header node/.style = {
    Minimum Width = header nodes,
    font          = \strut\Large\ttfamily,
    text depth    = +0pt,
    fill          = white,
    draw},
  header/.style = {%
    inner ysep = +1.5em,
    append after command = {
      \pgfextra{\let\TikZlastnode\tikzlastnode}
      node [header node] (header-\TikZlastnode) at (\TikZlastnode.north) {#1}
      node [span = (\TikZlastnode)(header-\TikZlastnode)]
        at (fit bounding box) (h-\TikZlastnode) {}
    }
  },
  hv/.style = {to path = {-|(\tikztotarget)\tikztonodes}},
  vh/.style = {to path = {|-(\tikztotarget)\tikztonodes}},
  fat blue line/.style = {ultra thick, blue}
}
\begin{document}
\begin{tikzpicture}[node distance = 1.2cm, thick, nodes = {align = center},
    >=latex]
  \node[Minimum Width = loop, shape = ellipse, fill = red] (imp-sol)
    {ellipsoid box};
  \node[Minimum Width = loop, fill = yellow, below = of imp-sol] (rec-box)
    {rectangular box, and very wiiiiiiiiiiiiiiide\\2nd line};
  \node[shift = (left:.5*x_node_dist)] at
    ($(imp-sol.west|-imp-sol.south)!.5!(rec-box.north west)$) (for-1)
    {formula 1};
  \node[shift = (right:.5*x_node_dist)] at
    ($(imp-sol.east|-imp-sol.south)!.5!(rec-box.north east)$) (for-2)
    {formula 2};
  \begin{scope}[on background layer]
    \node[fit = (for-1)(for-2)(imp-sol)(rec-box), basic box = blue,
      header = DMFT loop] (dmft-l) {};
  \end{scope}
  \path[very thick, blue, hv] (rec-box) edge[->] (for-1) edge[<-] (for-2)
                              (imp-sol) edge[->] (for-2) edge[<-] (for-1);

  \node[east above = of dmft-l, basic box = green, header = DMFT prelude]
    (dmft-p) {Math and text math and text math and text\\
              math and text math and text math and text};
  \node[north left = of dmft-l, basic box = green, header = $\rho$ update,
     shift = (down:y_node_dist)] (rho)
    {Much more text much more text\\much more text much more text};
  \node[basic box = blue, header = DFT part, anchor = north] at
    (dmft-p.north-|rho) (dft) {So much text so much text so much text\\
    I think I need \texttt{tikz-lipsum}\\or something like that.};
  \node[basic box = green, anchor = north] at
    ($(dft.north east)!.5!(dmft-p.north west)$) (upd) {update\\$math$};
  \path[fat blue line, <-, dashed, vh] (rho) edge
    ({$(rho.south)!.5!(dmft-l.south)$}-|dmft-l.south west);
  \path[fat blue line, ->]
    ({$(upd.south)!.5!(dmft-p.south)$}-|dmft-p.south west)
    coordinate (@) edge[<-, solid] coordinate[pos=.15] (@s)
    coordinate[pos=.9] (@e) (@-|dft.east)
    {[every edge/.append style=dashed, vh] (@s) edge[<-] (upd) (@e) edge (upd)}
    (h-rho) edge[dashed] (dft)
    ($(dmft-p.south)!.5!(dmft-p.south east)$)
    coordinate (@) edge (@|-dmft-l.north);
\end{tikzpicture}
\end{document}

### 2015-10-29 bin/pmArticleScore.pl

@@ -1,0 +1,444 @@
+#!/usr/bin/perl 
+
+#pmArticleScore.pl: 
+#      input: 1. list of "training" article from pubmed searches -- these include titles and abstracts
+#      input: 2. list of "background" articles from pubmed searches -- these include titles and abstracts (OPTIONAL: plain, further background text -- possibly non-scientific)
+#      input: 3. list of "candidate" articles from pubmed searches -- these may contain results similar to the "training" articles
+
+#      output: a ranked list of candidate articles, based upon how well the terms in candidate articles match the training articles
+
+#      Basic approach: 
+#                      TRAINING
+#                      1. build a dictionary for each training article (hash?)
+#                      2. exclude high frequency words (possibly use the further background/non-scientific text for this e.g. Alice in Wonderland)
+#                      3. score each remaining word based upon how frequently it is used in the training data relative to the background set 
+#                         (e.g. log2( [f(t.word)/t.size + d]/[f(b.word)/b.size + d] ) ...  )
+#                      
+#                      SCORING
+#                      1. build a dictionary for each candidate article (hash?)
+#                      2. exclude high frequency words
+#                      3. score each article using sum( c.word, log( [f(t.word)/t.size + d]/[f(b.word)/b.size + d] ) )
+
+#PUBMED SEARCHES:
+#                      TRAINING
+#                      less Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv | perl -lane '@F=split(/\t/); print "$F[0]\[uid\] OR " if $F[0]>0' | tr -d "\n"
[…]

### 2015-10-29 bin/prettyPlot.R

@@ -1,0 +1,147 @@
+#R CMD BATCH prettyPlot.R
+library(MASS)  # in case it is not already loaded 
+library(RColorBrewer)
+library(fields)
+k <- 11
+my.cols <- rev(brewer.pal(k, "RdYlBu"))
+d<-read.table("meanRankSpeedData", header=T)
+reg1  <- lm(d$speedRank~d$accuracyRank)
+reg1c <- coefficients(reg1)
+
+dr<-read.table("rawRankSpeedData", header=T)
+
+
+fudgeit <- function(){
+  xm <- get('xm', envir = parent.frame(1))
+  ym <- get('ym', envir = parent.frame(1))
+  z  <- get('dens', envir = parent.frame(1))
+  colramp <- get('colramp', parent.frame(1))
+  my.cols <- rev(brewer.pal(k, "RdYlBu"))
+
+  image.plot(xm,ym,z, col = my.cols, legend.only = T, add =F)
+}
+
+
+
+plotMe <- function() {
+       
+op<-par(mfrow=c(1,2),cex=1.1)
+
+
+op<-par(cex=1.1)
+plot(c(0,1),c(1,0), xlim=c(1,0),ylim=c(1,0),xlab="normalised accuracy rank", ylab="normalised speed rank",type="l",lty=2,lwd=3)
+
+#DIAGONALS
+#lines(c(0,1),c(0,1),lty=2)
+#lines(c(0,1),c(1,0),lty=2, lwd=3)
+
+#BOX
+lines(c(0,0),c(0,1),lty=1,lwd=2)#left
+lines(c(0,1),c(1,1),lty=1,lwd=2)#top
+lines(c(1,1),c(1,0),lty=1,lwd=2)#right
+lines(c(1,0),c(0,0),lty=1,lwd=2)#bottom
+
+for(id in unique(dr$testId)){       
+       regr1  <- lm(dr$normSpeedRank[ dr$testId == id]~dr$normAccuracyRank[ dr$testId == id])
+       regr1c <- coefficients(regr1)
+       if( regr1c[1]+regr1c[2] > 0 ){
+	   lines(c(0,1),regr1c[1]+regr1c[2]*c(0,1), lwd=5)
+       }
+       else{#y negative when x = 0
+            #y = mx + c = 0
+	    #  =>     x = -c/m
+	   lines(c(0, -1*regr1c[1]/regr1c[2] ),  c(regr1c[1], 0), lwd=5)	
+	   cat("Y<0!: ",id, "\n")
+       }
+}
+
[…]

### 2015-10-29 bin/tsv2data.pl

@@ -1,0 +1,119 @@
+#!/usr/bin/perl 
+
+use warnings;
+use strict;
+
+open(IN, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv"); 
+open(UT0, "> rawRankSpeedData");
+print UT0 "testId\tnormAccuracyRank\tnormSpeedRank\tmethod\n";
+
+#echo -ne "accuracyRank\tspeedRank\tnumMethods\n" > data && cut -f 7,8,9 Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv | grep -v N | tr -d "=" | perl -lane 'if(/^acc|^N/ or $F[0] !~ /\d+/ or $F[1] !~ /\d+/){next}elsif(defined($F[2])){$max=$F[2]} printf "%0.2f\t%0.2f\t$max\n", ($F[0]-1)/($max-1), ($F[1]-1)/($max-1); ' >> data
+    #  1	pubmedID
+    #  2	Title
+    #  3	accuracySource
+    #  4	accuracyMetric
+    #  5	speedSource
+    #  6	Method
+    #  7	accuracyRank
+    #  8	speedRank
+    #  9	numMethods
+    # 10	Data set (if applicable)
+    # 11	Bias
+
+
+#cat Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv | cut -f 1 | sort -d | perl -lane 'print "$F[0] => 1, "' | tr -d "\n"
+
[…]

### 2015-11-04 figures/Figure2-flowDiagram.tex
Added --ignore option. Iterated results.

@@
-% Flowchart
-% Author: Qrrbrbirlbel
-\documentclass[tikz,border=10pt]{standalone}
+\documentclass{article}
+
+\usepackage[latin1]{inputenc}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+
 %%%<
 \usepackage{verbatim}
+\usepackage[active,tightpage]{preview}
+\PreviewEnvironment{tikzpicture}
+\setlength\PreviewBorder{5pt}%
 %%%>
+
 \begin{comment}
-:Title: Flowchart
-:Tags: Diagrams;Flowcharts;Node positioning;Nodes and shapes
-:Author: Qrrbrbirlbel
-:Slug: flowchart
+:Title: Simple flow chart
+:Tags: Diagrams
 
-This example flowchart uses the positioning-plus library and the
-node-families library by the same author. The fit library is implicitly loaded
-by positioning-plus, the backgrounds library is used to draw stuff behind
-other stuff, the calc library for some coordinate calculations and the
-shapes.geometric library for the ellipse shape.
+With PGF/TikZ you can draw flow charts with relative ease. This flow chart from [1]_
+outlines an algorithm for identifying the parameters of an autonomous underwater vehicle model.
 
-This example was written by Qrrbrbirlbel answering a question on TeX.SE.
+Note that relative node
+placement has been used to avoid placing nodes explicitly. This feature was
+introduced in PGF/TikZ >= 1.09.
+
+.. [1] Bossley, K.; Brown, M. & Harris, C. Neurofuzzy identification of an autonomous underwater vehicle `International Journal of Systems Science`, 1999, 30, 901-913
+
 \end{comment}
-\usetikzlibrary{shapes.geometric,backgrounds,
-  positioning-plus,node-families,calc}
-\tikzset{
-  basic box/.style = {
-    shape = rectangle,
-    align = center,
-    draw  = #1,
-    fill  = #1!25,
-    rounded corners},
-  header node/.style = {
-    Minimum Width = header nodes,
-    font          = \strut\Large\ttfamily,
-    text depth    = +0pt,
-    fill          = white,
-    draw},
-  header/.style = {%
-    inner ysep = +1.5em,
-    append after command = {
-      \pgfextra{\let\TikZlastnode\tikzlastnode}
-      node [header node] (header-\TikZlastnode) at (\TikZlastnode.north) {#1}
-      node [span = (\TikZlastnode)(header-\TikZlastnode)]
-        at (fit bounding box) (h-\TikZlastnode) {}
-    }
-  },
-  hv/.style = {to path = {-|(\tikztotarget)\tikztonodes}},
-  vh/.style = {to path = {|-(\tikztotarget)\tikztonodes}},
-  fat blue line/.style = {ultra thick, blue}
-}
+
 \begin{document}
-\begin{tikzpicture}[node distance = 1.2cm, thick, nodes = {align = center},
-    >=latex]
-  \node[Minimum Width = loop, shape = ellipse, fill = red] (imp-sol)
-    {ellipsoid box};
-  \node[Minimum Width = loop, fill = yellow, below = of imp-sol] (rec-box)
-    {rectangular box, and very wiiiiiiiiiiiiiiide\\2nd line};
-  \node[shift = (left:.5*x_node_dist)] at
-    ($(imp-sol.west|-imp-sol.south)!.5!(rec-box.north west)$) (for-1)
-    {formula 1};
-  \node[shift = (right:.5*x_node_dist)] at
-    ($(imp-sol.east|-imp-sol.south)!.5!(rec-box.north east)$) (for-2)
-    {formula 2};
-  \begin{scope}[on background layer]
-    \node[fit = (for-1)(for-2)(imp-sol)(rec-box), basic box = blue,
-      header = DMFT loop] (dmft-l) {};
-  \end{scope}
-  \path[very thick, blue, hv] (rec-box) edge[->] (for-1) edge[<-] (for-2)
-                              (imp-sol) edge[->] (for-2) edge[<-] (for-1);
+% Define the layers to draw the diagram
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{background,main,foreground}
+\pagestyle{empty}
+\renewcommand{\arraystretch}{0.75}
 
-  \node[east above = of dmft-l, basic box = green, header = DMFT prelude]
-    (dmft-p) {Math and text math and text math and text\\
-              math and text math and text math and text};
-  \node[north left = of dmft-l, basic box = green, header = $\rho$ update,
-     shift = (down:y_node_dist)] (rho)
-    {Much more text much more text\\much more text much more text};
-  \node[basic box = blue, header = DFT part, anchor = north] at
-    (dmft-p.north-|rho) (dft) {So much text so much text so much text\\
-    I think I need \texttt{tikz-lipsum}\\or something like that.};
-  \node[basic box = green, anchor = north] at
-    ($(dft.north east)!.5!(dmft-p.north west)$) (upd) {update\\$math$};
-  \path[fat blue line, <-, dashed, vh] (rho) edge
-    ({$(rho.south)!.5!(dmft-l.south)$}-|dmft-l.south west);
-  \path[fat blue line, ->]
-    ({$(upd.south)!.5!(dmft-p.south)$}-|dmft-p.south west)
-    coordinate (@) edge[<-, solid] coordinate[pos=.15] (@s)
-    coordinate[pos=.9] (@e) (@-|dft.east)
-    {[every edge/.append style=dashed, vh] (@s) edge[<-] (upd) (@e) edge (upd)}
-    (h-rho) edge[dashed] (dft)
-    ($(dmft-p.south)!.5!(dmft-p.south east)$)
-    coordinate (@) edge (@|-dmft-l.north);
+%POSITIVE TERMS:
+%egrep 'befit|winner|encourag|metric|noisy|algorithms$|kernel$|methods$|procedure$|programs$|software$|technologies$|accuracy$|analysis$|benchmark|comparing$|comparison$|comprehensive$|estimation$|evaluation$|effectiveness$|metrics$|efficiency$|performance$|perspective$|quality$|rated$|suitable$|suitability$|superior$|survey$|cputime$|speed$|time$|duration$|fast$|slow$' wordScores.tsv | sort -nr | egrep -v 'mrsfast|mrfast|alignmentmethods' | head -n 5  | perl -lane 'print "{\\tiny $F[0]} & {\\tiny $F[1]} & {\\tiny $F[2]} & {\\tiny $F[3]}\\\\"'
+
+%NEGATIVE TERMS:
+%egrep 'mirnas$|associated$|disease$|genes$|human$|proteins$|genetic$|pathways$|mrna$|cdna$|functional$|clinical$|regulation$|predicted$|interaction$|review$|network' 
+%cat wordScores.tsv | sort -nr | egrep -v 'xxxxxx' | tail -n 5  | perl -lane 'print "{\\tiny $F[0]} & {\\tiny $F[1]} & {\\tiny $F[2]} & {\\tiny $F[3]}\\\\"'
+
+% Define block styles
+\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
+    text width=5em, text centered, rounded corners, minimum height=4em]
+\tikzstyle{line} = [draw, line width=1mm, -latex']
+\tikzstyle{cloud} = [draw, rectangle,fill=red!20, 
+       text width=14em, text centered, rounded corners, minimum height=4em]
+\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
+    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
+
+\begin{tikzpicture}[->,node distance = 6cm, auto,thick,main node/.style={font=\sffamily\Large\bfseries}]
+    % Place nodes
+    \node [align=center]                              (dummy)  {}; 
+%    \node [block]                  (dummy)  {}; 
+    \node [block, left of=dummy,   node distance=4.5cm] (init1)  {Training articles};
+    \node [block, right of=dummy,  node distance=4.5cm] (init2)  {Background articles};
+    \node [cloud, below of=dummy,  node distance=1cm] (remove) {Remove high freq. words};
+    \node [cloud, below of=remove, node distance=2.5cm] (freqs)  {Compute word \& di-word freqs};
+    \node [cloud, below of=freqs,  node distance=4cm] (lods)   {Compute word scores: $s(word)=\log_2\left(\frac{f_{training}(word)+\delta}{f_{background}(word)+\delta}\right)$\\
+\begin{tabular}{llll}
+{\tiny logOdds} & {\tiny tnFreq} & {\tiny bgFreq} & {\tiny word}\\
+{\tiny 5.28}    & {\tiny 0.0019} & {\tiny 0.0000} & {\tiny benchmarking}\\
+{\tiny 5.21}    & {\tiny 0.0061} & {\tiny 0.0002} & {\tiny benchmark}\\
+{\tiny 4.91}    & {\tiny 0.0011} & {\tiny 0.0000} & {\tiny noisy}\\
+{\tiny 4.85}    & {\tiny 0.0022} & {\tiny 0.0001} & {\tiny metrics}\\
+{\tiny 4.85}    & {\tiny 0.0003} & {\tiny 0.0000} & {\tiny encouragingly}\\
+...\\
+{\tiny -7.90}   & {\tiny 0.0000} & {\tiny 0.0024} & {\tiny disease}\\
+{\tiny -8.02}   & {\tiny 0.0000} & {\tiny 0.0026} & {\tiny associated}\\
+{\tiny -8.09}   & {\tiny 0.0000} & {\tiny 0.0027} & {\tiny mirnas}\\
+\end{tabular}
+};
+    \node [cloud, below of=lods, node distance=4cm]     (score)  {Score \& rank candidate articles: $\sum_{word\in article} s(word)$};
+    \node [block, left  of=score, node distance=4.5cm]    (cand)   {Candidate articles};
+    \node [decision, below of=score, node distance=3.5cm] (decide) {Manually evaluate high scoring articles};
+
+    % Draw edges
+    \path [line] (init1)  -- (remove);
+    \path [line] (init2)  -- (remove);
+    \path [line] (remove) -- (freqs);
+    \path [line] (freqs)  -- (lods);
+    \path [line] (lods)  -- (score);
+    \path [line] (cand)  -- (score);
+    \path [line] (score)  -- (decide);
+    \path (decide) edge[bend right=90, line width=1mm] node[anchor=init2,right]{no }(init2);
+    \path (decide) edge[bend  left=90, line width=1mm] node[anchor=init1,right]{yes}(init1);
+
+%    \path (remove.north) +(0,-2.3) node (clouds) {Build model};
+    \path (remove.north) +(-1.5,-2.0) node (clouds) {Build model};
+
+    \begin{pgfonlayer}{background}
+        \path (remove.west |- remove.north)+(-0.5,0.3) node (a) {};
+        \path (lods.south -| lods.east)+(+0.5,-0.3)    node (b) {};
+        \path (clouds.east |- clouds.east)+(+3.5,-7.8)   node (c) {};
+
+        \path[fill=yellow!20,rounded corners, draw=black!50, dashed]
+            (a) rectangle (c);           
+        \path (remove.north west)+(-0.2,0.2) node (a) {};
+
+    \end{pgfonlayer}
+
 \end{tikzpicture}
+
 \end{document}

### 2015-11-04 bin/pmArticleScore.pl

@@ -10,7 +10,11 @@
 #      Basic approach: 
 #                      TRAINING
 #                      1. build a dictionary for each training article (hash?)
-#                      2. exclude high frequency words (possibly use the further background/non-scientific text for this e.g. Alice in Wonderland)
+#                      2. exclude high frequency words (possibly use the further background/non-scientific text for this e.g. Alice in Wonderland & The Hobbit)
+#                                 --add a list of words to ignore e.g. methods, sub-disciplines of bioinformatics
+#                                 --echo -e "assemblers\nmappers\nmapper\nassembler\ngabenchtob\nclustal\nenme\nsimprot\ncuresim\ncoffee\nseal\nsmidgen\nmsbi\nbrat\nidba\nfirefly\nsoapv\nreas\nhartigan\nhalle\ngsas\ncaap" | sort -d  > ignore.tsv
+#                                 --cat meanRankSpeedData | cut -f 4 | grep -v method >>ignore.tsv
+#
 #                      3. score each remaining word based upon how frequently it is used in the training data relative to the background set 
 #                         (e.g. log2( [f(t.word)/t.size + d]/[f(b.word)/b.size + d] ) ...  )
 #                      
@@ -22,7 +26,7 @@
 #PUBMED SEARCHES:
 #                      TRAINING
 #                      less Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv | perl -lane '@F=split(/\t/); print "$F[0]\[uid\] OR " if $F[0]>0' | tr -d "\n"
-#                      17151342[uid] OR 20047664[uid] OR 25198770[uid] OR 21483869[uid] OR 24526711[uid] OR 24839440[uid] OR 17062146[uid] OR 21423806[uid] OR 25511303[uid] OR 20617200[uid] OR 26475308[uid] OR 25521762[uid] OR 20375457[uid] OR 23593445[uid] OR 21525877[uid] OR 24708189[uid] OR 18287116[uid] OR 24602402[uid] OR 24086547[uid] OR 18793413[uid] OR 23393030[uid] OR 22132132[uid] OR 15701525[uid] OR 22152123[uid] OR 19046431[uid] OR 25760244[uid] OR 23758764[uid]
[…]

### 2015-11-04 bin/prettyPlot.R

@@ -83,6 +83,9 @@ text( 1.2,-0.2, "fast+inaccurate",pos=4, cex=0.5)
 lines(c(0,1),reg1c[1]+reg1c[2]*c(0,1), lwd=5)
 xr<-0.2; text(xr, reg1c[1]+reg1c[2]*xr, "regression",pos=3, cex=0.8)
 
+#lowess
+lines(lowess(d$accuracyRank, d$speedRank, f = .2), col = 2, lwd=5)
+
 #DIAGONALS
 #lines(c(0,1),c(0,1),lty=2)
 lines(c(0,1),c(1,0),lty=2, lwd=3)
@@ -96,11 +99,11 @@ lines(c(1,0),c(0,0),lty=1,lwd=2)#bottom
 }
 
 
-png(filename="../figures/Figure1.png", width = 720, height = 360)
+pdf(file=    "../figures/Figure1.pdf", width = 10, height = 5)
 plotMe()
 dev.off()
 
-pdf(file=    "../figures/Figure1.pdf", width = 10, height = 5)
+png(filename="../figures/Figure1.png", width = 720, height = 360)
 plotMe()
 dev.off()
 

### 2015-11-04 bin/tsv2data.pl

@@ -33,6 +33,7 @@ my ($testId,$pmid,$accuracySource,$accuracyMetric,$speedSource)=("","","","","")
 
 while(my $in = <IN>){
     chomp($in);
+    next if $in =~ /^pubmedID/; 
     $in =~  s/\r//g;
     my @in = split(/\t/, $in); 
     if (isNumeric($in[8])){
@@ -101,7 +102,7 @@ foreach my $meth (keys %ranks){
 close(UT);
 close(UT0);
 
-system("R CMD BATCH --no-save prettyPlot.R");
+system("R CMD BATCH --no-save ../bin/prettyPlot.R");
 
 exit(0);
 

### 2015-11-05 bin/pmArticleScore.pl
Updated viz. Adding method data to analysis.

@@ -12,8 +12,8 @@
 #                      1. build a dictionary for each training article (hash?)
 #                      2. exclude high frequency words (possibly use the further background/non-scientific text for this e.g. Alice in Wonderland & The Hobbit)
 #                                 --add a list of words to ignore e.g. methods, sub-disciplines of bioinformatics
-#                                 --echo -e "assemblers\nmappers\nmapper\nassembler\ngabenchtob\nclustal\nenme\nsimprot\ncuresim\ncoffee\nseal\nsmidgen\nmsbi\nbrat\nidba\nfirefly\nsoapv\nreas\nhartigan\nhalle\ngsas\ncaap" | sort -d  > ignore.tsv
-#                                 --cat meanRankSpeedData | cut -f 4 | grep -v method >>ignore.tsv
+#                                 --echo -e "assemblers\nmappers\nmapper\nassembler\ngabenchtob\nclustal\nenme\nsimprot\ncuresim\ncoffee\nseal\nsmidgen\nmsbi\nbrat\nidba\nfirefly\nsoapv\nreas\nhartigan\nhalle\ngsas\ncaap\nsegmod\nencad" | sort -d  > ignore.tsv
+#                                 --cat meanRankSpeedData | cut -f 4 | grep -v method | perl -lane 's/[\s+\d+\.,;:\!\?\&\$\@\%\=\|\"()\[\]\-\/\_\*]//g; print;' >>ignore.tsv
 #
 #                      3. score each remaining word based upon how frequently it is used in the training data relative to the background set 
 #                         (e.g. log2( [f(t.word)/t.size + d]/[f(b.word)/b.size + d] ) ...  )
@@ -26,7 +26,7 @@
 #PUBMED SEARCHES:
 #                      TRAINING
 #                      less Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv | perl -lane '@F=split(/\t/); print "$F[0]\[uid\] OR " if $F[0]>0' | tr -d "\n"
[…]

### 2015-11-05 bin/prettyPlot.R

@@ -1,15 +1,66 @@
-#R CMD BATCH prettyPlot.R
+#R CMD BATCH ../bin/prettyPlot.R
 library(MASS)  # in case it is not already loaded 
 library(RColorBrewer)
 library(fields)
 k <- 11
 my.cols <- rev(brewer.pal(k, "RdYlBu"))
-d<-read.table("meanRankSpeedData", header=T)
-reg1  <- lm(d$speedRank~d$accuracyRank)
-reg1c <- coefficients(reg1)
+d <-read.table("meanRankSpeedData.tsv", header=T)
+dr<-read.table( "rawRankSpeedData.tsv", header=T)
+#reg1  <- lm(d$speedRank~d$accuracyRank)
+#reg1c <- coefficients(reg1)
 
-dr<-read.table("rawRankSpeedData", header=T)
+plotMe <- function() {
+       
+op<-par(mfrow=c(1,2),cex=1.1)
 
+
+
+# plot(c(0,1),c(1,0), xlim=c(1,0),ylim=c(1,0),xlab="normalised accuracy rank", ylab="normalised speed rank",type="l",lty=2,lwd=3)
+
+# #DIAGONALS
+# #lines(c(0,1),c(0,1),lty=2)
+# #lines(c(0,1),c(1,0),lty=2, lwd=3)
+
+# #BOX
+# lines(c(0,0),c(0,1),lty=1,lwd=2)#left
+# lines(c(0,1),c(1,1),lty=1,lwd=2)#top
+# lines(c(1,1),c(1,0),lty=1,lwd=2)#right
+# lines(c(1,0),c(0,0),lty=1,lwd=2)#bottom
+
+# for(id in unique(dr$testId)){       
+#        regr1  <- lm(dr$normSpeedRank[ dr$testId == id]~dr$normAccuracyRank[ dr$testId == id])
+#        regr1c <- coefficients(regr1)
+#        if( regr1c[1]+regr1c[2] > 0 ){
+# 	   lines(c(0,1),regr1c[1]+regr1c[2]*c(0,1), lwd=5)
+#        }
+#        else{#y negative when x = 0
+#             #y = mx + c = 0
+# 	    #  =>     x = -c/m
+# 	   lines(c(0, -1*regr1c[1]/regr1c[2] ),  c(regr1c[1], 0), lwd=5)	
+# 	   cat("Y<0!: ",id, "\n")
[…]

### 2015-11-05 bin/tsv2data.pl

@@ -3,10 +3,18 @@
 use warnings;
 use strict;
 
-open(IN, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv"); 
-open(UT0, "> rawRankSpeedData");
-print UT0 "testId\tnormAccuracyRank\tnormSpeedRank\tmethod\n";
+#Read 
+open(IN0, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv"); 
+while(my $in = <IN0>){
+    next if $in =~ /^pubmedID/; 
+    chomp($in);
+    $in =~  s/\r//g;
+    my @in = split(/\t/, $in); 
 
+
+}
+close(IN0);
+######################################################################
 #echo -ne "accuracyRank\tspeedRank\tnumMethods\n" > data && cut -f 7,8,9 Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv | grep -v N | tr -d "=" | perl -lane 'if(/^acc|^N/ or $F[0] !~ /\d+/ or $F[1] !~ /\d+/){next}elsif(defined($F[2])){$max=$F[2]} printf "%0.2f\t%0.2f\t$max\n", ($F[0]-1)/($max-1), ($F[1]-1)/($max-1); ' >> data
     #  1	pubmedID
     #  2	Title
@@ -21,9 +29,11 @@ print UT0 "testId\tnormAccuracyRank\tnormSpeedRank\tmethod\n";
     # 11	Bias
 
 
+#ADD NEW METHODS TO "methodInfo" SHEET:
 #cat Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv | cut -f 1 | sort -d | perl -lane 'print "$F[0] => 1, "' | tr -d "\n"
[…]

### 2015-11-05 bin/prettyPlot.R
Added Spearmans rho to Fig.1. Added code to do it.

@@ -11,7 +11,7 @@ dr<-read.table( "rawRankSpeedData.tsv", header=T)
 
 plotMe <- function() {
        
-op<-par(mfrow=c(1,2),cex=1.1)
+op<-par(mfrow=c(1,2),cex=1.1,las=2)
 
 
 
@@ -41,22 +41,36 @@ op<-par(mfrow=c(1,2),cex=1.1)
 #        }
 # }
 
-smoothScatter(dr$accuracyRank, dr$speedRank, nbin=1000, nrpoints=0, colramp=colorRampPalette(my.cols), pch=19, cex=.8, xlab="normalised accuracy rank", ylab="normalised speed rank",xlim=c(1.2,-0.2),ylim=c(1.2,-0.2)) 
-lines(lowess(dr$accuracyRank, dr$speedRank, f = .2), col = 2, lwd=5)
+#smoothScatter(dr$accuracyRank, dr$speedRank, nbin=1000, nrpoints=0, colramp=colorRampPalette(my.cols), pch=19, cex=.8, xlab="normalised accuracy rank", ylab="normalised speed rank",xlim=c(1.2,-0.2),ylim=c(1.2,-0.2)) 
+#lines(lowess(dr$accuracyRank, dr$speedRank, f = .2), col = 2, lwd=5)
+#boxit()
 
-boxit()
+######################################################################
+#sumRanks	accuracyRank	speedRank	method	numTests	yearPublished	IF	H5	cites	hindex	mindex
+cites <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$cites), method = "spearman")
+IF    <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$IF), method = "spearman")
+H5    <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$H5), method = "spearman")
+hindex<-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$hindex), method = "spearman")
+mindex<-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$mindex), method = "spearman")
+
+barplot(t(c(hindex$estimate, mindex$estimate, cites$estimate, H5$estimate, IF$estimate )), names=c("Corr. Author\nH-index", "Corr. Auth\nM-index", 'journal.H5', 'tot.Cites', 'journal.IF'), ylab="Spearman's rho",ylim=c(-1,1))
+lines(c(-100,100),c(0,0))
+
 
 par(mar = c(5,4,4,5) + .1)
[…]

### 2015-11-05 bin/tsv2data.pl

@@ -3,17 +3,92 @@
 use warnings;
 use strict;
 
-#Read 
-open(IN0, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv"); 
+#Read method info:
+my %methodInfo; 
+my @methodInfoKeys = qw(yearPublished IF H5 cites hindex mindex); 
+# head -n 1 Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv | tr "\t" "\n" | nl
+#      1	method
+#      2	yearPublished
+#      3	journal
+#      4	impactFactor
+#      5	Journal H5-index
+#      6	totalCitations
+#      7	Corresponding author: H-index
+#      8	Corresponding author: M-index
+#      9	fullCite
+open(IN0, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv");
 while(my $in = <IN0>){
-    next if $in =~ /^pubmedID/; 
+    next if $in =~ /^method/; 
     chomp($in);
     $in =~  s/\r//g;
     my @in = split(/\t/, $in); 
+    
+    if(defined($in[0]) && length($in[0]) > 1){
+	
+	if(defined($in[1])){
+	    my @yrs = split(/;\s*/, $in[1]);
+	    $methodInfo{$in[0]}{'yearPublished'}='NA';
+	    $methodInfo{$in[0]}{'yearPublished'}=minA(@yrs) if ($in[1] !~ 'NA'); #using the first publication
+	}
+	
+	if(defined($in[3])){
+	    my @ifs = split(/;\s*/, $in[3]);	    
+	    $methodInfo{$in[0]}{'IF'}='NA';
+	    $methodInfo{$in[0]}{'IF'}=maxA(@ifs) if ($in[3] !~ 'NA'); #using the highest impact factor
+	}
+	
+	if(defined($in[4])){
+	    my @h5s = split(/;\s*/, $in[4]);
+	    $methodInfo{$in[0]}{'H5'}='NA';
+	    $methodInfo{$in[0]}{'H5'}=maxA(@h5s) if ($in[4] !~ 'NA'); #using the highest impact factor
+	}
+	
+	if(defined($in[5])){
+	    my @cites = split(/;\s*/, $in[5]);
+	    $methodInfo{$in[0]}{'cites'}='NA';
[…]

### 2015-11-06 bin/prettyPlot.R
Added relative age and relative #cites to analysis.

@@ -47,22 +47,28 @@ op<-par(mfrow=c(1,2),cex=1.1,las=2)
 
 ######################################################################
 #sumRanks	accuracyRank	speedRank	method	numTests	yearPublished	IF	H5	cites	hindex	mindex
-cites <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$cites), method = "spearman")
-IF    <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$IF), method = "spearman")
-H5    <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$H5), method = "spearman")
-hindex<-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$hindex), method = "spearman")
-mindex<-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$mindex), method = "spearman")
+cites   <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$cites),    method = "spearman")
+IF      <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$IF),       method = "spearman")
+H5      <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$H5),       method = "spearman")
+hindex  <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$hindex),   method = "spearman")
+mindex  <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$mindex),   method = "spearman")
+relAge  <-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$relAge),   method = "spearman")
+relCites<-cor.test(d$accuracyRank+d$speedRank, as.numeric(d$relCites), method = "spearman")
 
-barplot(t(c(hindex$estimate, mindex$estimate, cites$estimate, H5$estimate, IF$estimate )), names=c("Corr. Author\nH-index", "Corr. Auth\nM-index", 'journal.H5', 'tot.Cites', 'journal.IF'), ylab="Spearman's rho",ylim=c(-1,1))
[…]

### 2015-11-06 bin/tsv2data.pl

@@ -3,6 +3,9 @@
 use warnings;
 use strict;
 
+
+#TODO: add relative age & relative cites & relative IF/H5, & relative corresponding author impact (H/M), ??????? 
+
 #Read method info:
 my %methodInfo; 
 my @methodInfoKeys = qw(yearPublished IF H5 cites hindex mindex); 
@@ -19,6 +22,7 @@ my @methodInfoKeys = qw(yearPublished IF H5 cites hindex mindex);
 open(IN0, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv");
 while(my $in = <IN0>){
     next if $in =~ /^method/; 
+    $in =~ s/[^[:ascii:]]//g; #strip of fucking unicode chars
     chomp($in);
     $in =~  s/\r//g;
     my @in = split(/\t/, $in); 
@@ -29,6 +33,7 @@ while(my $in = <IN0>){
 	    my @yrs = split(/;\s*/, $in[1]);
 	    $methodInfo{$in[0]}{'yearPublished'}='NA';
 	    $methodInfo{$in[0]}{'yearPublished'}=minA(@yrs) if ($in[1] !~ 'NA'); #using the first publication
+	    #printf "[$in[0]]\tyrsArray[@yrs]\tyearStr[$in[1]]\tminA[%d]\n", minA(@yrs);
 	}
 	
 	if(defined($in[3])){
@@ -114,17 +119,22 @@ close(IN0);
 my ($max,$numBench)=(1,0);
 my %ranks;
 my %methodCounts;
+my %benchMethod;
 my ($testId,$pmid,$accuracySource,$accuracyMetric,$speedSource)=("","","","","");
 
 open(IN1, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv"); 
 open(UT0, "> rawRankSpeedData.tsv");
 print UT0 "testId\taccuracyRank\tspeedRank\tmethod\n";
-
 while(my $in = <IN1>){
     next if $in =~ /^pubmedID/; 
+    $in =~ s/[^[:ascii:]]//g; #strip of fucking unicode chars
     chomp($in);
     $in =~  s/\r//g;
     my @in = split(/\t/, $in); 
+    #fix method name:
+    $in[5] =~  s/[ -=\/0-9]//g;
+    $in[5]=lc($in[5]);
+    
     if (isNumeric($in[8])){
         $max=$in[8];
         $numBench++;
[…]

### 2015-12-17 figures/Figure2-flowDiagram.tex
Did a lot of stuff. Updated analysis. Regularised IFs.

@@
 \tikzstyle{block} = [rectangle, draw, fill=blue!20, 
-    text width=5em, text centered, rounded corners, minimum height=4em]
+    text width=12em, text centered, rounded corners, minimum height=4em]
 \tikzstyle{line} = [draw, line width=1mm, -latex']
 \tikzstyle{cloud} = [draw, rectangle,fill=red!20, 
-       text width=14em, text centered, rounded corners, minimum height=4em]
+       text width=32em, text centered, rounded corners, minimum height=4em]
 \tikzstyle{decision} = [diamond, draw, fill=blue!20, 
-    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
+    text width=9em, text badly centered, node distance=3cm, inner sep=0pt]
 
-\begin{tikzpicture}[->,node distance = 6cm, auto,thick,main node/.style={font=\sffamily\Large\bfseries}]
+\begin{tikzpicture}[->,node distance = 20cm, auto,thick,main node/.style={font=\sffamily\Large\bfseries}]
     % Place nodes
@@
 %    \node [block]                  (dummy)  {}; 
-    \node [block, left of=dummy,   node distance=4.5cm] (init1)  {Training articles};
-    \node [block, right of=dummy,  node distance=4.5cm] (init2)  {Background articles};
-    \node [cloud, below of=dummy,  node distance=1cm] (remove) {Remove high freq. words};
-    \node [cloud, below of=remove, node distance=2.5cm] (freqs)  {Compute word \& di-word freqs};
-    \node [cloud, below of=freqs,  node distance=4cm] (lods)   {Compute word scores: $s(word)=\log_2\left(\frac{f_{training}(word)+\delta}{f_{background}(word)+\delta}\right)$\\
+    \node [block, left of=dummy,   node distance=9cm] (init1)  {\Huge Training articles};
+    \node [block, right of=dummy,  node distance=9cm] (init2)  {\Huge Background articles};
+    \node [cloud, below of=dummy,  node distance=2cm] (remove)   {\Huge Remove high freq. words};
+    \node [cloud, below of=remove, node distance=3cm] (freqs)  {\Huge Compute word \& di-word freqs};
+    \node [cloud, below of=freqs,  node distance=8cm] (lods)     {\Huge Compute word scores: $lo(word)=\log_2\left(\frac{f_{training}(word)+\delta}{f_{background}(word)+\delta}\right)$\\
 \begin{tabular}{llll}
-{\tiny logOdds} & {\tiny tnFreq} & {\tiny bgFreq} & {\tiny word}\\
-{\tiny 5.28}    & {\tiny 0.0019} & {\tiny 0.0000} & {\tiny benchmarking}\\
-{\tiny 5.21}    & {\tiny 0.0061} & {\tiny 0.0002} & {\tiny benchmark}\\
-{\tiny 4.91}    & {\tiny 0.0011} & {\tiny 0.0000} & {\tiny noisy}\\
-{\tiny 4.85}    & {\tiny 0.0022} & {\tiny 0.0001} & {\tiny metrics}\\
-{\tiny 4.85}    & {\tiny 0.0003} & {\tiny 0.0000} & {\tiny encouragingly}\\
+{\normalsize logOdds} & {\normalsize tnFreq} & {\normalsize bgFreq} & {\normalsize word}\\
+{\normalsize 5.28}    & {\normalsize 0.0019} & {\normalsize 0.0000} & {\normalsize benchmarking}\\
+{\normalsize 5.21}    & {\normalsize 0.0061} & {\normalsize 0.0002} & {\normalsize benchmark}\\
+{\normalsize 4.91}    & {\normalsize 0.0011} & {\normalsize 0.0000} & {\normalsize noisy}\\
+{\normalsize 4.85}    & {\normalsize 0.0022} & {\normalsize 0.0001} & {\normalsize metrics}\\
+{\normalsize 4.85}    & {\normalsize 0.0003} & {\normalsize 0.0000} & {\normalsize encouragingly}\\
 ...\\
-{\tiny -7.90}   & {\tiny 0.0000} & {\tiny 0.0024} & {\tiny disease}\\
-{\tiny -8.02}   & {\tiny 0.0000} & {\tiny 0.0026} & {\tiny associated}\\
-{\tiny -8.09}   & {\tiny 0.0000} & {\tiny 0.0027} & {\tiny mirnas}\\
+{\normalsize -7.90}   & {\normalsize 0.0000} & {\normalsize 0.0024} & {\normalsize disease}\\
+{\normalsize -8.02}   & {\normalsize 0.0000} & {\normalsize 0.0026} & {\normalsize associated}\\
+{\normalsize -8.09}   & {\normalsize 0.0000} & {\normalsize 0.0027} & {\normalsize mirnas}\\
 \end{tabular}
 };
-    \node [cloud, below of=lods, node distance=4cm]     (score)  {Score \& rank candidate articles: $\sum_{word\in article} s(word)$};
-    \node [block, left  of=score, node distance=4.5cm]    (cand)   {Candidate articles};
-    \node [decision, below of=score, node distance=3.5cm] (decide) {Manually evaluate high scoring articles};
+    \node [cloud, below of=lods, node distance=8cm]     (score)  {\Huge Score \& rank candidate articles: $\sum_{i} lo(w_i)$};
+    \node [block, left  of=score, node distance=10cm]    (cand)   {\Huge Candidate articles};
+    \node [decision, below of=score, node distance=7cm] (decide) {\Huge Manually evaluate high scoring articles};
 
@@
     \path [line] (score)  -- (decide);
-    \path (decide) edge[bend right=90, line width=1mm] node[anchor=init2,right]{no }(init2);
-    \path (decide) edge[bend  left=90, line width=1mm] node[anchor=init1,right]{yes}(init1);
+    \path (decide) edge[bend right=90, line width=1mm] node[anchor=init2,right]{\Huge no }(init2);
+    \path (decide) edge[bend  left=90, line width=1mm] node[anchor=init1,right]{\Huge yes}(init1);
 
-%    \path (remove.north) +(0,-2.3) node (clouds) {Build model};
-    \path (remove.north) +(-1.5,-2.0) node (clouds) {Build model};
+    \path (remove.north) +(7,-2.0) node[rotate=-90] (clouds) {\Huge Build model};
 
     \begin{pgfonlayer}{background}
-        \path (remove.west |- remove.north)+(-0.5,0.3) node (a) {};
-        \path (lods.south -| lods.east)+(+0.5,-0.3)    node (b) {};
-        \path (clouds.east |- clouds.east)+(+3.5,-7.8)   node (c) {};
+        \path (remove.west |- remove.north)+(-1,0.3) node (a) {};
+        \path (lods.south -| lods.east)+(-0.5,-0.3)    node (b) {};
+        \path (clouds.east |- clouds.east)+(+1.75,-17.8)   node (c) {};
 

### 2015-12-17 bin/pmArticleScore.pl

@@ -97,7 +97,7 @@ if(defined($fictionFile)){
 
 
 #read training data, exclude common words, compute word frequencies for the remainder: 
-open(TD, "< $trainingFile"); 
+open(TD, "< $trainingFile") or die "[missing training file]: [$trainingFile]\n[$!]"; 
 my @training = <TD>;
 close(TD);
 my %empty;
@@ -283,7 +283,7 @@ sub filterCommonWords {
     my $sum = 0;
     foreach my $word (keys %words){
 	next if (defined($commonWordDict{$word}));
-	next if (length($word) < 4);#filter short words
+	#next if (length($word) < 4);#filter short words
 	next if (defined($skipWords{$word}));
         #next if ($words{$word} < 1);#FILTER WORDS ONLY SEEN ONCE -- MAY REVISIT THIS!!!
 	$filteredTrainingWords{$word} = $words{$word}; 

### 2015-12-17 bin/prettyPlot.R

@@ -1,4 +1,5 @@
 #R CMD BATCH ../bin/prettyPlot.R
+
 library(MASS)  # in case it is not already loaded 
 library(RColorBrewer)
 library(fields)
@@ -9,57 +10,54 @@ dr<-read.table( "rawRankSpeedData.tsv", header=T)
 #reg1  <- lm(d$speedRank~d$accuracyRank)
 #reg1c <- coefficients(reg1)
 
-plotMe <- function() {
-       
-op<-par(mfrow=c(1,2),cex=1.1,las=2)
-
-
-
-# plot(c(0,1),c(1,0), xlim=c(1,0),ylim=c(1,0),xlab="normalised accuracy rank", ylab="normalised speed rank",type="l",lty=2,lwd=3)
-
-# #DIAGONALS
-# #lines(c(0,1),c(0,1),lty=2)
-# #lines(c(0,1),c(1,0),lty=2, lwd=3)
-
-# #BOX
-# lines(c(0,0),c(0,1),lty=1,lwd=2)#left
-# lines(c(0,1),c(1,1),lty=1,lwd=2)#top
-# lines(c(1,1),c(1,0),lty=1,lwd=2)#right
-# lines(c(1,0),c(0,0),lty=1,lwd=2)#bottom
-
-# for(id in unique(dr$testId)){       
-#        regr1  <- lm(dr$normSpeedRank[ dr$testId == id]~dr$normAccuracyRank[ dr$testId == id])
-#        regr1c <- coefficients(regr1)
-#        if( regr1c[1]+regr1c[2] > 0 ){
-# 	   lines(c(0,1),regr1c[1]+regr1c[2]*c(0,1), lwd=5)
-#        }
-#        else{#y negative when x = 0
-#             #y = mx + c = 0
-# 	    #  =>     x = -c/m
-# 	   lines(c(0, -1*regr1c[1]/regr1c[2] ),  c(regr1c[1], 0), lwd=5)	
-# 	   cat("Y<0!: ",id, "\n")
-#        }
-# }
-
-#smoothScatter(dr$accuracyRank, dr$speedRank, nbin=1000, nrpoints=0, colramp=colorRampPalette(my.cols), pch=19, cex=.8, xlab="normalised accuracy rank", ylab="normalised speed rank",xlim=c(1.2,-0.2),ylim=c(1.2,-0.2)) 
-#lines(lowess(dr$accuracyRank, dr$speedRank, f = .2), col = 2, lwd=5)
-#boxit()
-
 ######################################################################
[…]

### 2015-12-17 bin/tsv2data.pl

@@ -229,7 +229,7 @@ open(UT, "> meanRankSpeedData.tsv");
 my $methInfo = join("\t", @methodInfoKeys); 
 print UT "sumRanks\taccuracyRank\tspeedRank\tmethod\tnumTests\t$methInfo\trelAge\trelCites\n";
 
-foreach my $meth (keys %ranks){ 
+foreach my $meth (sort keys %ranks){ 
     
     printf UT "%0.2f\t%0.2f\t%0.2f\t%s\t%d", $ranks{$meth}[0]/$ranks{$meth}[2] + $ranks{$meth}[1]/$ranks{$meth}[2],  $ranks{$meth}[0]/$ranks{$meth}[2], $ranks{$meth}[1]/$ranks{$meth}[2], $meth, $ranks{$meth}[2];     
     

### 2016-06-16 bin/prettyPlot.R
Added new viz. Permutation tests.

@@ -5,6 +5,8 @@
 library(MASS)  # in case it is not already loaded 
 library(RColorBrewer)
 library(fields)
+library(gplots)
+
 k <- 11
 my.cols <- rev(brewer.pal(k, "RdYlBu"))
 d <-read.table("meanRankSpeedData.tsv", header=T)
@@ -17,233 +19,16 @@ regM <- lm(accuracyRank ~ speedRank + IF + H5 + cites + hindex + mindex + relAge
 summary(regM)
 
 ######################################################################
-testCorners <- function(){
-#Test all corners: Fast & inaccurate, slow & accurate, ...
-cat("Test all corners")
-for(delta in seq(0,0.6,by=0.005)){
-#delta<-0.25
-          fANDa <- length(d$accuracyRank[(d$speedRank <=   delta) & (d$accuracyRank <=   delta)])
-          fANDi <- length(d$accuracyRank[(d$speedRank <=   delta) & (d$accuracyRank >= 1-delta)])
-          sANDa <- length(d$accuracyRank[(d$speedRank >= 1-delta) & (d$accuracyRank <=   delta)])
-          sANDi <- length(d$accuracyRank[(d$speedRank >= 1-delta) & (d$accuracyRank >= 1-delta)])
-
-          corners <-
-          matrix(c(sANDa, fANDa, sANDi, fANDi), nrow = 2, dimnames =
-                          list(c("Slow", "Fast"),
-                          c("Accurate", "Inaccurate")))
-corners
-	  ft <- fisher.test(corners)#, alternative = "less") 
-          cs <-  chisq.test(corners)
-          if(ft$p < 0.1 || cs$p.value < 0.1){
-                    cat(delta,"\tfet.P = ",ft$p, "\todds-ratio =", ft$estimate, "\tcst.P = ",cs$p.value, "\tX-squared =", cs$statistic, "\tN =", sANDa+fANDa+sANDi+fANDi,"\n")
-          }
-}
-
-cat("Fast & Inaccurate")
-#Fast & Inaccurate vs the rest?
-for(delta in seq(0,0.6,by=0.005)){
-#delta<-0.25
-         nfANDni <- length(d$accuracyRank[(d$speedRank >   delta) & (d$accuracyRank  < 1-delta)])
[…]

### 2016-06-17 bin/tsv2data.pl
Added permutation tests to the tsv2data script.

@@ -119,9 +119,9 @@ close(IN0);
 my ($max,$numBench)=(1,0);
 my %ranks;
 my %methodCounts;
-my %benchMethod;
 my ($testId,$pmid,$accuracySource,$accuracyMetric,$speedSource)=("","","","","");
-my (@permAccuracy, @permSpeed) = ((),());
+my (%benchAccuracy, %benchSpeed, %benchMethod);
+
 
 open(IN1, "< Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv"); 
 open(UT0, "> rawRankSpeedData.tsv");
@@ -169,23 +169,30 @@ while(my $in = <IN1>){
 	
         if($pmid && length($accuracySource) && length($accuracyMetric) && length($speedSource) ){
 	    $testId = "$pmid:$accuracySource:$accuracyMetric:$speedSource";
+	    
+	    my $accNorm = ($in[6]-1)/($max-1);
+	    my $spdNorm = ($in[7]-1)/($max-1);
 	    push(@{$benchMethod{$testId}}, $in[5]);
-	}
-	        
-        if (not defined $ranks{$in[5]}){
-            #sum(accuracy) sum(speed) numEntries whichBenchmark
-            $ranks{$in[5]} = [0.0,0.0, 0, 0]; 
-            $numBench++;
-        }
-
-        #ranks: key: methodName, 0: sum of normalised accuracy rank, 1: sum of normalised speed rank, 2: number of times method benchmarked, 3: benchmark number
-        $ranks{$in[5]}[0] += ($in[6]-1)/($max-1); #normalised accuracy rank
-        $ranks{$in[5]}[1] += ($in[7]-1)/($max-1); #normalised speed rank
-        $ranks{$in[5]}[2]++;
-        $ranks{$in[5]}[3]  = $numBench;
+	    push(@{$benchAccuracy{$testId}}, $accNorm);
+	    push(@{$benchSpeed{   $testId}}, $spdNorm);
 	
-	#            testId   normAccuracyRank   normSpeedRank   method
-	printf UT0 "$testId\t%0.3f\t%0.3f\t$in[5]\n", $ranks{$in[5]}[0], $ranks{$in[5]}[1];#($in[6]-1)/($max-1), ($in[7]-1)/($max-1);
+	        
+	    if (not defined $ranks{$in[5]}){
[…]

### 2016-06-17 manuscript/supplementary.tex
Added permutation tsv files & supplementary document.

\documentclass[fleqn,10pt]{SelfArx} % Document font size and equations flushed left

\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\graphicspath{{../figures/}}

\definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
\definecolor{color2}{RGB}{0,20,20} % Color of the boxes behind the abstract and headings

\JournalInfo{Supplementary Text} % Journal information ``Journal, Vol. XXI, No. 1, 1-5, 2015''
\Archive{ } % Additional notes (e.g. copyright, DOI, review/research article)

\PaperTitle{Supplementary results for: A meta-analysis of computational biology benchmarks reveals publication bias affects on speed and accuracy} % Article title

\Authors{Paul P. Gardner\textsuperscript{1,2}*, James Paterson\textsuperscript{1,2}, Fatemeh Ashari Ghomi\textsuperscript{1,2}, Sinan Uğur Umu\textsuperscript{1,2}, Stephanie McGimpsey\textsuperscript{1,2}} % Authors
\affiliation{\textsuperscript{1}\textit{School of Biological Sciences, University of Canterbury, Christchurch, New Zealand.}} % Author affiliation
\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre and the Bio-Protection Research Centre, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{*\textbf{Corresponding author}: paul.gardner@canterbury.ac.nz} % Corresponding author

\Keywords{computational biology --- accuracy --- benchmarks --- meta-analysis --- software development} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{.}

\begin{document}

\flushbottom % Makes all text pages the same height
\maketitle % Print the title and abstract box
%\tableofcontents % Print the contents section

\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section*{Introduction} % The \section*{} command stops section numbering

XXXX

\clearpage
\newpage

\onecolumn

\begin{figure}[H]
\centering
\includegraphics[width=0.55\textwidth]{supplementary-figures-small.pdf}
\caption{The distributions for the metrics used in this study. These
are, reading from left to right, top to bottom: Accuracy -- the mean
normalised accuracy rank for each benchmarked method; Speed -- the
mean normalised speed rank for each benchmarked method; Cites -- the number of
citations to the most cited manuscript describing a method, data from GoogleScholar;
JIF -- the Journal Impact Factor to the highest impact journal that has published
a manuscript describing a method, data from 2014 Thompson-Reuters Citation Reports;
JH5 -- the Journal H5 index to the highest impact journal that has published
a manuscript describing a method, data from GoogleScholar 2015 Metrics;
H-index -- the H-index for the highest profile corresponding author from the
manuscripts describing a method, data from GoogleScholar User Profiles;
M-index -- the M-index (H-index/(\#years since first publication)) for the
highest profile corresponding author from the manuscripts describing a method,
data from GoogleScholar User Profiles;}
\label{fig:S1}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.5\textwidth]{smoothScatter-speed-vs-accuracy.pdf}
\caption{...}
\label{fig:S2}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.85\textwidth]{smoothScatters.pdf}
\caption{...}
\label{fig:S3}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{relAge-speedAcc.pdf}
\includegraphics[width=0.45\textwidth]{relAge-SpeedVsAccuracy-heatmap.pdf}\\
\includegraphics[width=0.45\textwidth]{hindex-SpeedVsAccuracy-heatmap.pdf}
\caption{...}
\label{fig:S4}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.85\textwidth]{wordScores.pdf}
\caption{...}
\label{fig:S5}
\end{figure}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\bibliographystyle{unsrt}
\bibliography{paulall}

\end{document}

### 2016-06-23 manuscript/supplementary.tex
Increased the number of permutations. Reordered R code.

@@
 \onecolumn
+
+\section*{R code for Figure~1A\&B}
+
+{\tiny
+\begin{verbatim}
+
+#read data
+d <-read.table("meanRankSpeedData.tsv", header=T)
+
+#initialize matrices
+dNames <- c("IF", "H5", "relAge",       "yearPublished", "cites", "relCites",       "mindex",  "hindex",  "accuracyRank",  "speedRank" )
+pNames <- c("JIF", "JH5", "Rel. age", "Year",          "Cites", "Rel. cites", "M-index", "H-index", "Accuracy",      "Speed")
+pvalMatrix<-matrix(1, length(dNames), length(dNames))
+rhoMatrix <-matrix(0, length(dNames), length(dNames))
+sigMatrix <-matrix("",length(dNames), length(dNames))
+
+colnames(pvalMatrix)<-pNames
+rownames(pvalMatrix)<-pNames
+colnames(rhoMatrix) <-pNames
+rownames(rhoMatrix) <-pNames
+colnames(sigMatrix) <-pNames
+rownames(sigMatrix) <-pNames
+
+#loop through pairwise combinations, record rho and P-values 
+for(i in 1:length(dNames)){
+      for(j in 1:length(dNames)){
+	    spear<-cor.test(d[,dNames[i] == colnames(d)], d[,dNames[j] == colnames(d)], method = "spearman", exact = T)   #, alternative = "less")
+	    pvalMatrix[i,j] <- spear$p.value 
+	    rhoMatrix[i,j]  <- spear$estimate
+	    if(spear$p.value < 0.05){
+		sigMatrix[i,j]  <- "X"
+	    }
+      }
+}
+
+#generate plots
+pdf(file=    "../figures/spearmanHeatmap.pdf", width = 7,  height = 6)
+par(mar = c(8,4,4,4) + .1) #c(bottom, left, top, right). default: c(5, 4, 4, 2) + 0.1
+heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black", col=redblue(40), density.info="none", trace="none", dendrogram=c("column"), symm=F,symkey=T,symbreaks=T,
+	scale="none", key.title = "", srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), breaks=(-20:20)/20,
+margins = c(8, 8), cexRow=1.5, cexCol=1.5)
+dev.off()
+
+relCitesA<-cor.test(1-d$accuracyRank, as.numeric(d$relCites),     method = "spearman")
+hindexA  <-cor.test(1-d$accuracyRank, as.numeric(d$hindex),       method = "spearman")
+mindexA  <-cor.test(1-d$accuracyRank, as.numeric(d$mindex),       method = "spearman")
+H5A      <-cor.test(1-d$accuracyRank, as.numeric(d$H5),           method = "spearman")
+relAgeA  <-cor.test(1-d$accuracyRank, as.numeric(d$relAge),       method = "spearman")
+speedA   <-cor.test(1-d$accuracyRank, as.numeric(d$speedRank),    method = "spearman")
+citesA   <-cor.test(1-d$accuracyRank, as.numeric(d$cites),        method = "spearman")
+IFA      <-cor.test(1-d$accuracyRank, as.numeric(d$IF),           method = "spearman")
+yearA    <-cor.test(1-d$accuracyRank, as.numeric(d$yearPublished),method = "spearman")
+
+pdf(file=    "../figures/spearmanBarplot.pdf", width = 5,  height = 3)
+op<-par(mfrow=c(1,1),cex=1.0,las=2)
+barplot(t(c(mindexA$estimate, hindexA$estimate, relAgeA$estimate, H5A$estimate, speedA$estimate, citesA$estimate, relCitesA$estimate, yearA$estimate, IFA$estimate)),
+	names=c("M-index", "H-index", "Rel. age", 'JH5', "Speed", "Cites", "Rel. cites", "Year", "JIF"), ylab="Spearman's rho",ylim=c(-0.1,0.1),main="Correlates with accuracy rank")
+lines(c(-100,100),c(0,0))
+dev.off()
+\end{verbatim}
+}
+
+\clearpage
+\newpage
 

### 2016-06-23 bin/prettyPlot.R

@@ -1,135 +1,42 @@
 #!/usr/bin/Rscript
 
+#Rscript ../bin/prettyPlot.R
 #R CMD BATCH ../bin/prettyPlot.R
+
 
 library(MASS)  # in case it is not already loaded 
 library(RColorBrewer)
 library(fields)
+library(gplots)
+library(metap)
+library("CombinePValue")
 library(gplots)
 
 k <- 11
 my.cols <- rev(brewer.pal(k, "RdYlBu"))
 d <-read.table("meanRankSpeedData.tsv", header=T)
 dr<-read.table( "rawRankSpeedData.tsv", header=T)
-reg1  <- lm(d$speedRank~d$accuracyRank)
+
+
+######################################################################
+#TESTING IF SPEED inv. ACCURACY IS TRUE:
+
+reg1  <- lm(d$accuracyRank ~ d$speedRank)
 #reg1c <- coefficients(reg1)
 summary(reg1)
 
 regM <- lm(accuracyRank ~ speedRank + IF + H5 + cites + hindex + mindex + relAge + relCites, data=d)
 summary(regM)
 
-######################################################################
-plotMe <- function(r) {
-       
-op<-par(mfrow=c(1,1),cex=1.1,las=2)
-
-par(mar = c(5,4,4,5) + .1)
-smoothScatter(d$speedRank, d$accuracyRank, nbin=1000, nrpoints=0, colramp=colorRampPalette(my.cols), postPlotHook = fudgeit, pch=19, cex=.85, ylab="mean normalised accuracy rank", xlab="mean normalised speed rank",xlim=c(1.2,-0.2),ylim=c(1.2,-0.2), xaxt = "n", yaxt = "n",main="Accuracy vs. Speed") 
-lines(lowess(d$speedRank, d$accuracyRank, f = .2), col = 2, lwd=5)
-axis(1,at=(0:5)/5)
-axis(2,at=(0:5)/5)
-boxit()
-
-}
-
-######################################################################
-fudgeit <- function(){
-  xm <- get('xm', envir = parent.frame(1))
-  ym <- get('ym', envir = parent.frame(1))
-  z  <- get('dens', envir = parent.frame(1))
-  colramp <- get('colramp', parent.frame(1))
-  my.cols <- rev(brewer.pal(k, "RdYlBu"))
-
[…]

### 2016-06-23 bin/tsv2data.pl

@@ -204,7 +204,7 @@ while(my $in = <IN1>){
 close(IN1);
 close(UT0);
 
-my $numPermutations = 1000;
+my $numPermutations = 10000;
 my (%benchAccuracyPerms, %benchSpeedPerms, %ranksPerms);
 #PERMUTATIONS & RELATIVE VALUES FOR EACH BENCHMARK:
 my (%datesRelRanks, %citesRelRanks);

### 2016-06-25 bin/prettyPlot.R
Renamed flow diagram files.

@@ -109,12 +109,6 @@ heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black", col=redblue
 margins = c(8, 8), cexRow=1.5, cexCol=1.5)
 dev.off()
 
-png(file=    "../figures/spearmanHeatmap.png", width = 420,  height = 360)
-par(mar = c(8,4,4,4) + .1, cex=0.5) #c(bottom, left, top, right). default: c(5, 4, 4, 2) + 0.1
-heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black", col=redblue(75), density.info="none", trace="none", dendrogram=c("row"), symm=F,symkey=T,symbreaks=T, scale="none", cexRow=0.95, cexCol=0.95, key.title = "Spearman\'s rho") 
-dev.off()
-
-
 ############
 #BARPLOT OF SPEARMAN RHO VALUES:
 

### 2016-12-06 manuscript/supplementary.tex
Renamed flow diagram files.
Overdue update. Mostly worked on the manuscript.

--- 2016-06-17 manuscript/supplementary.tex
@@
 \graphicspath{{../figures/}}
+
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S\arabic{figure}}%
+     }
 
@@
 
-\Abstract{.}
+\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
 
@@
 
-\section*{Introduction} % The \section*{} command stops section numbering
+\onecolumn
+\beginsupplement
 
-XXXX
+\section*{Literature mining} % The \section*{} command stops section numbering
+
+\begin{figure}[H]
+\centering
+\includegraphics[height=0.7\textheight]{litMiningFlowDiagram-edited.pdf}
+\caption{In order to improve the identification of benchmark articles
+  that rank both accuracy and speed we developed a tool for ranking
+  PubMed articles based upon word association scores (measured in
+  ‘bits’). In brief, keywords were extracted from titles and abstracts
+  for both training (in this case benchmark articles) and background
+  articles (articles published between 2013 and 2015 with
+  “bioinformatics” in the title or abstract). Log-odds ratios were
+  computed for each keyword (measured in ‘bits’). Candidate articles
+  that matched a hand-selected list of keywords associated with
+  benchmarks were then scored and ranked with a “sum of bits”
+  score. High ranking articles were then inspected, those that met our
+  criteria were added to the training set, those that didn’t were
+  added to the background set of articles.}
+\label{fig:flow}
+\end{figure}
+
+\begin{figure}[H] \centering
+\includegraphics[width=1.05\textwidth]{wordScores.pdf} \caption{The 40
+  highest and lowest scoring words that are associated with
+  bioinformatic benchmark articles from the training benchmark articles,
+  compared to the background articles. The log-odds ratios (measured
+  in bits) are indicated on the y-axis. }
+\label{fig:wordScores} \end{figure}
+
+%% \clearpage
+%% \newpage
+
+%% \section*{R code ...}
+
+%% {\tiny
+%% \begin{verbatim}
+
+%% \end{verbatim}
+%% }
 
@@
 
-\onecolumn
+\section*{Data mining} % The \section*{} command stops section numbering
 
@@
 \centering
-\includegraphics[width=0.55\textwidth]{supplementary-figures-small.pdf}
-\caption{The distributions for the metrics used in this study. These
+\includegraphics[width=0.75\textwidth]{supplementary-figures-small.pdf}
+\caption{The distributions for the measures we expected to be
+  predictive of software quality used in this study. These
 are, reading from left to right, top to bottom: Accuracy -- the mean
@@
 data from GoogleScholar User Profiles;}
-\label{fig:S1}
+\label{fig:metricDistributions}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.45\textwidth]{spearmanBarplot.pdf}
+\includegraphics[width=0.45\textwidth]{spearmanBarplotSpeed.pdf}
+\caption{The correlation between method accuracy (on the left) and
+  method speed (on the right) and measures we expected to be
+  predictive of software quality. E.g. author reputation measures
+  (H-index, M-index), journal reputation (JH5 and JIF), number of
+  users (citations and relative citations) or the recency of methods
+  (year and relative age).  The correlations were estimated using
+  Spearman's $\rho$. The signficant relationships are indicated with a
+  ``*''. }
+\label{fig:S2}
 \end{figure}
@@
 \includegraphics[width=0.5\textwidth]{smoothScatter-speed-vs-accuracy.pdf}
-\caption{...}
-\label{fig:S2}
+\caption{A smoothed color density representation of a scatterplot of
+  normalised speed ranks and normalised accuracy ranks. Dark red
+  regions are indicative of of a high density of points, blue shades
+  indicate the reverse. The expected inverse relationship between
+  speed and accuracy is indicated with a dashed black line, points
+  above this line could be considered comparatively fast and accurate,
+  points below are the reverse. A locally weighted regression (lowess)
+  curve is shown in red.}
+\label{fig:smoothScatterAccSpeed}
 \end{figure}
@@
 \includegraphics[width=0.85\textwidth]{smoothScatters.pdf}
-\caption{...}
-\label{fig:S3}
+\caption{A smoothed color density representation of a scatterplot of
+  journal impact factors (JIF) on the left and number of citations on
+  the right vs normalised accuracy ranks. As above, dark red regions
+  are indicative of of a high density of points, blue shades indicate
+  the reverse. The JIF for journals that are major publishers of
+  bioinformatic software are indicated on the left. }
+\label{fig:smoothScatters}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.45\textwidth]{relAge-speedAcc.pdf}
-\includegraphics[width=0.45\textwidth]{relAge-SpeedVsAccuracy-heatmap.pdf}\\
 \includegraphics[width=0.45\textwidth]{hindex-SpeedVsAccuracy-heatmap.pdf}
-\caption{...}
-\label{fig:S4}
+\includegraphics[width=0.45\textwidth]{cites-SpeedVsAccuracy-heatmap.pdf}
+\caption{Heatmaps of normalised speed ranks and normalised accuracy
+  ranks, both $x$ and $y$ dimensions are discretised into a $10 \times
+  10$ matrix. The shading indicates median H-indices for corresponding
+  authors (darker shade indicates a higher H) in the figure on the
+  left. The shading indicates median citations (on a log scale), for
+  software tools in the figure on the right. }
+\label{fig:heatmaps}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.85\textwidth]{wordScores.pdf}
-\caption{...}
-\label{fig:S5}
+\includegraphics[width=0.95\textwidth]{relAge-SpeedVsAccuracy-heatmap-edited.pdf}
+\caption{{\bf A.} Violin plots for the relative age distribution for software tools in each of the 5 2x2 cells indicated in B.
+The five boxes correspond to the four extreme corners of the speed vs accuracy spectrum (i.e.
+slow and inaccurate, slow and accurate, fast and inaccurate, fast and accurate) and the central box (medium speed and medium accuracy).
+{\bf B.} A heatmap indicating the relative age of software in a range of relative accuracy and speed rankings.
+Blue colours indicate an abundance of older software tools in an accuracy and speed category, while light colours indicate younger software in an accuracy and speed category.
+}
+\label{fig:ageplot}
 \end{figure}

### 2016-12-06 bin/prettyPlot.R

@@ -11,11 +11,15 @@ library(gplots)
 library(metap)
 library("CombinePValue")
 library(gplots)
+library("hash")
+library("vioplot")
 
 k <- 11
 my.cols <- rev(brewer.pal(k, "RdYlBu"))
 d <-read.table("meanRankSpeedData.tsv", header=T)
 dr<-read.table( "rawRankSpeedData.tsv", header=T)
+
+
 
 
 ######################################################################
@@ -27,6 +31,14 @@ summary(reg1)
 
 regM <- lm(accuracyRank ~ speedRank + IF + H5 + cites + hindex + mindex + relAge + relCites, data=d)
 summary(regM)
+
+
+cbind(d$accuracyRank, d$speedRank, d$IF, d$H5, d$cites, d$hindex, d$mindex, d$relAge, d$relCites)
+pca<-prcomp(na.omit(cbind(d$accuracyRank, d$speedRank, d$IF, d$H5, d$cites, d$hindex, d$mindex, d$relAge, d$relCites, 2016-d$yearPublished+1)), center = TRUE,scale=TRUE)
+pca
+summary(pca)
+
+
 
 ##########################
 #Combining P-values and Z-scores:
@@ -105,8 +117,8 @@ for(i in 1:length(dNames)){
 
 pdf(file=    "../figures/spearmanHeatmap.pdf", width = 7,  height = 6)
 par(mar = c(8,4,4,4) + .1) #c(bottom, left, top, right). default: c(5, 4, 4, 2) + 0.1
-heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black", col=redblue(40), density.info="none", trace="none", dendrogram=c("column"), symm=F,symkey=T,symbreaks=T, scale="none", key.title = "", srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), breaks=(-20:20)/20,
-margins = c(8, 8), cexRow=1.5, cexCol=1.5)
+heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black", col=rev(redblue(40)), density.info="none", trace="none", dendrogram=c("column"), symm=F,symkey=T,symbreaks=T, scale="none", key.title = "", srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), breaks=(-20:20)/20,
[…]

### 2016-12-07 manuscript/manuscript-speed-accuracy.tex
Converted the manuscript to latex

\documentclass[fleqn,10pt]{SelfArx} % Document font size and equations flushed left

\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\graphicspath{{../figures/}}
\setlength{\abovecaptionskip}{15pt plus 3pt minus 2pt}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
\definecolor{color2}{RGB}{200,200,200} % Color of the boxes behind the abstract and headings
\definecolor{color3}{RGB}{5,5,5} % Color of the boxes behind the abstract and headings

\JournalInfo{.} % Journal information ``Journal, Vol. XXI, No. 1, 1-5, 2015''
\Archive{ } % Additional notes (e.g. copyright, DOI, review/research article)

\PaperTitle{A meta-analysis of bioinformatics software benchmarks reveals that publication-bias unduly influences software accuracy} % Article title

\Authors{Paul P. Gardner\textsuperscript{1,2,3}*, James M. Paterson\textsuperscript{1,2}, Fatemeh Ashari-Ghomi\textsuperscript{1,2}, Sinan U. Umu\textsuperscript{1,2}, Stephanie McGimpsey\textsuperscript{4,5}, Aleksandra Pawlik\textsuperscript{6}} % Authors
\affiliation{\textsuperscript{1}\textit{School of Biological Sciences, University of Canterbury, Christchurch, New Zealand.}} % Author affiliation
\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{\textsuperscript{3}\textit{Bio-Protection Research Centre, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{\textsuperscript{4}\textit{McConomy School of Dance, Derry, Ireland.}}
\affiliation{\textsuperscript{5}\textit{Research for Good, Limavady, Ireland.}}
\affiliation{\textsuperscript{6}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
\affiliation{*\textbf{Corresponding author}: paul.gardner@canterbury.ac.nz} % Corresponding author

\Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

%\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
\Abstract{
Computational biology has provided widely used and powerful software tools for testing and making inferences about biological data. In the face of increasing volumes of data, heuristic methods that trade software speed for mathematical completeness must be employed. 
We are interested in whether trade-offs between speed and accuracy are reasonable. Also, what factors are indicative of accurate software?

In this work we mine published benchmarks of computational biology software, we collect data on the relative accuracy and speed of different software and then test to see what factors influence accuracy e.g. speed, author reputation, journal impact or recency.

We found that author reputation, journal impact, the number of citations, software speed and age are not reliable predictors of software accuracy. This implies that useful bioinformatics software is not only the domain of famous researchers, and that any researchers are capable of producing good software. In addition, we found that there exists an excessive number of slow and inaccurate software tools across multiple sub-disciplines of bioinformatics. Meanwhile, there are very few tools of middling accuracy and speed. We hypothesise that a strong publication bias is unduly influencing the publication and development of bioinformatic software tools. In other words, at present software that is not highly ranked on speed and not highly ranked on accuracy is difficult to publish due to editorial and reviewer practices. This leaves an unfortunate gap in the literature upon which future software refinements could be constructed.
}

\begin{document}

\flushbottom % Makes all text pages the same height
\maketitle % Print the title and abstract box
%\tableofcontents % Print the contents section

\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

%\onecolumn
%\beginsupplement

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section*{Background}
Computational biology software is widely used and has produced some of the most cited publications in the scientific corpus \cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. This software includes implementations of methods for sequence alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga}, phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, statistical analysis of survival patterns in biomedicine \cite{Kaplan1958-ju,Cox1972-nu}, biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}. However, the popularity of a software tool does not necessarily imply that it is either accurate or computationally efficient, instead usability, ease of installation, operating system and other factors may play a greater role.

Progress in the biological sciences is increasingly limited by the ability to analyse increasing volumes of data, therefore the dependence of biologists on software is also increasing \cite{Marx2013-zi}. There is an increasing use of technological solutions for automating biological data generation (e.g. next-generation sequencing, mass-spectroscopy, cell-tracking and species tracking), therefore the biological sciences have become increasingly dependent upon computational software for processing large quantities of data \cite{Marx2013-zi}. As a consequence, the computational efficiency of data processing and analysis software is of great importance to decrease the energy and time costs of research \cite{Gombiner2011-md}. Furthermore, even small error rates can have a major impact on the number of false inferences as datasets become larger \cite{Storey2003-cv}.

The gold-standard for determining accuracy is for independent researchers to conduct benchmarks, which can serve a useful role in reducing the over-optimistic reporting of software accuracy \cite{Boulesteix2010-te,Jelizarow2010-zf} and the self-assessment trap \cite{Norel2011-cq}. Benchmark studies typically use a number of positive and negative control datasets, predictions can then be partitioned into true or false groups and a variety of metrics can be used to evaluate the performance of different predictions \cite{Egan1975-nd,Hall2012-kg}. Some benchmarks now use live, or frequently updated, results to indicate the latest developments in software performance \cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}. The aim of these benchmarks is to independently identify tools that make acceptable compromises in terms of scoring schemes and the resulting potential for false predictions, true predictions and speed, and are therefore suited for wide adoption by the community.

For common computational biology tasks, a proliferation of software-based solutions often exists \cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut,Wikipedia_contributors2015-vj,Wikipedia_contributors2015-hr}. While this may generally be a good problem to have, and points to a diversity of options from which practical solutions can be selected, many possible options creates a dilemma for users. In the absence of any recent gold-standard benchmarks, how should scientific software be selected? In the following work, we presume that ``biological accuracy'' is the most desirable feature of software. Biological accuracy is the degree to which predictions or measurements reflect the truths of biological systems, this is usually determined by comparing results of software to the results of established reference data. These are commonly expert-derived curated datasets. In some fields biological accuracy is very difficult to ascertain, for example, in phylogenetics it is nearly impossible to know the ancestral relationships between organisms. In situations like this, researchers can use a mix of simulated or high-confidence datasets.

A number of possible predictors of software quality are used by the community of computational biology software users. Some accessible, quantifiable and frequently used proxies for identifying high quality software include: \textbf{1. Recency:} recently published software tools may have built upon the results of past work or be an update to an existing software. Therefore, these could be more accurate and faster. \textbf{2. Wide adoption:} a software tool may be widely used because it is fast and accurate, or because it is well-supported and user-friendly. In fact, “large user base”, “word-of-mouth”, “wide-adoption”, "personal recommendation," and "recommendation from a close colleague," are frequent responses to surveys of “how do scientists select software?” \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal impact:} high profile journals are run by editors and reviewers who carefully select and curate the best manuscripts. Therefore, high impact journals may be more likely to select manuscripts describing good software \cite{Garfield1955-wf}. \textbf{4. Author/Group reputation:} the key to any project is the skills of the people involved, including maintaining a high collective intelligence \cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a consequence, an argument could be made that well respected and high-profile authors will produce better software \cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software is frequently said to trade accuracy for speed. For example, heuristic software such as the popular homology search tool, BLAST, compromise the mathematical guarantee of optimal solutions for more speed \cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively interpret this fact as slower software is likely to be more accurate. But speed may also be influenced by the programming language \cite{Fourment2008-vl}, and/or level of hardware optimisation \cite{Farrar2007-ky,Dematte2010-ph}; In general the implementation is likely to have more of an impact (e.g. brute-force approaches versus rapid and sensitive pre-filtering \cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).

Other factors that influence whether a software tool is selected include: whether the documentation is good, user-friendly, word-of-mouth and ``used in a similar analysis'' \cite{Loman2015-bw}. This sort of information is not as readily quantifiable as the above measures. However, citation metrics may be a useful proxy. The word-of-mouth factor may also explain the reason why some software continues to be used, in spite of poor relative performance \cite{Wadi2016-dj}.

In the following study, we explore factors that may be indicative of software accuracy. This, in our opinion, should be one of the prime reasons for selecting a software tool. We have mined the large and freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks of computational biology software, and manually extracted accuracy and speed rankings for \textbf{{\color{red}243}} software packages. For each software tool, we have collected measures that may be predictive of accuracy, and may be subjectively employed by the researcher community as a proxy for software quality. These include relative speed, relative age, the productivity and impact of the corresponding authors, journal impact and the number of citations.

\section*{Results}
We have collected relative accuracy and speed ranks for \textbf{{\color{red}243}} distinct software tools. This software has been developed for solving a broad cross-section computational biology tasks. These include software for homology search \cite{Freyhult2007-et}, genome sequence analysis (e.g. read mapping or sequence assembly) \cite{Junemann2014-mb,Tran2014-pe,Zhang2011-nd,Abbas2014-gu,Bao2011-lv,Caboche2014-lj,Kleftogiannis2013-wi,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Yang2013-aj,Holtgrewe2011-fd,Rackham2015-ag,Huang2015-wu}, multiple sequence alignment \cite{Thompson2011-rf,Nuin2006-nk,Pais2014-sr,Pervez2014-zp,Liu2010-rp}, cell tracking \cite{Maska2014-ak}, transcriptome analysis \cite{Li2012-wr,Lu2013-fs,Liu2014-kz,Kumar2016-xz}, RNA interaction prediction \cite{Pain2015-gr}, protein interactions \cite{Tikk2010-qd}, protein structure prediction \cite{Kolodny2005-ry,Wallner2005-qi}, epistasis \cite{Shang2011-vy},  metagenomic analysis \cite{Lindgreen2016-tt,Bazinet2012-wf}, repetitive sequence prediction \cite{Saha2008-kd}, proteomics \cite{Lange2008-pt,Yang2009-oc} and phylogenetics \cite{Liu2011-pz,Yang2011-dv,Oscamou2008-md,Bayzid2013-hc,Liu2009-lx}. Each software tool was benchmarked in at least one of \textbf{{\color{red}43}} publications that satisfy the Boulesteix criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria are: 1. the main focus of the article is a benchmark. 2. the authors are reasonably neutral. 3. the test data and evaluation criteria are sensible.

For each of the publications describing these methods, we have (when possible) identified the 2014 journal impact factor (JIF), published by Thomson Reuters \cite{Garfield2006-lv} and the H5-index published by \href{https://scholar.google.co.nz/citations?view_op=top_venues&hl=en}{Google Scholar Metrics}. We have collected the H-indices and M-indices \cite{Hirsch2005-mt} for the corresponding authors for each method, and the number of times the publication(s) associated with a method has been cited using Google Scholar (data collected over a 1 month period in early 2016).

We have computed the Spearman’s correlation coefficient for each pairwise combination of the mean normalised accuracy and speed ranks, the year published, mean relative age (compared to software in the same benchmarks), journal IF and H5 metrics, the total number of citations, the relative number of citations (compared to software in the same benchmarks) and the maximum H and M indices for the corresponding authors. The results are presented in Figure~\ref{fig:allfactors}A. We found significant associations between most of the citation-based metrics (journal H5, JIF, citations, relative citations, H-index and M-index). There is also a strong association between the year of publication, the relative age and many of the citation-based metrics.

We found that author reputation metrics, journal impacts and the age of methods were \textbf{not} significantly correlated with either method accuracy or speed (see Figure~\ref{fig:allfactors}). The strongest association was between accuracy and journal impact factor (Spearman’s rho = 0.1, P-value = 0.16). A linear model of these parameters and accuracy also failed to identify a correlation between these (accuracy: R2 = -0.03, P-value = 0.94; speed: R2 = 0.04, P-value = 0.66).

\begin{equation*}
\begin{split}
accuracy=& c_0+c_1\times speed+c_2\times JIF+c_3\times H5+\\
& c_4\times citations+c_5\times Hindex+\\
& c_6\times Mindex+c_7\times relativeAge+\\
& c_8\times relativeCitations
\end{split}
\end{equation*}

\begin{equation*}
\begin{split}
speed=& c_9+c_{10}\times accuracy+c_{11}\times JIF+c_{12}\times H5+\\
& c_{13}\times citations+c_{14}\times Hindex+\\
& c_{15}\times Mindex+c_{16}\times relativeAge+\\
& c_{17}\times relativeCitations
\end{split}
\end{equation*}
To further validate this result, we compute a correlation between speed and accuracy for each benchmark and used weighted sum Z-tests \cite{Zaykin2011-tj}. This also failed to identify a significant relationship (sum Z=-0.1, P-value=0.5).

In order to gain a deeper understanding of the distribution of available bioinformatic software tools on a speed versus accuracy landscape, we ran a Monte Carlo permutation test. The ranks extracted from each benchmark were randomly permuted, generating 10,000 randomized speed and accuracy ranks. In the cells of a $10\times10$ grid spanning the normalised speed and accuracy ranks we computed a Z-score for the observed number of methods in a cell, compared to the expected distributions generated by the randomized ranks. The results of which are shown in Figure~\ref{fig:speedaccuracy}. We identified 21 bins where there was a significant excess or dearth of methods. For example, there was an excess of “slow and inaccurate” software (Z=1.6, P-value=0.05) and “slow and accurate” software (Z=1.9, P-value=0.03). We find that the amount of software classed as “fast and accurate” and “fast and inaccurate” are at approximately the expected proportions based upon the permutation test. The number of significant results is in excess and not due to multiple testing, as the probability of finding 21 of 100 tests significant by chance is low (P-value = $2\times 10^{-8}$, exact binomial test) \cite{Moran2003-ve}.

There is a large reduction in the number of software tools that are classed as intermediate in terms of both speed and accuracy based upon our permutation test (Figure~\ref{fig:speedaccuracy}). The cells corresponding to the four central underrepresented deciles are highlighted (Z = -1.7, -2.1, -2.1 and -2.4, P-values = 0.04, 0.02, 0.02 and 0.008, respectively, reading from top to bottom, left to right). We also tested the relative age of the software tools in significantly over- or under-represented regions of the speed vs accuracy plot. We found that the “slow and inaccurate” methods were generally published earlier than other methods (W = 31, P=0.007, one-tailed Wilcoxon test) (Figure S8).

\section*{Discussion}
We have gathered data on the relative speed and accuracies of \textbf{{\color{red}243}} bioinformatic methods from \textbf{{\color{red}43}} benchmarks that were published between 2005 and 2016. We show that there is an \textbf{under-representation of software that has both intermediate levels of accuracy and speed}. There may be a number of factors that drive this phenomena. One likely explanation is that \textbf{bioinformatic software tools suffer from a form of publication bias} \cite{Boulesteix2015-am}. Our community of developers, reviewers and editors may be unwilling to publish software that is neither the fastest nor the most accurate (we have anecdotal evidence to support this, N=1). If correct, this is unfortunate. Some problems that lack fast and accurate solutions force researchers to make unnecessary compromises in terms of accuracy and time. If our hypothesis that this underrepresentation is due to publication bias is valid then why is there an enrichment of slow and inaccurate software (P-value=0.05, empirical distributions from permutation tests)? How could these methods be published? We have found that slow and inaccurate software is generally published earlier than alternative methods (P=0.007, one-tailed Wilcoxon test), therefore comparisons were not required to publish these tools and the accuracies may have been over-optimistically reported \cite{Boulesteix2015-nc}.

We found that no commonly used proxy for study quality is correlated with software accuracy. Neither, author reputation, number of citations, journal impact, relative age or speed appear to be associated with accurate software. Linear mixed models of these values also fail to identify predictors of software accuracy, as do methods for combining P-values. A great deal of criticism has been leveled specifically at journal impact factors \cite{Editors2006-ue,Rossner2007-yk,Lariviere2010-kx}. Our finding show that this measure is not reflective of software quality. But neither is the H5 index which is thought to be a more robust measures of journal impact \cite{Harzing2008-wb}. The poor relationship between accuracy and both author reputation and the number of citations is particularly troubling. Both are related to “word of mouth” and “previously used in a similar analysis” which a recent survey of researchers suggested is a major influence on the selection of software tools \cite{Loman2015-bw}. This implies that the recorded high citation rates for bioinformatic software  \cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a reflection of user-friendliness and the Matthew Effect \cite{Lariviere2010-kx,Merton1968-cb}. The Matthew Effect is a biblical reference that can be paraphrased as “the rich get richer and the poor get poorer”. In this context, highly cited software is more likely to be used and cited again, irrespective of relative performance.

The lack of any relationship between software speed and accuracy is surprising. 
The slower software tools are found to be overrepresented at both high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}). Likewise accurate software was found in both high and low speed ranges. A simple gedankenexperiment may be sufficient to prove that slow software is less thoroughly tested than fast software. This is because typical software development is an iterative process, where methods are refined by successive rounds of coding, testing and evaluation \cite{Wilson2006-ih}. We can assume that similar time-spans are spent on most projects, for example, the span of a MSc or PhD degree.  As a consequence slow software undergoes fewer development cycles than fast methods, and is therefore tested less, often resulting in less accurate and slower software. The fact that fast and inaccurate software is relatively rare does support this argument. The lack of any further apparent relationship between speed and accuracy implies there are stronger influences on accuracy rankings than speed alone.

\section*{Conclusions}
Scientific progress is a grinding process made by testing hypotheses using appropriate positive and negative controls, in combination with independent experiments and replication \cite{Ioannidis2005-xh,Moonesinghe2007-kk}. We think that software and analysis tools should be no exception to these principles. The continual increases in the complexity of software tools creates an increased likelihood that software bugs are introduced \cite{Darriba2015-sz}. Scientific software should therefore be thoroughly tested by developers, however, the results of developer and author-derived tests should be treated with caution due to a range of potential conscious and unconscious biases \cite{Norel2011-cq}. Our study shows that no commonly-used impact-based metrics are related to software accuracy, therefore the only reasonable way to select software tools is through software benchmarks.

We have shown that slow and inaccurate software is typically published early in the development of a field (see Figure S8). Subsequent software tools may be high accuracy or high speed. However, as further software tools become available, those that do not rank highly in terms of accuracy or speed are under-represented in the literature. Our conjecture is that this is due to a publication bias in computational biology software literature, with publishers, editors and reviewers implicitly requiring software to be notable in at least one of these regards in order to be published. This hole in the literature leaves an unfortunate gap upon which further software refinements cannot be made.

The software that could be used in building research workflows can make use of software tools that are of medium accuracy and speed. These may prove to be useful rapid pre-filters for more accurate, yet computationally-demanding approaches. These could also be used when datasets are too large to be processed by more accurate software tools, i.e. when accurate tools are slow or memory hungry \cite{Lowe1997-hq,Weinberg2006-tm}. These tools may also serve as an indication that some approaches may not be worth pursuing further.

We propose that the full spectrum of software tool accuracies and speeds serve a useful purpose to the research community. Like negative results, if honestly reported, illustrate to the research community that certain approaches are not practical research avenues \cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}. The current practises of publishers, editors, reviewers and authors of software tools appear to have  deprived our community of tools for building effective and productive workflows.

A potential avenue for further exploration is to compare the starting point for software development projects as certain approaches may produce more rapid gains than others. For example, starting with biologically plausible and fast methods will theoretically allow more rapid gains in accuracy through iterative method refinement than starting with slow and mathematically complete approaches. At the very least these may be used as rapid data filters for reducing the size and complexity of problems, prior to employing more rigorous methods.

We have shown that accurate software is not necessarily the most recently released method or the product of high profile lab groups or selected by high impact journals. Software that is widely used or is either slow or fast is also not necessarily the most accurate. Therefore, accurate software may be the product of features we have not been able to capture. Possibly, hard work, good ideas and sound method testing, as well as technical ability, experience and education levels in software development \cite{Wilson2006-ih}.

Finally, we think that the field of computational biology could benefit from embracing an increased number of independent software comparison studies \cite{Boulesteix2013-vb}, in parallel with the popular challenge-based benchmarks such as CASP, Assemblathon, Alignathon, DREAM and RGASP \cite{Boutros2014-zm}. We are hopeful that this, along with the relaxation of the publication bias we have described, will reduce the over-optimistic and misleading reporting of method accuracy \cite{Boulesteix2010-te,Jelizarow2010-zf,Norel2011-cq}.

\section*{Methods}
In order to evaluate predictors of computational biology software accuracy, we mined the published literature, extracted data from articles, connected these with bibliometric databases, and tested for correlates with accuracy. We outline these steps in further detail below.

\textbf{Criteria for inclusion:} We are interested in using computational biology benchmarks that satisfy Anne-Laure Boulesteix’s (ALB) three criteria for a “neutral comparison study” \cite{Boulesteix2013-vb}. Firstly, the main focus of the article is the comparison and \textbf{not} the introduction of a new method. Secondly, the authors should be reasonably neutral, which means that the authors should not generally have been involved in the development of the methods included in the benchmark. Thirdly, the test data and evaluation criteria should be sensible. This means that the test data should be independent of data that methods have been trained upon, and that the evaluation measures appropriately quantify correct and incorrect predictions.

Literature mining: We identified an initial list of 10 benchmark articles that satisfy the ALB-criteria. These were identified based upon previous knowledge of published articles and were supplemented with several literature searches (e.g. ‘“benchmark” AND “cputime”’ was used to query both GoogleScholar and Pubmed \cite{Sayers2010-vm,McEntyre2001-fl}). We used these articles to seed a machine-learning approach for identifying further candidate articles and to identify new search terms to include.

For our machine-learning-based literature screening, we computed a score ($s(a)$) for each article that tells us the likelihood that it is a benchmark. In brief, our approaches uses 3 stages:
\begin{enumerate}
\item Remove high frequency words from the title and abstract of candidate articles (e.g. ‘the’, ‘and’, ‘of’, ‘to’, ‘a’, …) 
\item Compute a log-odds score for the remaining words 
\item Use a sum of log-odds scores to give a total score for candidate articles
\end{enumerate}
For stage 1, we identified a list of high frequency (e.g. f(word) $>$ 1/10,000) words by pooling the content of two control texts \cite{Carroll1865-hk,Tolkien1937-ke}.

For stage 2, in order to compute a log-odds score for bioinformatic words, we computed the frequency of words that were not removed by our high frequency filter in two different groups of articles: bioinformatics-background and bioinformatics-benchmark articles. The text from bioinformatics-background articles were drawn from the bioinformatics literature, but these were not necessarily associated with benchmark studies. For background text we used Pubmed (\cite{Sayers2010-vm,McEntyre2001-fl} to select 8,908 articles that contained the word “bioinformatics” in the title or abstract and were published between 2013 and 2015. We computed frequencies for each word by combining text from titles and abstracts for the background and training articles. A log-odds score is computed for each word using the following formula: $lo(w)=\log_2\frac{•f_{tr}(word)+\delta}{f_{bg}(word)+\delta}$, where  is a prior probability ($\delta = 10^{-5}$, by default), $f_{bg}(word)$ and $f_{tr}(word)$ are the frequencies of a $word$ in the background and training datasets respectively. Word frequencies are computed by counting the number of times a word appears in the pool of titles and abstracts, the counts are normalised by the total number of words in each set.

Thirdly, we also collected a group of candidate benchmark articles by mining Pubmed for articles that are likely to be benchmarks of bioinformatic software, these may match the terms: “((bioinformatics) AND (algorithms OR programs OR software)) AND (accuracy OR assessment OR benchmark OR comparison OR performance) AND (speed OR time)”. Further terms used in this search were progressively added as relevant enriched terms were identified in later iterations. The final query is given in \textbf{supplementary materials}.

A score is computed for each candidate article by summing the log-odds scores for the words in title and abstract, i.e. $s(a)=\sum_i^Nlo(w_i)$. The high scoring candidate articles are then manually evaluated against the ALB-criteria. Accuracy and speed ranks are extracted from the articles that meet the criteria, and these are also added to the set of training articles. The evaluated candidate articles that do not meet the ALB-criteria are incorporated into the set of background articles. This process is iterated a number of times and has resulted in the identification of \textbf{{\color{red}43}} benchmark articles, that contain \textbf{{\color{red}102}} different benchmarks, together these rank \textbf{{\color{red}243}} distinct software packages.

There is a potential for bias to have been introduced into this dataset. Some possible forms of bias include converging on a niche group of benchmark studies due to the literature mining technique that we have used. A further possibility is that benchmark studies themselves are biased, either including very high performing or very low performing software tools. To address each of these concerns we have attempted to be as comprehensive as possible in terms of benchmark inclusion, as well as include comprehensive benchmarks. By which we mean studies that include all available software tools that address a biological problem.

\textbf{Data extraction:} for each article that met the ALB-criteria and contained data on both the accuracy and speed from their tests we extracted ranks for each method. Many articles contained multiple benchmarks, in these cases we selected a range of these, the provenance of which is stored with the accuracy metric and raw speed and accuracy rank data for each method. In line with rank-based statistics, the cases where methods were tied are resolved by using a midpoint rank (e.g. if method 3 and 4 are tied, the rank 3.5 is used) \cite{Mann1947-re}. Each rank extraction was independently verified by at least one other co-author to ensure both the provenance of the data could be established and that the ranks were correct. The ranks for each benchmark were then normalised to lie between 0 and 1 using the formula $\frac{r-1}{n-1}$ where ‘r’ is a method’s rank and ‘n’ is the number of methods in the benchmark. For methods that were benchmarked multiple times with multiple metrics (e.g. BWA is evaluated in 6 different articles \cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd}) a mean normalised rank is used to summarise the performance.

For each method we identified the corresponding publications in GoogleScholar, the total number of citations was recorded, the corresponding authors were also identified and if these had public GoogleScholar profiles we extracted their H-index and calculated a M-index ($\frac{H-index}{y}$) where ‘$y$’ is the number of years since their first publication. For the journals that each method is published in we extracted the “journal impact factor” (JIF) and the H5-index from Thompson-Reuters and GoogleScholar Metrics databases respectively. The year of publication was also recorded for each method. A “relative age” and “relative citations” was also computed for each method. For each benchmark, software was ranked by year of first publication (or number of citations), ranks were assigned and then normalised as described above. Methods ranked in multiple evaluations were then assigned a mean value for “relative age” and “relative citations”.

\textbf{Statistical analysis:} For each method we have up to 10 statistics (1. corresponding author’s H-index, 2. corresponding author’s M-index, 3. journal H5 index, 4. journal impact factor, 5. normalised accuracy rank, 6. normalised speed rank, 7. number of citations, 8. relative age, 9. relative number of citations, 10. year first published). These have been evaluated in a pairwise fashion to produce Figure~\ref{fig:allfactors} A\&B, the R code for these is given in the supplement.

For each benchmark of three or more methods, we extracted the published accuracy and speed ranks. In order to identify if there is an enrichment of certain accuracy and speed pairings we constructed a permutation test. The individual accuracy and speed ranks were reassigned to methods in a random fashion and each new accuracy and speed rank pairing was recorded. For each benchmark this procedure was repeated 10,000 times. These permuted rankings were normalised and compared to the real rankings to produce the ‘x’ points in Figure~\ref{fig:allfactors}B and the heatmap and histograms in Figure~\ref{fig:speedaccuracy}. The heatmap in Figure~\ref{fig:speedaccuracy} is based upon Z-scores ($Z=\frac{x-\bar{x}}{\sigma}$). For each cell in a $10\times 10$ grid a Z-score is computed to illustrate the abundance or lack of methods in a cell relative to the permuted data.

\section*{Data availability}
The raw datasets are available here:\\
\fussy
\url{https://docs.google.com/spreadsheets/d/14xIY2PHNvxmV9MQLpbzSfFkuy1RlzDHbBOCZLJKcGu8/edit?usp=sharing}\\
\sloppy
Additional documentation, code, figures and raw data is available here:\\
\fussy
\url{https://github.com/UCanCompBio/speed-vs-accuracy-meta-analysis}
\sloppy

\section*{Acknowledgements}
The authors acknowledge the valued contribution of invaluable discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia Palmer and Jason Tylianakis. Michael A. Black,  Murray Cox, Raquel Norel, Alexandros Stamatakis, Jens Stoye,Tandy Warnow, provided valuable feedback on drafts of the manuscript.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\bibliographystyle{unsrt}
\bibliography{references.bib}

\begin{figure*}
\includegraphics[scale=0.7]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between proposed predictors of software quality. Spearman’s rho is used to infer correlations between metrics such as the H and M indices of corresponding authors, number of citations, journal impact factors and H5 indices, the year and relative age of software and the mean relative rankings of software speed and accuracy. Red colours indicate a high positive correlation, blue colours indicate a high negative correlation. Correlations with a P-value less than 0.05 are indicated with a ‘X’. The dashed rectangular area is illustrated in more detail in \textbf{B}, the bold square is shown in more detail in \textbf{Figure 2}.The dendrogram was computed using the default ‘heatmap.2 function in R, which computes a Euclidean distance matrix and a complete-linkage hierarchical clustering. \textbf{B.} A barplot illustrating the correlation as measured by Spearman’s rho between and normalised accuracy ranks and potential factors that may be predictive of accuracy. In order to give an appreciation of the difference between the observed effect-sizes and significant effect sizes, we generated 10,000 permuted accuracy ranks for each benchmark and recorded Spearman’s rho for the significant correlations (P<0.05). These values are marked with “x”s in the barplot.}
\label{fig:allfactors}
\end{figure*}

\begin{figure*}
\includegraphics[scale=0.55]{figure2.pdf}
\caption{A heatmap indicating the relative paucity or abundance of software in the range of possible accuracy and speed rankings. Red colours indicate an abundance of software tools in an accuracy and speed category, while blue colours indicate scarcity of software in an accuracy and speed category. The abundance is quantified using a Z-score computation for each bin, this is derived from 10,000 random permutations of the speed and accuracy ranks from each benchmark. Mean normalised ranks of accuracy and speed have been binned into 100 classes (a $10\times10$ grid) that range from comparatively slow/inaccurate to comparatively fast/accurate. Z-scores with a P-value less than 0.05 are indicated with a ‘X’.}
\label{fig:speedaccuracy}
\end{figure*}

\end{document}

### 2016-12-08 manuscript/manuscript-speed-accuracy.tex
Added licence, tidied README...

@@
 This, in our opinion, should be one of the prime reasons for selecting a software tool.
-We have mined the large and freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks of computational biology software, and manually extracted accuracy and speed rankings for \textbf{{\color{red}243}} software packages.
+We have mined the large and freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks of computational biology software, and manually extracted accuracy and speed rankings for \textbf{{\color{black}243}} software packages.
 For each software tool, we have collected measures that may be predictive of accuracy, and may be subjectively employed by the researcher community as a proxy for software quality.
@@
 \section*{Results}
-We have collected relative accuracy and speed ranks for \textbf{{\color{red}243}} distinct software tools.
+We have collected relative accuracy and speed ranks for \textbf{{\color{black}243}} distinct software tools.
 This software has been developed for solving a broad cross-section computational biology tasks.
@@
 read mapping or sequence assembly) \cite{Junemann2014-mb,Tran2014-pe,Zhang2011-nd,Abbas2014-gu,Bao2011-lv,Caboche2014-lj,Kleftogiannis2013-wi,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Yang2013-aj,Holtgrewe2011-fd,Rackham2015-ag,Huang2015-wu}, multiple sequence alignment \cite{Thompson2011-rf,Nuin2006-nk,Pais2014-sr,Pervez2014-zp,Liu2010-rp}, cell tracking \cite{Maska2014-ak}, transcriptome analysis \cite{Li2012-wr,Lu2013-fs,Liu2014-kz,Kumar2016-xz}, RNA interaction prediction \cite{Pain2015-gr}, protein interactions \cite{Tikk2010-qd}, protein structure prediction \cite{Kolodny2005-ry,Wallner2005-qi}, epistasis \cite{Shang2011-vy},  metagenomic analysis \cite{Lindgreen2016-tt,Bazinet2012-wf}, repetitive sequence prediction \cite{Saha2008-kd}, proteomics \cite{Lange2008-pt,Yang2009-oc} and phylogenetics \cite{Liu2011-pz,Yang2011-dv,Oscamou2008-md,Bayzid2013-hc,Liu2009-lx}.
-Each software tool was benchmarked in at least one of \textbf{{\color{red}43}} publications that satisfy the Boulesteix criteria \cite{Boulesteix2013-vb}.
+Each software tool was benchmarked in at least one of \textbf{{\color{black}43}} publications that satisfy the Boulesteix criteria \cite{Boulesteix2013-vb}.
 In brief, the Boulesteix criteria are: 1.
@@
 
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{figure1.pdf}
+\caption{\textbf{A.} A heatmap indicating the relationships between proposed predictors of software quality.
+Spearman’s rho is used to infer correlations between metrics such as the H and M indices of corresponding authors, number of citations, journal impact factors and H5 indices, the year and relative age of software and the mean relative rankings of software speed and accuracy.
+Red colours indicate a high positive correlation, blue colours indicate a high negative correlation.
+Correlations with a P-value less than 0.05 are indicated with a ‘X’.
+The dashed rectangular area is illustrated in more detail in \textbf{B}, the bold square is shown in more detail in \textbf{Figure 2}.The dendrogram was computed using the default ‘heatmap.2 function in R, which computes a Euclidean distance matrix and a complete-linkage hierarchical clustering.
+\textbf{B.} A barplot illustrating the correlation as measured by Spearman’s rho between and normalised accuracy ranks and potential factors that may be predictive of accuracy.
+In order to give an appreciation of the difference between the observed effect-sizes and significant effect sizes, we generated 10,000 permuted accuracy ranks for each benchmark and recorded Spearman’s rho for the significant correlations (P<0.05).
+These values are marked with “x”s in the barplot.}
+\label{fig:allfactors}
+\end{figure*}
+
 We found that author reputation metrics, journal impacts and the age of methods were \textbf{not} significantly correlated with either method accuracy or speed (see Figure~\ref{fig:allfactors}).
@@
 A linear model of these parameters and accuracy also failed to identify a correlation between these (accuracy: R2 = -0.03, P-value = 0.94; speed: R2 = 0.04, P-value = 0.66).
-
-\begin{equation*}
-\begin{split}
-accuracy=& c_0+c_1\times speed+c_2\times JIF+c_3\times H5+\\
-& c_4\times citations+c_5\times Hindex+\\
-& c_6\times Mindex+c_7\times relativeAge+\\
-& c_8\times relativeCitations
-\end{split}
-\end{equation*}
-
-\begin{equation*}
-\begin{split}
-speed=& c_9+c_{10}\times accuracy+c_{11}\times JIF+c_{12}\times H5+\\
-& c_{13}\times citations+c_{14}\times Hindex+\\
-& c_{15}\times Mindex+c_{16}\times relativeAge+\\
-& c_{17}\times relativeCitations
-\end{split}
-\end{equation*}
 To further validate this result, we compute a correlation between speed and accuracy for each benchmark and used weighted sum Z-tests \cite{Zaykin2011-tj}.
@@
 
+\begin{figure*}
+%\includegraphics[scale=0.55]{figure2.pdf}
+\includegraphics[width=\textwidth]{figure2.pdf}
+\caption{A heatmap indicating the relative paucity or abundance of software in the range of possible accuracy and speed rankings.
+Red colours indicate an abundance of software tools in an accuracy and speed category, while blue colours indicate scarcity of software in an accuracy and speed category.
+The abundance is quantified using a Z-score computation for each bin, this is derived from 10,000 random permutations of the speed and accuracy ranks from each benchmark.
+Mean normalised ranks of accuracy and speed have been binned into 100 classes (a $10\times10$ grid) that range from comparatively slow/inaccurate to comparatively fast/accurate.
+Z-scores with a P-value less than 0.05 are indicated with a ‘X’.}
+\label{fig:speedaccuracy}
+\end{figure*}
+
 There is a large reduction in the number of software tools that are classed as intermediate in terms of both speed and accuracy based upon our permutation test (Figure~\ref{fig:speedaccuracy}).
@@
 \section*{Discussion}
-We have gathered data on the relative speed and accuracies of \textbf{{\color{red}243}} bioinformatic methods from \textbf{{\color{red}43}} benchmarks that were published between 2005 and 2016.
-We show that there is an \textbf{under-representation of software that has both intermediate levels of accuracy and speed}.
+
+We have gathered data on the relative speed and accuracies of \textbf{{\color{black}243}} bioinformatic methods from \textbf{{\color{black}43}} benchmarks that were published between 2005 and 2016.
+We show that there is an under-representation of software that has both intermediate levels of accuracy and speed.
 There may be a number of factors that drive this phenomena.
-One likely explanation is that \textbf{bioinformatic software tools suffer from a form of publication bias} \cite{Boulesteix2015-am}.
+One likely explanation is that bioinformatic software tools suffer from a form of publication bias \cite{Boulesteix2015-am}.
 Our community of developers, reviewers and editors may be unwilling to publish software that is neither the fastest nor the most accurate (we have anecdotal evidence to support this, N=1).
@@
 Like negative results, if honestly reported, illustrate to the research community that certain approaches are not practical research avenues \cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
-The current practises of publishers, editors, reviewers and authors of software tools appear to have  deprived our community of tools for building effective and productive workflows.
+The classes of tools that we find in the under-represented accuracy and speed region include some of the most used tools in bioinformatics, such as the homology search tools NCBI-BLAST \cite{Altschul1990-ht,Altschul1997-ga} and HMMER \cite{Eddy:2009} and genome assembly and mapping tools such as AbySS \cite{Simpson:2009}, MAQ \cite{Li:2008} and Newbler \cite{Margulies:2005}.
+The current practises of publishers, editors, reviewers and authors of software tools therefore deprive our community of tools for building effective and productive workflows.
 
@@
 The evaluated candidate articles that do not meet the ALB-criteria are incorporated into the set of background articles.
-This process is iterated a number of times and has resulted in the identification of \textbf{{\color{red}43}} benchmark articles, that contain \textbf{{\color{red}102}} different benchmarks, together these rank \textbf{{\color{red}243}} distinct software packages.
+This process is iterated a number of times and has resulted in the identification of \textbf{{\color{black}43}} benchmark articles, that contain \textbf{{\color{black}102}} different benchmarks, together these rank \textbf{{\color{black}243}} distinct software packages.
 
@@
 
-\textbf{Data extraction:} for each article that met the ALB-criteria and contained data on both the accuracy and speed from their tests we extracted ranks for each method.
+\textbf{Data extraction and processing:} for each article that met the ALB-criteria and contained data on both the accuracy and speed from their tests we extracted ranks for each method.
 Many articles contained multiple benchmarks, in these cases we selected a range of these, the provenance of which is stored with the accuracy metric and raw speed and accuracy rank data for each method.
@@
 These have been evaluated in a pairwise fashion to produce Figure~\ref{fig:allfactors} A\&B, the R code for these is given in the supplement.
+
+The linear models that we used to test for relationships between speed, accuracy and the above measures are:
+
+\begin{equation*}
+\begin{split}
+accuracy=& c_0+c_1\times speed+c_2\times JIF+c_3\times H5+\\
+& c_4\times citations+c_5\times Hindex+\\
+& c_6\times Mindex+c_7\times relativeAge+\\
+& c_8\times relativeCitations
+\end{split}
+\end{equation*}
+
+\begin{equation*}
+\begin{split}
+speed=& c_0+c_{1}\times accuracy+c_{2}\times JIF+c_{3}\times H5+\\
+& c_{4}\times citations+c_{5}\times Hindex+\\
+& c_{6}\times Mindex+c_{7}\times relativeAge+\\
+& c_{8}\times relativeCitations
+\end{split}
+\end{equation*}
 
@@
 
-\begin{figure*}
-\includegraphics[scale=0.7]{figure1.pdf}
-\caption{\textbf{A.} A heatmap indicating the relationships between proposed predictors of software quality.
-Spearman’s rho is used to infer correlations between metrics such as the H and M indices of corresponding authors, number of citations, journal impact factors and H5 indices, the year and relative age of software and the mean relative rankings of software speed and accuracy.
-Red colours indicate a high positive correlation, blue colours indicate a high negative correlation.
-Correlations with a P-value less than 0.05 are indicated with a ‘X’.
-The dashed rectangular area is illustrated in more detail in \textbf{B}, the bold square is shown in more detail in \textbf{Figure 2}.The dendrogram was computed using the default ‘heatmap.2 function in R, which computes a Euclidean distance matrix and a complete-linkage hierarchical clustering.
-\textbf{B.} A barplot illustrating the correlation as measured by Spearman’s rho between and normalised accuracy ranks and potential factors that may be predictive of accuracy.
-In order to give an appreciation of the difference between the observed effect-sizes and significant effect sizes, we generated 10,000 permuted accuracy ranks for each benchmark and recorded Spearman’s rho for the significant correlations (P<0.05).
-These values are marked with “x”s in the barplot.}
-\label{fig:allfactors}
-\end{figure*}
-
-\begin{figure*}
-\includegraphics[scale=0.55]{figure2.pdf}
-\caption{A heatmap indicating the relative paucity or abundance of software in the range of possible accuracy and speed rankings.
-Red colours indicate an abundance of software tools in an accuracy and speed category, while blue colours indicate scarcity of software in an accuracy and speed category.
-The abundance is quantified using a Z-score computation for each bin, this is derived from 10,000 random permutations of the speed and accuracy ranks from each benchmark.
-Mean normalised ranks of accuracy and speed have been binned into 100 classes (a $10\times10$ grid) that range from comparatively slow/inaccurate to comparatively fast/accurate.
-Z-scores with a P-value less than 0.05 are indicated with a ‘X’.}
-\label{fig:speedaccuracy}
-\end{figure*}
-
 \end{document}

### 2016-12-14 manuscript/manuscript-speed-accuracy.tex
Removed humiliating file conflict info.

@@
 Like negative results, if honestly reported, illustrate to the research community that certain approaches are not practical research avenues \cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
-The classes of tools that we find in the under-represented accuracy and speed region include some of the most used tools in bioinformatics, such as the homology search tools NCBI-BLAST \cite{Altschul1990-ht,Altschul1997-ga} and HMMER \cite{Eddy:2009} and genome assembly and mapping tools such as AbySS \cite{Simpson:2009}, MAQ \cite{Li:2008} and Newbler \cite{Margulies:2005}.
+The classes of tools that we find in the under-represented accuracy and speed region include some of the most used tools in bioinformatics, such as the homology search tools NCBI-BLAST \cite{Altschul1990-ht,Altschul1997-ga} and HMMER2 \cite{Eddy:2009} and genome assembly and mapping tools such as AbySS \cite{Simpson:2009}, MAQ \cite{Li:2008} and Newbler \cite{Margulies:2005}.
 The current practises of publishers, editors, reviewers and authors of software tools therefore deprive our community of tools for building effective and productive workflows.
@@
 Each rank extraction was independently verified by at least one other co-author to ensure both the provenance of the data could be established and that the ranks were correct.
-The ranks for each benchmark were then normalised to lie between 0 and 1 using the formula $\frac{r-1}{n-1}$ where ‘r’ is a method’s rank and ‘n’ is the number of methods in the benchmark.
+The ranks for each benchmark were then normalised to lie between 0 and 1 using the formula $1-\frac{r-1}{n-1}$ where ‘r’ is a method’s rank and ‘n’ is the number of methods in the benchmark.
 For methods that were benchmarked multiple times with multiple metrics (e.g.
@@
 \section*{Data availability}
-The raw datasets are available here:\\
+Raw datasets, software and documents are available under a CC-BY license:\\
 \fussy
 \url{https://docs.google.com/spreadsheets/d/14xIY2PHNvxmV9MQLpbzSfFkuy1RlzDHbBOCZLJKcGu8/edit?usp=sharing}\\
+and here:\\
+\fussy
+\url{https://dx.doi.org/10.6084/m9.figshare.4299320.v1}\\
 \sloppy
@@
 
+PPG, FAG and NEW are supported by a Rutherford Discovery Fellowship,
+administered by the Royal Society of New Zealand. NEW is supported by
+a PhD scholarship from the University of Canterbury.
+
 %----------------------------------------------------------------------------------------

### 2020-07-03 manuscript/manuscript-speed-accuracy.tex
Updated README. Re-running analysis.

@@
 We found that author reputation, journal impact, the number of citations, software speed and age are not reliable predictors of software accuracy.
-This implies that useful bioinformatics software is not only the domain of famous researchers, and that any researchers are capable of producing good software.
-In addition, we found that there exists an excessive number of slow and inaccurate software tools across multiple sub-disciplines of bioinformatics.
+Unfortunately, citations, author and journal reputation are frequently cited reasons for selecting software tools.
+Our results imply that useful bioinformatics software is not only the domain of famous senior researchers.
+In addition, we found that there is an excess of slow and inaccurate software tools across multiple sub-disciplines of bioinformatics.
 Meanwhile, there are very few tools of middling accuracy and speed.
-We hypothesise that a strong publication bias is unduly influencing the publication and development of bioinformatic software tools.
+We hypothesise that a form of publication-bias unduly influences the publication and development of bioinformatic software tools.
 In other words, at present software that is not highly ranked on speed and not highly ranked on accuracy is difficult to publish due to editorial and reviewer practices.
-This leaves an unfortunate gap in the literature upon which future software refinements could be constructed.
+This leaves an unfortunate gap in the literature upon which future software refinements cannot be constructed.
 }
@@
 
+%%REFERENCES::
+%reference Stefan Buchka$^{1}$, Rory Wilson$^2$, Alexander Hapfelmeier$^3$,\\  Paul Gardner$^4$,  Anne-Laure Boulesteix PAPER!!!
+
+%% A comprehensive analysis of the usability and archival stability of omics computational tools and resources
+%% https://www.biorxiv.org/content/10.1101/452532v1
+
+%% Tools and best practices for data processing in allelic expression analysis
+%% https://genomebiology.biomedcentral.com/articles/10.1186/s13059-015-0762-6
+
+%% Towards reproducible, transparent, and systematic benchmarking of omics computational tools
+%% https://osf.io/p8yd9
+
+%% Essential guidelines for computational method benchmarking
+%% https://arxiv.org/abs/1812.00661
+
 \section*{Background}
-Computational biology software is widely used and has produced some of the most cited publications in the scientific corpus \cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}.
-This software includes implementations of methods for sequence alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga}, phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, statistical analysis of survival patterns in biomedicine \cite{Kaplan1958-ju,Cox1972-nu}, biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}.
-However, the popularity of a software tool does not necessarily imply that it is either accurate or computationally efficient, instead usability, ease of installation, operating system and other factors may play a greater role.
+Computational biology software is widely used and has produced some of
+the most cited publications in the scientific corpus
+\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. This
+software includes implementations of methods for sequence alignment
+and homology inference
+\cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
+phylogenetic analysis
+\cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei},
+statistical analysis of survival patterns in biomedicine
+\cite{Kaplan1958-ju,Cox1972-nu}, biomolecular structure analysis
+\cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj},
+visualization and data collection
+\cite{Kraulis1991-lt,Berman2000-to}. However, the popularity of a
+software tool does not necessarily imply that it is either accurate or
+computationally efficient, instead usability, ease of installation,
+operating system and other factors may play a greater role. There have
+been several notable incidences where inaccurate software has caused
+considerable harm \cite{leveson1993investigation}.
+%Fergusson covid predictions?
 
-Progress in the biological sciences is increasingly limited by the ability to analyse increasing volumes of data, therefore the dependence of biologists on software is also increasing \cite{Marx2013-zi}.
-There is an increasing use of technological solutions for automating biological data generation (e.g.
-next-generation sequencing, mass-spectroscopy, cell-tracking and species tracking), therefore the biological sciences have become increasingly dependent upon computational software for processing large quantities of data \cite{Marx2013-zi}.
-As a consequence, the computational efficiency of data processing and analysis software is of great importance to decrease the energy and time costs of research \cite{Gombiner2011-md}.
-Furthermore, even small error rates can have a major impact on the number of false inferences as datasets become larger \cite{Storey2003-cv}.
+Progress in the biological sciences is increasingly limited by the
+ability to analyse increasing volumes of data, therefore the
+dependence of biologists on software is also increasing
+\cite{Marx2013-zi}. There is an increasing use of technological
+solutions for automating biological data generation
+(e.g. next-generation sequencing, mass-spectroscopy, cell-tracking and
+species tracking), therefore the biological sciences have become
+increasingly dependent upon computational software for processing
+large quantities of data \cite{Marx2013-zi}. As a consequence, the
+computational efficiency of data processing and analysis software is
+of great importance to decrease the energy and time costs of research
+\cite{Gombiner2011-md}. Furthermore, even small error rates can have a
+major impact on the number of false inferences as datasets become
+larger \cite{Storey2003-cv}.
 
-The gold-standard for determining accuracy is for independent researchers to conduct benchmarks, which can serve a useful role in reducing the over-optimistic reporting of software accuracy \cite{Boulesteix2010-te,Jelizarow2010-zf} and the self-assessment trap \cite{Norel2011-cq}.
-Benchmark studies typically use a number of positive and negative control datasets, predictions can then be partitioned into true or false groups and a variety of metrics can be used to evaluate the performance of different predictions \cite{Egan1975-nd,Hall2012-kg}.
-Some benchmarks now use live, or frequently updated, results to indicate the latest developments in software performance \cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
-The aim of these benchmarks is to independently identify tools that make acceptable compromises in terms of scoring schemes and the resulting potential for false predictions, true predictions and speed, and are therefore suited for wide adoption by the community.
+The gold-standard for determining accuracy is for independent
+researchers to conduct benchmarks, which can serve a useful role in
+reducing the over-optimistic reporting of software accuracy
+\cite{Boulesteix2010-te,Jelizarow2010-zf} and the self-assessment trap
+\cite{Norel2011-cq}. Benchmark studies typically use a number of
+positive and negative control datasets, predictions can then be
+partitioned into true or false groups and a variety of metrics can be
+used to evaluate the performance of different predictions
+\cite{Egan1975-nd,Hall2012-kg}. Some benchmarks now use live, or
+frequently updated, results to indicate the latest developments in
+software performance
+\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}. The aim of
+these benchmarks is to independently identify tools that make
+acceptable compromises in terms of scoring schemes and the resulting
+potential for false predictions, true predictions and speed, and are
+therefore suited for wide adoption by the community.
 
-For common computational biology tasks, a proliferation of software-based solutions often exists \cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut,Wikipedia_contributors2015-vj,Wikipedia_contributors2015-hr}.
-While this may generally be a good problem to have, and points to a diversity of options from which practical solutions can be selected, many possible options creates a dilemma for users.
-In the absence of any recent gold-standard benchmarks, how should scientific software be selected?
-In the following work, we presume that ``biological accuracy'' is the most desirable feature of software.
-Biological accuracy is the degree to which predictions or measurements reflect the truths of biological systems, this is usually determined by comparing results of software to the results of established reference data.
-These are commonly expert-derived curated datasets.
-In some fields biological accuracy is very difficult to ascertain, for example, in phylogenetics it is nearly impossible to know the ancestral relationships between organisms.
-In situations like this, researchers can use a mix of simulated or high-confidence datasets.
+For common computational biology tasks, a proliferation of
+software-based solutions often exists
+\cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut,Wikipedia_contributors2015-vj,Wikipedia_contributors2015-hr}. While
+this may generally be a good problem to have, and points to a
+diversity of options from which practical solutions can be selected,
+many possible options creates a dilemma for users. In the absence of
+any recent gold-standard benchmarks, how should scientific software be
+selected? In the following work, we presume that ``biological
+accuracy'' is the most desirable feature of software. Biological
+accuracy is the degree to which predictions or measurements reflect
+the truths of biological systems, this is usually determined by
+comparing results of software to the results of established reference
+data. These are commonly expert-derived curated datasets. In some
+fields biological accuracy is very difficult to ascertain, for
+example, in phylogenetics it is nearly impossible to know the
+ancestral relationships between organisms. In situations like this,
+researchers can use a mix of simulated or high-confidence datasets.
 
-A number of possible predictors of software quality are used by the community of computational biology software users.
-Some accessible, quantifiable and frequently used proxies for identifying high quality software include: \textbf{1.
-Recency:} recently published software tools may have built upon the results of past work or be an update to an existing software.
-Therefore, these could be more accurate and faster.
-\textbf{2.
-Wide adoption:} a software tool may be widely used because it is fast and accurate, or because it is well-supported and user-friendly.
-In fact, “large user base”, “word-of-mouth”, “wide-adoption”, "personal recommendation," and "recommendation from a close colleague," are frequent responses to surveys of “how do scientists select software?” \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}.
-\textbf{3.
-Journal impact:} high profile journals are run by editors and reviewers who carefully select and curate the best manuscripts.
-Therefore, high impact journals may be more likely to select manuscripts describing good software \cite{Garfield1955-wf}.
-\textbf{4.
-Author/Group reputation:} the key to any project is the skills of the people involved, including maintaining a high collective intelligence \cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}.
-As a consequence, an argument could be made that well respected and high-profile authors will produce better software \cite{Hirsch2005-mt,Bornmann2008-il}.
-\textbf{5.
-Speed:} software is frequently said to trade accuracy for speed.
-For example, heuristic software such as the popular homology search tool, BLAST, compromise the mathematical guarantee of optimal solutions for more speed \cite{Altschul1990-ht,Altschul1997-ga}.
-Some researchers may naively interpret this fact as slower software is likely to be more accurate.
-But speed may also be influenced by the programming language \cite{Fourment2008-vl}, and/or level of hardware optimisation \cite{Farrar2007-ky,Dematte2010-ph}; In general the implementation is likely to have more of an impact (e.g.
-brute-force approaches versus rapid and sensitive pre-filtering \cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).
+A number of possible predictors of software quality are used by the
+community of computational biology software users. Some accessible,
+quantifiable and frequently used proxies for identifying high quality
+software include: \textbf{1. Recency:} recently published software
+tools may have built upon the results of past work or be an update to
+an existing software. Therefore, these could be more accurate and
+faster. \textbf{2. Wide adoption:} a software tool may be widely used
+because it is fast and accurate, or because it is well-supported and
+user-friendly. In fact, “large user base”, “word-of-mouth”,
+“wide-adoption”, "personal recommendation," and "recommendation from a
+close colleague," are frequent responses to surveys of “how do
+scientists select software?”
+\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
+  impact:} high profile journals are run by editors and reviewers who
+carefully select and curate the best manuscripts. Therefore, high
+impact journals may be more likely to select manuscripts describing
+good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
+  reputation:} the key to any project is the skills of the people
+involved, including maintaining a high collective intelligence
+\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
+consequence, an argument could be made that well respected and
+high-profile authors will produce better software
+\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software is
+frequently said to trade accuracy for speed. For example, heuristic
+software such as the popular homology search tool, BLAST, compromise
+the mathematical guarantee of optimal solutions for more speed
+\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
+interpret this fact as slower software is likely to be more
+accurate. But speed may also be influenced by the programming language
+\cite{Fourment2008-vl}, and/or level of hardware optimisation
+\cite{Farrar2007-ky,Dematte2010-ph}; In general the implementation is
+likely to have more of an impact (e.g. brute-force approaches versus
+rapid and sensitive pre-filtering
+\cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).
 
-Other factors that influence whether a software tool is selected include: whether the documentation is good, user-friendly, word-of-mouth and ``used in a similar analysis'' \cite{Loman2015-bw}.
-This sort of information is not as readily quantifiable as the above measures.
-However, citation metrics may be a useful proxy.
-The word-of-mouth factor may also explain the reason why some software continues to be used, in spite of poor relative performance \cite{Wadi2016-dj}.
+Other factors that are less quantifiable that influence whether a
+software tool is selected include: whether the documentation is good,
+user-friendly, word-of-mouth and ``used in a similar analysis''
+\cite{Loman2015-bw}. This sort of information is not as readily
+quantifiable as the above measures. However, citation metrics may be a
+useful proxy. The word-of-mouth factor may also explain the reason why
+some software continues to be used, in spite of poor relative
+performance \cite{Wadi2016-dj}. We note that none of these factors
+are directly linked to the top ranked features of successful software
+projects, as determined by literature survey \cite{nasir2011critical}.
+
+{\bf EDITS:\\
+  Number of releases. GitHub statistics (https://www.frontiersin.org/articles/10.3389/fbioe.2018.00198/abstract).\\
+  The influence of organizational structure on software quality https://ieeexplore.ieee.org/iel5/4814109/4814110/04814163.pdf
+  }
 
@@
 
-For each of the publications describing these methods, we have (when possible) identified the 2014 journal impact factor (JIF), published by Thomson Reuters \cite{Garfield2006-lv} and the H5-index published by \href{https://scholar.google.co.nz/citations?view_op=top_venues&hl=en}{Google Scholar Metrics}.
+For each of the publications describing these methods, we have (when possible) identified the 2014 journal impact factor (JIF), published by Thomson Reuters \cite{Garfield2006-lv} and the H5-index published by \href{https://scholar.google.co.nz/citations?view_op=top_venues\&hl=en}{Google Scholar Metrics}.
 We have collected the H-indices and M-indices \cite{Hirsch2005-mt} for the corresponding authors for each method, and the number of times the publication(s) associated with a method has been cited using Google Scholar (data collected over a 1 month period in early 2016).
@@
 The results of which are shown in Figure~\ref{fig:speedaccuracy}.
-We identified 21 bins where there was a significant excess or dearth of methods.
+We identified 19 bins where there was a significant excess or dearth of methods.
 For example, there was an excess of “slow and inaccurate” software (Z=1.6, P-value=0.05) and “slow and accurate” software (Z=1.9, P-value=0.03).
 We find that the amount of software classed as “fast and accurate” and “fast and inaccurate” are at approximately the expected proportions based upon the permutation test.
-The number of significant results is in excess and not due to multiple testing, as the probability of finding 21 of 100 tests significant by chance is low (P-value = $2\times 10^{-8}$, exact binomial test) \cite{Moran2003-ve}.
+The number of significant results is in excess and not due to multiple testing, as the probability of finding 19 of 100 tests significant by chance is low (P-value = $5\times 10^{-7}$, exact binomial test).
+
+% binom.test(sigCount,gridRes^2, p = 0.05)
 
@@
 There may be a number of factors that drive this phenomena.
-One likely explanation is that bioinformatic software tools suffer from a form of publication bias \cite{Boulesteix2015-am}.
-Our community of developers, reviewers and editors may be unwilling to publish software that is neither the fastest nor the most accurate (we have anecdotal evidence to support this, N=1).
+One possible explanation for this is that benchmarkers select software that fills the extremes of each accuracy/speed niche.
+This seems unlikely since based upon our word frequency analysis the terms ``broad'', ``comprehensive'', ``inclusive'' and ``thorough'' are over-represented by between 200\% and 4,400\%.
+A more likely explanation, in our opinion, is that bioinformatic software tools suffer from a form of publication-bias \cite{Boulesteix2015-am,Nissen:2016}.
+That is, the probability of a study being published is influenced by the results it contains \cite{sterling1959publication}.
+The community of developers, reviewers and editors appears to be unwilling to publish software that is neither the fastest nor the most accurate (we have anecdotal evidence to support this, N=1).
 If correct, this is unfortunate.
 Some problems that lack fast and accurate solutions force researchers to make unnecessary compromises in terms of accuracy and time.
-If our hypothesis that this underrepresentation is due to publication bias is valid then why is there an enrichment of slow and inaccurate software (P-value=0.05, empirical distributions from permutation tests)?
+If our hypothesis that this underrepresentation is due to publication-bias is valid then why is there an enrichment of slow and inaccurate software (P-value=0.05, empirical distributions from permutation tests)?
 How could these methods be published?
 We have found that slow and inaccurate software is generally published earlier than alternative methods (P=0.007, one-tailed Wilcoxon test), therefore comparisons were not required to publish these tools and the accuracies may have been over-optimistically reported \cite{Boulesteix2015-nc}.
+
+% 0.91	0.0004054328	0.0002104564	broad             100*(0.0004054328/0.0002104564) = 192.6446
+% 2.08	0.0028380296	0.0006656728	comprehensive     100*(0.0028380296/0.0006656728) = 426.34
+% 2.23	0.0002027164	0.0000352306	thorough          100*(0.0002027164/0.0000352306) = 575.3987
+% 3.86	0.0002027164	0.0000046356	inclusive         100*(0.0002027164/0.0000046356) = 4373.035
 
@@
 Linear mixed models of these values also fail to identify predictors of software accuracy, as do methods for combining P-values.
-A great deal of criticism has been leveled specifically at journal impact factors \cite{Editors2006-ue,Rossner2007-yk,Lariviere2010-kx}.
-Our finding show that this measure is not reflective of software quality.
+A great deal of criticism has been leveled specifically at journal impact factors \cite{Editors2006-ue,Rossner2007-yk,Lariviere2010-kx,Brembs:2013}.
+Our finding show that this measure is also not reflective of software quality.
 But neither is the H5 index which is thought to be a more robust measures of journal impact \cite{Harzing2008-wb}.
@@
 
-The lack of any relationship between software speed and accuracy is surprising. 
-The slower software tools are found to be overrepresented at both high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
-Likewise accurate software was found in both high and low speed ranges.
-A simple gedankenexperiment may be sufficient to prove that slow software is less thoroughly tested than fast software.
-This is because typical software development is an iterative process, where methods are refined by successive rounds of coding, testing and evaluation \cite{Wilson2006-ih}.
-We can assume that similar time-spans are spent on most projects, for example, the span of a MSc or PhD degree.
-As a consequence slow software undergoes fewer development cycles than fast methods, and is therefore tested less, often resulting in less accurate and slower software.
-The fact that fast and inaccurate software is relatively rare does support this argument.
-The lack of any further apparent relationship between speed and accuracy implies there are stronger influences on accuracy rankings than speed alone.
+The lack of any relationship between software speed and accuracy is
+surprising.  The slower software tools are overrepresented at both
+high and low levels of accuracy
+(Figure~\ref{fig:speedaccuracy}). Likewise accurate software was found
+in both high and low speed ranges. A gedankenexperiment may be
+sufficient to show that slow software is less thoroughly tested than
+fast software. This is because typical academic software development
+is an iterative process, where methods are refined by successive
+rounds of coding, testing and evaluation \cite{Wilson2006-ih}. We can
+assume that similar time-spans are spent on most projects, for
+example, the span of a PhD degree or Postdoctoral Fellowship.  As a
+consequence, slow software typically undergoes fewer development
+cycles than fast methods, and is tested less, this may
+result in less accurate and slower software. The fact that fast and
+inaccurate software is relatively rare supports this argument.
+%The
+%lack of any further apparent relationship between speed and accuracy
+%implies there are stronger influences on accuracy rankings than speed
+%alone.
 
 \section*{Conclusions}
-Scientific progress is a grinding process made by testing hypotheses using appropriate positive and negative controls, in combination with independent experiments and replication \cite{Ioannidis2005-xh,Moonesinghe2007-kk}.
-We think that software and analysis tools should be no exception to these principles.
-The continual increases in the complexity of software tools creates an increased likelihood that software bugs are introduced \cite{Darriba2015-sz}.
-Scientific software should therefore be thoroughly tested by developers, however, the results of developer and author-derived tests should be treated with caution due to a range of potential conscious and unconscious biases \cite{Norel2011-cq}.
-Our study shows that no commonly-used impact-based metrics are related to software accuracy, therefore the only reasonable way to select software tools is through software benchmarks.
+Scientific progress is a slow process made by testing hypotheses using
+appropriate positive and negative controls, in combination with
+independent experiments and replication
+\cite{Ioannidis2005-xh,Moonesinghe2007-kk}. We propose that software
+and analysis tools should be no exception to these principles. The
+continual increases in the complexity of software tools creates an
+increased likelihood that software bugs are introduced
+\cite{Darriba2015-sz}. Scientific software should therefore be
+thoroughly tested by developers, however, the results of developer and
+author-derived tests should be treated with caution due to a range of
+potential conscious and unconscious biases \cite{Norel2011-cq}. Our
+study shows that no commonly-used impact-based metrics are related to
+software accuracy, therefore the only reasonable way to select
+software tools is through software benchmarks.
 
-We have shown that slow and inaccurate software is typically published early in the development of a field (see Figure S8).
-Subsequent software tools may be high accuracy or high speed.
-However, as further software tools become available, those that do not rank highly in terms of accuracy or speed are under-represented in the literature.
-Our conjecture is that this is due to a publication bias in computational biology software literature, with publishers, editors and reviewers implicitly requiring software to be notable in at least one of these regards in order to be published.
-This hole in the literature leaves an unfortunate gap upon which further software refinements cannot be made.
+We have found that slow and inaccurate software is typically published
+early in the development of a field (see Figure S8). Subsequent
+software tools may be high accuracy or high speed. However, as further
+software tools become available, those that do not rank highly in
+terms of accuracy or speed are under-represented in the
+literature. Our conjecture is that this is due to a publication-bias
+in the computational biology software literature, with publishers, editors
+and reviewers implicitly requiring software to be notable in at least
+one of these regards in order to be published. This hole in the
+literature leaves an unfortunate gap upon which further software
+refinements cannot be made.
 
-The software that could be used in building research workflows can make use of software tools that are of medium accuracy and speed.
-These may prove to be useful rapid pre-filters for more accurate, yet computationally-demanding approaches.
-These could also be used when datasets are too large to be processed by more accurate software tools, i.e.
-when accurate tools are slow or memory hungry \cite{Lowe1997-hq,Weinberg2006-tm}.
-These tools may also serve as an indication that some approaches may not be worth pursuing further.
+The enrichment of slow and inaccurate software may be the primary
+reason for masking the association between software accuracy and speed
+that we had anticipated before embarking upon this study.
 
-We propose that the full spectrum of software tool accuracies and speeds serve a useful purpose to the research community.
-Like negative results, if honestly reported, illustrate to the research community that certain approaches are not practical research avenues \cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
-The classes of tools that we find in the under-represented accuracy and speed region include some of the most used tools in bioinformatics, such as the homology search tools NCBI-BLAST \cite{Altschul1990-ht,Altschul1997-ga} and HMMER2 \cite{Eddy:2009} and genome assembly and mapping tools such as AbySS \cite{Simpson:2009}, MAQ \cite{Li:2008} and Newbler \cite{Margulies:2005}.
-The current practises of publishers, editors, reviewers and authors of software tools therefore deprive our community of tools for building effective and productive workflows.
+The software that could be used in building research workflows can
+make use of software tools that are of intermediate accuracy and
+speed. These may prove to be useful rapid pre-filters for more
+accurate, yet computationally-demanding approaches. These could also
+be used when datasets are too large to be processed by more accurate
+software tools, i.e. when accurate tools are CPU or memory intensive
+\cite{Lowe1997-hq,Weinberg2006-tm}. These tools may also serve as an
+indication that some approaches may not be worth pursuing further.
 
-A potential avenue for further exploration is to compare the starting point for software development projects as certain approaches may produce more rapid gains than others.
-For example, starting with biologically plausible and fast methods will theoretically allow more rapid gains in accuracy through iterative method refinement than starting with slow and mathematically complete approaches.
-At the very least these may be used as rapid data filters for reducing the size and complexity of problems, prior to employing more rigorous methods.
+We propose that the full spectrum of software tool accuracies and
+speeds serve a useful purpose to the research community. Like negative
+results, if honestly reported, illustrate to the research community
+that certain approaches are not practical research avenues
+\cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}. The classes of
+tools that we find in the under-represented accuracy and speed region
+include some of the most widely used and useful tools in
+bioinformatics. These include the homology search tools NCBI-BLAST
+\cite{Altschul1990-ht,Altschul1997-ga} and HMMER2 \cite{Eddy:2009} and
+genome assembly and mapping tools such as AbySS \cite{Simpson:2009},
+MAQ \cite{Li:2008} and Newbler \cite{Margulies:2005}. The current
+practises of publishers, editors, reviewers and authors of software
+tools therefore deprive our community of tools for building effective
+and productive workflows.
 
-We have shown that accurate software is not necessarily the most recently released method or the product of high profile lab groups or selected by high impact journals.
-Software that is widely used or is either slow or fast is also not necessarily the most accurate.
-Therefore, accurate software may be the product of features we have not been able to capture.
-Possibly, hard work, good ideas and sound method testing, as well as technical ability, experience and education levels in software development \cite{Wilson2006-ih}.
+A potential avenue for further exploration is to compare the starting
+point for software development projects as certain approaches may
+produce more rapid gains than others. For example, starting with
+biologically plausible and fast methods will theoretically allow more
+rapid gains in accuracy through iterative method refinement than
+starting with slow and mathematically complete approaches. At the very
+least these may be used as rapid data filters for reducing the size
+and complexity of problems, prior to employing more rigorous methods.
 
-Finally, we think that the field of computational biology could benefit from embracing an increased number of independent software comparison studies \cite{Boulesteix2013-vb}, in parallel with the popular challenge-based benchmarks such as CASP, Assemblathon, Alignathon, DREAM and RGASP \cite{Boutros2014-zm}.
-We are hopeful that this, along with the relaxation of the publication bias we have described, will reduce the over-optimistic and misleading reporting of method accuracy \cite{Boulesteix2010-te,Jelizarow2010-zf,Norel2011-cq}.
+We have shown that accurate software is not necessarily the most
+recently released method or the product of high profile lab groups or
+selected by high impact journals. Software that is widely used or is
+either slow or fast is also not necessarily the most
+accurate. Therefore, accurate software may be the product of features
+we have not been able to capture such as scientific inspiration, hard work,
+sound method testing, technical ability and
+education levels in software development. % \cite{Wilson2006-ih}.
+
+Finally, we think that the field of computational biology could
+benefit from embracing an increased number of independent software
+comparison studies \cite{Boulesteix2013-vb}, in parallel with the
+popular challenge-based benchmarks such as CASP, Assemblathon,
+Alignathon, DREAM and RGASP \cite{Boutros2014-zm}. We are hopeful that
+this, along with steps to reduce the publication-bias we have
+described, will reduce the over-optimistic and misleading reporting of
+method accuracy
+\cite{Boulesteix2010-te,Jelizarow2010-zf,Norel2011-cq}.
 
@@
 \section*{Acknowledgements}
-The authors acknowledge the valued contribution of invaluable discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia Palmer and Jason Tylianakis.
-Michael A.
-Black,  Murray Cox, Raquel Norel, Alexandros Stamatakis, Jens Stoye,Tandy Warnow, provided valuable feedback on drafts of the manuscript.
 
-PPG, FAG and NEW are supported by a Rutherford Discovery Fellowship,
-administered by the Royal Society of New Zealand. NEW is supported by
-a PhD scholarship from the University of Canterbury.
+The authors acknowledge the valued contribution of invaluable
+discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia
+Palmer and Jason Tylianakis. Michael A. Black, Murray Cox, Raquel
+Norel, Alexandros Stamatakis, Jens Stoye,Tandy Warnow, provided
+valuable feedback on drafts of the manuscript.
+
+PPG is supported by a Rutherford Discovery Fellowship,
+administered by the Royal Society of New Zealand.
 

### 2020-07-03 bin/pmArticleScore.pl

@@ -26,7 +26,7 @@
 #PUBMED SEARCHES:
 #                      TRAINING
 #                      less Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ Data.tsv | perl -lane '@F=split(/\t/); print "$F[0]\[uid\] OR " if $F[0]>0' | grep -v 999999999 | tr -d "\n" && echo
-#                      17151342[uid] OR 20047664[uid] OR 25198770[uid] OR 21483869[uid] OR 24526711[uid] OR 24839440[uid] OR 17062146[uid] OR 21423806[uid] OR 25511303[uid] OR 20617200[uid] OR 26778510[uid] OR 25521762[uid] OR 23593445[uid] OR 21525877[uid] OR 24708189[uid] OR 18287116[uid] OR 24602402[uid] OR 24086547[uid] OR 18793413[uid] OR 23393030[uid] OR 22132132[uid] OR 15701525[uid] OR 22152123[uid] OR 19046431[uid] OR 25760244[uid] OR 23758764[uid] OR 22172045[uid] OR 25574120[uid] OR 22506536[uid] OR 21856737[uid] OR 21113338[uid] OR 23842808[uid] OR 15840834[uid] OR 19179695[uid] OR 26862001[uid] OR 22492192[uid] OR 21615913[uid] OR 19126200[uid] OR 22574964[uid]
[…]

### 2021-02-24 manuscript/manuscript-speed-accuracy.tex
Major update: larger dataset, revised and simplified code, new figures.

@@
 
+\def\numTools{499}
+%cat meanRankSpeedData.tsv | cut -f 4 | grep -v ^method$ | wc -l
+\def\numBenchmarkPubs{69}
+%cat speed-vs-accuracy-toolRanks2005-2020.tsv | cut -f 1 | sort  | uniq | grep -E '[0-9]'  | wc -l
+\def\numBenchmarks{134}
+%cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l
+
 \PaperTitle{A meta-analysis of bioinformatics software benchmarks reveals that publication-bias unduly influences software accuracy} % Article title
@@
 \Authors{Paul P.
-Gardner\textsuperscript{1,2,3}*, James M.
-Paterson\textsuperscript{1,2}, Fatemeh Ashari-Ghomi\textsuperscript{1,2}, Sinan U.
-Umu\textsuperscript{1,2}, Stephanie McGimpsey\textsuperscript{4,5}, Aleksandra Pawlik\textsuperscript{6}} % Authors
-\affiliation{\textsuperscript{1}\textit{School of Biological Sciences, University of Canterbury, Christchurch, New Zealand.}} % Author affiliation
+Gardner\textsuperscript{1,2}*, James M.
+Paterson\textsuperscript{?}, Fatemeh Ashari-Ghomi\textsuperscript{?}, Sinan U.
+Umu\textsuperscript{?}, Stephanie McGimpsey\textsuperscript{3,4}, Aleksandra Pawlik\textsuperscript{5},
+Alex Gavryushkin\textsuperscript{6}, Michael A Black\textsuperscript{1}
+} % Authors
+
+\affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
 \affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
-\affiliation{\textsuperscript{3}\textit{Bio-Protection Research Centre, University of Canterbury, Christchurch, New Zealand.}}
-\affiliation{\textsuperscript{4}\textit{McConomy School of Dance, Derry, Ireland.}}
-\affiliation{\textsuperscript{5}\textit{Research for Good, Limavady, Ireland.}}
-\affiliation{\textsuperscript{6}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
-\affiliation{*\textbf{Corresponding author}: paul.gardner@canterbury.ac.nz} % Corresponding author
+\affiliation{\textsuperscript{3}\textit{McConomy School of Dance, Derry, Ireland.}}
+\affiliation{\textsuperscript{4}\textit{Research for Good, Limavady, Ireland.}}
+\affiliation{\textsuperscript{5}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
+\affiliation{\textsuperscript{6}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+\affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author
 
@@
 %\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
-\Abstract{
-Computational biology has provided widely used and powerful software tools for testing and making inferences about biological data.
-In the face of increasing volumes of data, heuristic methods that trade software speed for mathematical completeness must be employed.
-We are interested in whether trade-offs between speed and accuracy are reasonable. Also, what factors are indicative of accurate software?
+\Abstract{ Computational biology has provided widely used and powerful
+  software tools for testing and making inferences about biological
+  data. In the face of rapidly increasing volumes of data, heuristic
+  methods that trade software speed for mathematical completeness must
+  be employed.  We are interested in testing for trade-offs between
+  speed and accuracy and for other factors are indicative of accurate
+  software.
 
-In this work we mine published benchmarks of computational biology software, we collect data on the relative accuracy and speed of different software and then test to see what factors influence accuracy e.g.
-speed, author reputation, journal impact or recency.
+We have extracted accuracy and speed ranks from independent benchmarks of
+computational biology software of different software tools, and evaluate the
+factors that are likely to influence accuracy e.g. speed, author
+reputation, journal impact, recency or developer efforts.
 
-We found that author reputation, journal impact, the number of citations, software speed and age are not reliable predictors of software accuracy.
-Unfortunately, citations, author and journal reputation are frequently cited reasons for selecting software tools.
-Our results imply that useful bioinformatics software is not only the domain of famous senior researchers.
-In addition, we found that there is an excess of slow and inaccurate software tools across multiple sub-disciplines of bioinformatics.
-Meanwhile, there are very few tools of middling accuracy and speed.
-We hypothesise that a form of publication-bias unduly influences the publication and development of bioinformatic software tools.
-In other words, at present software that is not highly ranked on speed and not highly ranked on accuracy is difficult to publish due to editorial and reviewer practices.
-This leaves an unfortunate gap in the literature upon which future software refinements cannot be constructed.
-}
+We found that software speed, author reputation, journal impact, the
+number of citations and age are all unreliable predictors of software
+accuracy. This is unfortunate because citations, author and journal
+reputation are frequently used reasons for selecting software
+tools. Our results do show that accurate bioinformatic software tools
+are generally the product many gradual improvements, often from
+multiple developers, based upon github records. In addition, we find
+that bioinformatics has an excess of slow and inaccurate software
+tools across many sub-disciplines. Meanwhile, there are few tools that
+are middle-of-road in terms of accuracy and speed trade-offs. We
+hypothesise that a form of publication-bias unduly influences the
+publication and development of bioinformatic software tools. In other
+words, at present software that is not highly ranked on speed and not
+highly ranked on accuracy is difficult to publish due to author,
+editor and reviewer practices. This leaves an unfortunate gap in the
+literature upon which future software refinements cannot be
+constructed.  }
 
@@
 Computational biology software is widely used and has produced some of
-the most cited publications in the scientific corpus
+the most cited publications in the entire scientific corpus
 \cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. This
-software includes implementations of methods for sequence alignment
-and homology inference
+highly-cited software includes implementations of methods for sequence
+alignment and homology inference
 \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
@@
 \cite{Kraulis1991-lt,Berman2000-to}. However, the popularity of a
-software tool does not necessarily imply that it is either accurate or
+software tool does not necessarily mean that it is accurate or
 computationally efficient, instead usability, ease of installation,
-operating system and other factors may play a greater role. There have
-been several notable incidences where inaccurate software has caused
-considerable harm \cite{leveson1993investigation}.
+operating system and other factors may play a greater role in a software
+tool's popularity. There have been several notable incidences where
+inaccurate software has caused considerable harm
+\cite{leveson1993investigation}.
 %Fergusson covid predictions?
@@
 reducing the over-optimistic reporting of software accuracy
-\cite{Boulesteix2010-te,Jelizarow2010-zf} and the self-assessment trap
-\cite{Norel2011-cq}. Benchmark studies typically use a number of
-positive and negative control datasets, predictions can then be
+\cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
+\cite{Norel2011-cq,Buchka:2021}. Benchmark studies typically use a number of
+positive and negative control datasets, predictions from different software tools can then be
 partitioned into true or false groups and a variety of metrics can be
-used to evaluate the performance of different predictions
-\cite{Egan1975-nd,Hall2012-kg}. Some benchmarks now use live, or
-frequently updated, results to indicate the latest developments in
-software performance
-\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}. The aim of
-these benchmarks is to independently identify tools that make
-acceptable compromises in terms of scoring schemes and the resulting
-potential for false predictions, true predictions and speed, and are
-therefore suited for wide adoption by the community.
+used to evaluate the performance differences
+\cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
+%Some benchmarks now use live, or
+%frequently updated, results to indicate the latest developments in
+%software performance
+%\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
+The aim of these benchmarks is to independently identify tools that
+make acceptable compromises in terms of balancing speed with
+discriminating true and false predictions, and are therefore suited
+for wide adoption by the community.
 
@@
 \cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut,Wikipedia_contributors2015-vj,Wikipedia_contributors2015-hr}. While
-this may generally be a good problem to have, and points to a
+this is a good problem to have, and points to a
 diversity of options from which practical solutions can be selected,
@@
 any recent gold-standard benchmarks, how should scientific software be
-selected? In the following work, we presume that ``biological
-accuracy'' is the most desirable feature of software. Biological
+selected? In the following we presume that ``biological
+accuracy'' is the most desirable feature for a software tool. Biological
 accuracy is the degree to which predictions or measurements reflect
-the truths of biological systems, this is usually determined by
-comparing results of software to the results of established reference
-data. These are commonly expert-derived curated datasets. In some
+the biological truths based on expert-derived curated datasets. In some
 fields biological accuracy is very difficult to ascertain, for
 example, in phylogenetics it is nearly impossible to know the
-ancestral relationships between organisms. In situations like this,
-researchers can use a mix of simulated or high-confidence datasets.
+true ancestral relationships between organisms. In situations like this,
+researchers may use simulated or high-confidence datasets.
 
@@
 high-profile authors will produce better software
-\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software is
-frequently said to trade accuracy for speed. For example, heuristic
+\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
+frequently trade accuracy for speed. For example, heuristic
 software such as the popular homology search tool, BLAST, compromise
@@
 accurate. But speed may also be influenced by the programming language
-\cite{Fourment2008-vl}, and/or level of hardware optimisation
-\cite{Farrar2007-ky,Dematte2010-ph}; In general the implementation is
+\cite{Fourment2008-vl}, and the level of hardware optimisation
+\cite{Farrar2007-ky,Dematte2010-ph}; however, the implementation is
 likely to have more of an impact (e.g. brute-force approaches versus
@@
 user-friendly, word-of-mouth and ``used in a similar analysis''
-\cite{Loman2015-bw}. This sort of information is not as readily
-quantifiable as the above measures. However, citation metrics may be a
-useful proxy. The word-of-mouth factor may also explain the reason why
-some software continues to be used, in spite of poor relative
-performance \cite{Wadi2016-dj}. We note that none of these factors
-are directly linked to the top ranked features of successful software
-projects, as determined by literature survey \cite{nasir2011critical}.
+\cite{Loman2015-bw}.
+%This information is not as readily
+%quantifiable as the above measures. However,
+However, citation metrics may be a
+useful proxy for the above.
+%This may also explain the reason why
+%some software continues to be used, in spite of poor relative
+%performance \cite{Wadi2016-dj}.
+%We note that none of these factors
+%are directly linked to the top ranked features of successful software
+%projects, as based upon literature surveys \cite{nasir2011critical}.
 
-{\bf EDITS:\\
-  Number of releases. GitHub statistics (https://www.frontiersin.org/articles/10.3389/fbioe.2018.00198/abstract).\\
-  The influence of organizational structure on software quality https://ieeexplore.ieee.org/iel5/4814109/4814110/04814163.pdf
-  }
+With the wide adoption of Github ($\approx 47\%$ of the
+\textbf{{\color{red}\numTools}} tools included in this study could be
+linked to a Github repository), quantifiable data on software
+development such as the number of contributors to code, number of code
+changes and versions is also available
+\cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.
 
-In the following study, we explore factors that may be indicative of software accuracy.
-This, in our opinion, should be one of the prime reasons for selecting a software tool.
-We have mined the large and freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks of computational biology software, and manually extracted accuracy and speed rankings for \textbf{{\color{black}243}} software packages.
-For each software tool, we have collected measures that may be predictive of accuracy, and may be subjectively employed by the researcher community as a proxy for software quality.
-These include relative speed, relative age, the productivity and impact of the corresponding authors, journal impact and the number of citations.
+In the following study, we explore factors that may be indicative of
+software accuracy. This, in our opinion, should be one of the prime
+reasons for selecting a software tool. We have mined the large and
+freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks
+of computational biology software, and manually extracted accuracy and
+speed rankings for \textbf{{\color{red}\numTools~}} unique software tools. For
+each tool, we have collected measures that may be predictive
+of accuracy, and may be subjectively employed by the researcher
+community as a proxy for software quality. These include relative
+speed, relative age, the productivity and impact of the corresponding
+authors, journal impact and the number of citations.
 
 \section*{Results}
-We have collected relative accuracy and speed ranks for \textbf{{\color{black}243}} distinct software tools.
-This software has been developed for solving a broad cross-section computational biology tasks.
-These include software for homology search \cite{Freyhult2007-et}, genome sequence analysis (e.g.
-read mapping or sequence assembly) \cite{Junemann2014-mb,Tran2014-pe,Zhang2011-nd,Abbas2014-gu,Bao2011-lv,Caboche2014-lj,Kleftogiannis2013-wi,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Yang2013-aj,Holtgrewe2011-fd,Rackham2015-ag,Huang2015-wu}, multiple sequence alignment \cite{Thompson2011-rf,Nuin2006-nk,Pais2014-sr,Pervez2014-zp,Liu2010-rp}, cell tracking \cite{Maska2014-ak}, transcriptome analysis \cite{Li2012-wr,Lu2013-fs,Liu2014-kz,Kumar2016-xz}, RNA interaction prediction \cite{Pain2015-gr}, protein interactions \cite{Tikk2010-qd}, protein structure prediction \cite{Kolodny2005-ry,Wallner2005-qi}, epistasis \cite{Shang2011-vy},  metagenomic analysis \cite{Lindgreen2016-tt,Bazinet2012-wf}, repetitive sequence prediction \cite{Saha2008-kd}, proteomics \cite{Lange2008-pt,Yang2009-oc} and phylogenetics \cite{Liu2011-pz,Yang2011-dv,Oscamou2008-md,Bayzid2013-hc,Liu2009-lx}.
-Each software tool was benchmarked in at least one of \textbf{{\color{black}43}} publications that satisfy the Boulesteix criteria \cite{Boulesteix2013-vb}.
-In brief, the Boulesteix criteria are: 1.
-the main focus of the article is a benchmark.
-2.
-the authors are reasonably neutral.
-3.
-the test data and evaluation criteria are sensible.
+We have collected relative accuracy and speed ranks for
+\textbf{{\color{red}\numTools~}} distinct software tools. This software has
+been developed for solving a broad cross-section computational biology
+tasks.
+%% These include software for homology search
+%% \cite{Freyhult2007-et}, genome sequence analysis (e.g. read mapping or
+%% sequence assembly)
+%% \cite{Junemann2014-mb,Tran2014-pe,Zhang2011-nd,Abbas2014-gu,Bao2011-lv,Caboche2014-lj,Kleftogiannis2013-wi,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Yang2013-aj,Holtgrewe2011-fd,Rackham2015-ag,Huang2015-wu},
+%% multiple sequence alignment
+%% \cite{Thompson2011-rf,Nuin2006-nk,Pais2014-sr,Pervez2014-zp,Liu2010-rp},
+%% cell tracking \cite{Maska2014-ak}, transcriptome analysis
+%% \cite{Li2012-wr,Lu2013-fs,Liu2014-kz,Kumar2016-xz}, RNA interaction
+%% prediction \cite{Pain2015-gr}, protein interactions
+%% \cite{Tikk2010-qd}, protein structure prediction
+%% \cite{Kolodny2005-ry,Wallner2005-qi}, epistasis \cite{Shang2011-vy},
+%% metagenomic analysis \cite{Lindgreen2016-tt,Bazinet2012-wf},
+%% repetitive sequence prediction \cite{Saha2008-kd}, proteomics
+%% \cite{Lange2008-pt,Yang2009-oc} and phylogenetics
+%% \cite{Liu2011-pz,Yang2011-dv,Oscamou2008-md,Bayzid2013-hc,Liu2009-lx}.
+Each
+software tool was benchmarked in at least one of
+\textbf{{\color{red}\numBenchmarkPubs~}} publications that satisfy the Boulesteix
+criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria
+are: 1. the main focus of the article is a benchmark. 2. the authors
+are reasonably neutral. 3. the test data and evaluation criteria are
+sensible.
 
-For each of the publications describing these methods, we have (when possible) identified the 2014 journal impact factor (JIF), published by Thomson Reuters \cite{Garfield2006-lv} and the H5-index published by \href{https://scholar.google.co.nz/citations?view_op=top_venues\&hl=en}{Google Scholar Metrics}.
-We have collected the H-indices and M-indices \cite{Hirsch2005-mt} for the corresponding authors for each method, and the number of times the publication(s) associated with a method has been cited using Google Scholar (data collected over a 1 month period in early 2016).
+For each of the publications describing these tools, we have (where
+possible) the H5-index published by
+\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
+  Scholar Metrics}. We have collected the maximum H-index and
+corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
+authors for each tool, and the number of times the publication(s)
+associated with a tool has been cited using Google Scholar (data
+collected over a 6 month period in late 2020). Note that the citation metrics
+are not static and will change over time.
 
-We have computed the Spearman’s correlation coefficient for each pairwise combination of the mean normalised accuracy and speed ranks, the year published, mean relative age (compared to software in the same benchmarks), journal IF and H5 metrics, the total number of citations, the relative number of citations (compared to software in the same benchmarks) and the maximum H and M indices for the corresponding authors.
-The results are presented in Figure~\ref{fig:allfactors}A.
-We found significant associations between most of the citation-based metrics (journal H5, JIF, citations, relative citations, H-index and M-index).
-There is also a strong association between the year of publication, the relative age and many of the citation-based metrics.
+We have computed the Spearman’s correlation coefficient for each
+pairwise combination of the mean normalised accuracy and speed ranks,
+the year published, mean relative age (compared to software in the
+same benchmarks), journal H5 metrics, the total number of citations,
+the relative number of citations (compared to software in the same
+benchmarks) and the maximum H- and corresponding M-indices for the corresponding
+authors. Where possible we also extract the version number, the number
+of commits and number of contributors from Github repositories.  The
+results are presented in Figure~\ref{fig:allfactors}A. We find
+significant associations between most of the citation-based metrics
+(journal H5, citations, relative citations, H-index and
+M-index). There is also a negative correlation between the year of
+publication, the relative age and many of the citation-based metrics.
 
@@
 \includegraphics[width=\textwidth]{figure1.pdf}
-\caption{\textbf{A.} A heatmap indicating the relationships between proposed predictors of software quality.
-Spearman’s rho is used to infer correlations between metrics such as the H and M indices of corresponding authors, number of citations, journal impact factors and H5 indices, the year and relative age of software and the mean relative rankings of software speed and accuracy.
-Red colours indicate a high positive correlation, blue colours indicate a high negative correlation.
-Correlations with a P-value less than 0.05 are indicated with a ‘X’.
-The dashed rectangular area is illustrated in more detail in \textbf{B}, the bold square is shown in more detail in \textbf{Figure 2}.The dendrogram was computed using the default ‘heatmap.2 function in R, which computes a Euclidean distance matrix and a complete-linkage hierarchical clustering.
-\textbf{B.} A barplot illustrating the correlation as measured by Spearman’s rho between and normalised accuracy ranks and potential factors that may be predictive of accuracy.
-In order to give an appreciation of the difference between the observed effect-sizes and significant effect sizes, we generated 10,000 permuted accuracy ranks for each benchmark and recorded Spearman’s rho for the significant correlations (P<0.05).
-These values are marked with “x”s in the barplot.}
+\caption{\textbf{A.} A heatmap indicating the relationships between
+  proposed predictors of software quality. Spearman’s rho is used to
+  infer correlations between metrics such as the H- and M-indices of
+  corresponding authors, number of citations, journal impact factors
+  and H5 indices, the year and relative age of software and the mean
+  relative rankings of software speed and accuracy. Red colours
+  indicate a positive correlation, blue colours indicate a negative
+  correlation. Correlations with a P-value less than 0.05 are
+  indicated with a ‘X’. The dashed rectangular area is illustrated in
+  more detail in \textbf{B}, the relationship between speed and
+  accuracy is shown in more detail in \textbf{Figure 2}.
+  %The
+  %dendrogram was computed using the default ‘heatmap.2 function in R,
+  %which computes a Euclidean distance matrix and a complete-linkage
+  %hierarchical clustering.
+  \textbf{B.} A barplot illustrating the correlation, as measured by
+  Spearman’s rho, between normalised accuracy ranks and software tool
+  features that may be predictive of accuracy. In order to give an
+  appreciation of the difference between the observed effect-sizes and
+  significant effect sizes, we generated 1000 permuted accuracy ranks
+  for each benchmark and recorded Spearman’s rho for the significant
+  correlations (P<0.05). These values are marked with “x”s in the
+  barplot. Significant correlations are marked with red asterisks.}
 \label{fig:allfactors}
@@
 
-We found that author reputation metrics, journal impacts and the age of methods were \textbf{not} significantly correlated with either method accuracy or speed (see Figure~\ref{fig:allfactors}).
-The strongest association was between accuracy and journal impact factor (Spearman’s rho = 0.1, P-value = 0.16).
-A linear model of these parameters and accuracy also failed to identify a correlation between these (accuracy: R2 = -0.03, P-value = 0.94; speed: R2 = 0.04, P-value = 0.66).
-To further validate this result, we compute a correlation between speed and accuracy for each benchmark and used weighted sum Z-tests \cite{Zaykin2011-tj}.
-This also failed to identify a significant relationship (sum Z=-0.1, P-value=0.5).
+Data on the
+number of updates to software tools from Github such as the number of versions,
+commits and contributors \textbf{{\color{red}was significantly correlated with 
+accuracy (respective Spearman’s rhos = 0.13, 0.22, 0.21, respective P-values = 0.0027, 0.00079, 0.0013). 
+These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A).
+}}
+We also found that author reputation metrics, journal impacts and the age
+of tools were generally \textbf{not} correlated with either tool
+accuracy or speed (see Figure~\ref{fig:allfactors}). The strongest
+association was between accuracy and \textbf{{\color{red}the relative
+    number of citations}} (Spearman’s rho = \textbf{{\color{red}0.092,
+    P-value = 0.047}}). But the effect size is only just over the
+significance threshold and may be explained by multiple-testing.
 
-In order to gain a deeper understanding of the distribution of available bioinformatic software tools on a speed versus accuracy landscape, we ran a Monte Carlo permutation test.
-The ranks extracted from each benchmark were randomly permuted, generating 10,000 randomized speed and accuracy ranks.
-In the cells of a $10\times10$ grid spanning the normalised speed and accuracy ranks we computed a Z-score for the observed number of methods in a cell, compared to the expected distributions generated by the randomized ranks.
-The results of which are shown in Figure~\ref{fig:speedaccuracy}.
-We identified 19 bins where there was a significant excess or dearth of methods.
-For example, there was an excess of “slow and inaccurate” software (Z=1.6, P-value=0.05) and “slow and accurate” software (Z=1.9, P-value=0.03).
-We find that the amount of software classed as “fast and accurate” and “fast and inaccurate” are at approximately the expected proportions based upon the permutation test.
-The number of significant results is in excess and not due to multiple testing, as the probability of finding 19 of 100 tests significant by chance is low (P-value = $5\times 10^{-7}$, exact binomial test).
+%To further validate this result, we compute a correlation
+%between speed and accuracy for each benchmark and used weighted sum
+%Z-tests \cite{Zaykin2011-tj}. This also failed to identify a
+%significant relationship (sum Z=-0.1, P-value=0.5).
+
+In order to gain a deeper understanding of the distribution of
+available bioinformatic software tools on a speed versus accuracy
+landscape, we ran a permutation test. The ranks extracted
+from each benchmark were randomly permuted, generating 1,000
+randomized speed and accuracy ranks. In the cells of a $10\times10$
+grid spanning the normalised speed and accuracy ranks we computed a
+Z-score for the observed number of tools in a cell, compared to the
+expected distributions generated by 1,000 randomized ranks. The results
+of which are shown in Figure~\ref{fig:speedaccuracy}. We identified \textbf{{\color{red}18}}
+bins where there was a significant excess or dearth of tools. For
+example, there was an excess of “slow and inaccurate” software (\textbf{{\color{red}Z=2.50,
+P-value=0.0063}}). We find that the amount of software classed as “fast
+and accurate” and “fast and inaccurate” are at approximately the
+expected proportions based upon the permutation test. The number of
+significant results is in excess, and is not due to multiple testing, as
+the probability of finding \textbf{{\color{red}18}} of 100 tests significant by chance is
+low (P-value = \textbf{{\color{red}$2.2\times 10^{-6}$}}, exact binomial test).
 
@@
 \includegraphics[width=\textwidth]{figure2.pdf}
-\caption{A heatmap indicating the relative paucity or abundance of software in the range of possible accuracy and speed rankings.
-Red colours indicate an abundance of software tools in an accuracy and speed category, while blue colours indicate scarcity of software in an accuracy and speed category.
-The abundance is quantified using a Z-score computation for each bin, this is derived from 10,000 random permutations of the speed and accuracy ranks from each benchmark.
-Mean normalised ranks of accuracy and speed have been binned into 100 classes (a $10\times10$ grid) that range from comparatively slow/inaccurate to comparatively fast/accurate.
-Z-scores with a P-value less than 0.05 are indicated with a ‘X’.}
+\caption{A heatmap indicating the relative paucity or abundance of
+  software in the range of possible accuracy and speed rankings. Red
+  colours indicate an abundance of software tools in an accuracy and
+  speed category, while blue colours indicate scarcity of software in
+  an accuracy and speed category. The abundance is quantified using a
+  Z-score computation for each bin, this is derived from 10,000 random
+  permutations of the speed and accuracy ranks from each
+  benchmark. Mean normalised ranks of accuracy and speed have been
+  binned into 100 classes (a $10\times10$ grid) that range from
+  comparatively slow/inaccurate to comparatively
+  fast/accurate. Z-scores with a P-value less than 0.05 are indicated
+  with a ‘X’.}
 \label{fig:speedaccuracy}
@@
 
-There is a large reduction in the number of software tools that are classed as intermediate in terms of both speed and accuracy based upon our permutation test (Figure~\ref{fig:speedaccuracy}).
-The cells corresponding to the four central underrepresented deciles are highlighted (Z = -1.7, -2.1, -2.1 and -2.4, P-values = 0.04, 0.02, 0.02 and 0.008, respectively, reading from top to bottom, left to right).
-We also tested the relative age of the software tools in significantly over- or under-represented regions of the speed vs accuracy plot.
-We found that the “slow and inaccurate” methods were generally published earlier than other methods (W = 31, P=0.007, one-tailed Wilcoxon test) (Figure S8).
+There is a major reduction in the number of software tools that are
+classed as intermediate in terms of both speed and accuracy based on
+permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The cells
+corresponding to the four central deciles are
+highlighted \textbf{{\color{red}(Z = -1.9, -2.5, -2.9 and -1.1, P-values = 0.03, 0.006,
+0.002 and 0.14}}, respectively, reading from top to bottom, left to
+right). We also tested the relative age of the software tools in
+the indicated corners and central regions of the speed vs
+accuracy plot. We found that the “slow and inaccurate” tools ($N=27$) were
+generally published earlier than ``fast and accurate'' tools ($N=23$) \textbf{{\color{red}(W = 341, P=0.005,
+one-tailed Wilcoxon test)}} (Figure S7).
 
-\section*{Discussion}
+\section*{Conclusion}
 
-We have gathered data on the relative speed and accuracies of \textbf{{\color{black}243}} bioinformatic methods from \textbf{{\color{black}43}} benchmarks that were published between 2005 and 2016.
-We show that there is an under-representation of software that has both intermediate levels of accuracy and speed.
-There may be a number of factors that drive this phenomena.
-One possible explanation for this is that benchmarkers select software that fills the extremes of each accuracy/speed niche.
-This seems unlikely since based upon our word frequency analysis the terms ``broad'', ``comprehensive'', ``inclusive'' and ``thorough'' are over-represented by between 200\% and 4,400\%.
-A more likely explanation, in our opinion, is that bioinformatic software tools suffer from a form of publication-bias \cite{Boulesteix2015-am,Nissen:2016}.
-That is, the probability of a study being published is influenced by the results it contains \cite{sterling1959publication}.
-The community of developers, reviewers and editors appears to be unwilling to publish software that is neither the fastest nor the most accurate (we have anecdotal evidence to support this, N=1).
-If correct, this is unfortunate.
-Some problems that lack fast and accurate solutions force researchers to make unnecessary compromises in terms of accuracy and time.
-If our hypothesis that this underrepresentation is due to publication-bias is valid then why is there an enrichment of slow and inaccurate software (P-value=0.05, empirical distributions from permutation tests)?
-How could these methods be published?
-We have found that slow and inaccurate software is generally published earlier than alternative methods (P=0.007, one-tailed Wilcoxon test), therefore comparisons were not required to publish these tools and the accuracies may have been over-optimistically reported \cite{Boulesteix2015-nc}.
+We have gathered data on the relative speeds and accuracies of
+\textbf{{\color{red}\numTools~}} bioinformatic tools from
+\textbf{{\color{red}\numBenchmarkPubs}} benchmarks published between
+2005 and 2020. {\bf Our results suggest there are major benefits for
+  long term support for software \cite{siepel2019challenges}.} The
+finding of a strong relationship between the number of commits to
+github (i.e. software updates) and accuracy, highlights the benefits of
+long-term development.
 
-% 0.91	0.0004054328	0.0002104564	broad             100*(0.0004054328/0.0002104564) = 192.6446
-% 2.08	0.0028380296	0.0006656728	comprehensive     100*(0.0028380296/0.0006656728) = 426.34
-% 2.23	0.0002027164	0.0000352306	thorough          100*(0.0002027164/0.0000352306) = 575.3987
-% 3.86	0.0002027164	0.0000046356	inclusive         100*(0.0002027164/0.0000046356) = 4373.035
+Our study finds little evidence to support that impact-based metrics
+are poorly related to software quality, this is unfortunate as these
+are frequently cited reasons for selecting software tools
+\cite{Loman2015-bw}. This implies that the recorded high citation
+rates for bioinformatic software
+\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
+reflection of user-friendliness and possibly the Matthew Effect
+\cite{Lariviere2010-kx,Merton1968-cb} than accuracy.
 
-We found that no commonly used proxy for study quality is correlated with software accuracy.
-Neither, author reputation, number of citations, journal impact, relative age or speed appear to be associated with accurate software.
-Linear mixed models of these values also fail to identify predictors of software accuracy, as do methods for combining P-values.
-A great deal of criticism has been leveled specifically at journal impact factors \cite{Editors2006-ue,Rossner2007-yk,Lariviere2010-kx,Brembs:2013}.
-Our finding show that this measure is also not reflective of software quality.
-But neither is the H5 index which is thought to be a more robust measures of journal impact \cite{Harzing2008-wb}.
-The poor relationship between accuracy and both author reputation and the number of citations is particularly troubling.
-Both are related to “word of mouth” and “previously used in a similar analysis” which a recent survey of researchers suggested is a major influence on the selection of software tools \cite{Loman2015-bw}.
-This implies that the recorded high citation rates for bioinformatic software  \cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a reflection of user-friendliness and the Matthew Effect \cite{Lariviere2010-kx,Merton1968-cb}.
-The Matthew Effect is a biblical reference that can be paraphrased as “the rich get richer and the poor get poorer”.
-In this context, highly cited software is more likely to be used and cited again, irrespective of relative performance.
+We found the lack of a correlation between software speed and accuracy
+surprising.  The slower software tools are overrepresented at both
+high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
+In addition, we find an under-representation of software that has
+intermediate levels of both accuracy and speed. A possible explanation
+for this is that bioinformatic software tools are bound by a form of
+publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
+probability that a study being published is influenced by the results
+it contains \cite{sterling1959publication}. The community of
+developers, reviewers and editors may be unwilling to publish software
+that is not highly ranked on speed or accuracy. If correct, this may
+have unfortunate consequences.  We also found that slow and inaccurate
+software is generally published earlier than fast and accurate tools
+\textbf{{\color{red}(P=0.007, one-tailed Wilcoxon test)}}, therefore
+comparisons were not required to publish the slow and
+inaccurate tools.
 
-The lack of any relationship between software speed and accuracy is
-surprising.  The slower software tools are overrepresented at both
-high and low levels of accuracy
-(Figure~\ref{fig:speedaccuracy}). Likewise accurate software was found
-in both high and low speed ranges. A gedankenexperiment may be
-sufficient to show that slow software is less thoroughly tested than
-fast software. This is because typical academic software development
-is an iterative process, where methods are refined by successive
-rounds of coding, testing and evaluation \cite{Wilson2006-ih}. We can
-assume that similar time-spans are spent on most projects, for
-example, the span of a PhD degree or Postdoctoral Fellowship.  As a
-consequence, slow software typically undergoes fewer development
-cycles than fast methods, and is tested less, this may
-result in less accurate and slower software. The fact that fast and
-inaccurate software is relatively rare supports this argument.
-%The
-%lack of any further apparent relationship between speed and accuracy
-%implies there are stronger influences on accuracy rankings than speed
-%alone.
+%% A
+%% gedankenexperiment may be sufficient to show that slow software is
+%% less thoroughly tested than fast software. This is because typical
+%% academic software development is an iterative process, where tools are
+%% refined by successive rounds of coding, testing and evaluation
+%% \cite{Wilson2006-ih}. We can assume that similar time-spans are spent
+%% on most projects, for example, the span of a PhD degree or
+%% Postdoctoral Fellowship.  As a consequence, slow software may undergo
+%% fewer development cycles than faster tools.
 
-\section*{Conclusions}
-Scientific progress is a slow process made by testing hypotheses using
-appropriate positive and negative controls, in combination with
-independent experiments and replication
-\cite{Ioannidis2005-xh,Moonesinghe2007-kk}. We propose that software
-and analysis tools should be no exception to these principles. The
-continual increases in the complexity of software tools creates an
-increased likelihood that software bugs are introduced
-\cite{Darriba2015-sz}. Scientific software should therefore be
-thoroughly tested by developers, however, the results of developer and
-author-derived tests should be treated with caution due to a range of
-potential conscious and unconscious biases \cite{Norel2011-cq}. Our
-study shows that no commonly-used impact-based metrics are related to
-software accuracy, therefore the only reasonable way to select
-software tools is through software benchmarks.
-
-We have found that slow and inaccurate software is typically published
-early in the development of a field (see Figure S8). Subsequent
-software tools may be high accuracy or high speed. However, as further
-software tools become available, those that do not rank highly in
-terms of accuracy or speed are under-represented in the
-literature. Our conjecture is that this is due to a publication-bias
-in the computational biology software literature, with publishers, editors
-and reviewers implicitly requiring software to be notable in at least
-one of these regards in order to be published. This hole in the
-literature leaves an unfortunate gap upon which further software
-refinements cannot be made.
-
-The enrichment of slow and inaccurate software may be the primary
-reason for masking the association between software accuracy and speed
-that we had anticipated before embarking upon this study.
-
-The software that could be used in building research workflows can
-make use of software tools that are of intermediate accuracy and
-speed. These may prove to be useful rapid pre-filters for more
-accurate, yet computationally-demanding approaches. These could also
-be used when datasets are too large to be processed by more accurate
-software tools, i.e. when accurate tools are CPU or memory intensive
-\cite{Lowe1997-hq,Weinberg2006-tm}. These tools may also serve as an
-indication that some approaches may not be worth pursuing further.
+{\bf STUDY LIMITATIONS...}
 
@@
 that certain approaches are not practical research avenues
-\cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}. The classes of
-tools that we find in the under-represented accuracy and speed region
-include some of the most widely used and useful tools in
-bioinformatics. These include the homology search tools NCBI-BLAST
-\cite{Altschul1990-ht,Altschul1997-ga} and HMMER2 \cite{Eddy:2009} and
-genome assembly and mapping tools such as AbySS \cite{Simpson:2009},
-MAQ \cite{Li:2008} and Newbler \cite{Margulies:2005}. The current
+\cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
+The current
 practises of publishers, editors, reviewers and authors of software
@@
 
-A potential avenue for further exploration is to compare the starting
-point for software development projects as certain approaches may
-produce more rapid gains than others. For example, starting with
-biologically plausible and fast methods will theoretically allow more
-rapid gains in accuracy through iterative method refinement than
-starting with slow and mathematically complete approaches. At the very
-least these may be used as rapid data filters for reducing the size
-and complexity of problems, prior to employing more rigorous methods.
-
-We have shown that accurate software is not necessarily the most
-recently released method or the product of high profile lab groups or
-selected by high impact journals. Software that is widely used or is
-either slow or fast is also not necessarily the most
-accurate. Therefore, accurate software may be the product of features
-we have not been able to capture such as scientific inspiration, hard work,
-sound method testing, technical ability and
-education levels in software development. % \cite{Wilson2006-ih}.
-
-Finally, we think that the field of computational biology could
-benefit from embracing an increased number of independent software
-comparison studies \cite{Boulesteix2013-vb}, in parallel with the
-popular challenge-based benchmarks such as CASP, Assemblathon,
-Alignathon, DREAM and RGASP \cite{Boutros2014-zm}. We are hopeful that
+The most reliable way to select software tools is through neutral
+software benchmarks \cite{Boulesteix2013-vb}. We are hopeful that
 this, along with steps to reduce the publication-bias we have
 described, will reduce the over-optimistic and misleading reporting of
-method accuracy
-\cite{Boulesteix2010-te,Jelizarow2010-zf,Norel2011-cq}.
+tool accuracy \cite{Boulesteix2010-te,Jelizarow2010-zf,Norel2011-cq}.
 
 \section*{Methods}
-In order to evaluate predictors of computational biology software accuracy, we mined the published literature, extracted data from articles, connected these with bibliometric databases, and tested for correlates with accuracy.
-We outline these steps in further detail below.
+In order to evaluate predictors of computational biology software
+accuracy, we mined the published literature, extracted data from
+articles, connected these with bibliometric databases, and tested for
+correlates with accuracy. We outline these steps in further detail
+below.
 
-\textbf{Criteria for inclusion:} We are interested in using computational biology benchmarks that satisfy Anne-Laure Boulesteix’s (ALB) three criteria for a “neutral comparison study” \cite{Boulesteix2013-vb}.
-Firstly, the main focus of the article is the comparison and \textbf{not} the introduction of a new method.
-Secondly, the authors should be reasonably neutral, which means that the authors should not generally have been involved in the development of the methods included in the benchmark.
-Thirdly, the test data and evaluation criteria should be sensible.
-This means that the test data should be independent of data that methods have been trained upon, and that the evaluation measures appropriately quantify correct and incorrect predictions.
+\textbf{Criteria for inclusion:} We are interested in using
+computational biology benchmarks that satisfy Anne-Laure Boulesteix’s
+(ALB) three criteria for a “neutral comparison study”
+\cite{Boulesteix2013-vb}. Firstly, the main focus of the article is
+the comparison and \textbf{not} the introduction of a new
+tool. Secondly, the authors should be reasonably neutral, which means
+that the authors should not generally have been involved in the
+development of the tools included in the benchmark. Thirdly, the test
+data and evaluation criteria should be sensible. This means that the
+test data should be independent of data that tools have been trained
+upon, and that the evaluation measures appropriately quantify correct
+and incorrect predictions.
 
-Literature mining: We identified an initial list of 10 benchmark articles that satisfy the ALB-criteria.
-These were identified based upon previous knowledge of published articles and were supplemented with several literature searches (e.g.
-‘“benchmark” AND “cputime”’ was used to query both GoogleScholar and Pubmed \cite{Sayers2010-vm,McEntyre2001-fl}).
-We used these articles to seed a machine-learning approach for identifying further candidate articles and to identify new search terms to include.
+Literature mining: We identified an initial list of 10 benchmark
+articles that satisfy the ALB-criteria. These were identified based
+upon previous knowledge of published articles and were supplemented
+with several literature searches (e.g. ‘“benchmark” AND “cputime”’ was
+used to query both GoogleScholar and Pubmed
+\cite{Sayers2010-vm,McEntyre2001-fl}). We used these articles to seed
+a machine-learning approach for identifying further candidate articles
+and to identify new search terms to include.
 
-For our machine-learning-based literature screening, we computed a score ($s(a)$) for each article that tells us the likelihood that it is a benchmark. In brief, our approaches uses 3 stages:
+For our machine-learning-based literature screening, we computed a
+score ($s(a)$) for each article that tells us the likelihood that it
+is a benchmark. In brief, our approaches uses 3 stages:
 \begin{enumerate}
@@
 \end{enumerate}
-For stage 1, we identified a list of high frequency (e.g. f(word) $>$ 1/10,000) words by pooling the content of two control texts \cite{Carroll1865-hk,Tolkien1937-ke}.
+For stage 1, we identified a list of high frequency (e.g. f(word) $>$
+1/10,000) words by pooling the content of two control texts
+\cite{Carroll1865-hk,Tolkien1937-ke}.
 
-For stage 2, in order to compute a log-odds score for bioinformatic words, we computed the frequency of words that were not removed by our high frequency filter in two different groups of articles: bioinformatics-background and bioinformatics-benchmark articles.
-The text from bioinformatics-background articles were drawn from the bioinformatics literature, but these were not necessarily associated with benchmark studies.
-For background text we used Pubmed (\cite{Sayers2010-vm,McEntyre2001-fl} to select 8,908 articles that contained the word “bioinformatics” in the title or abstract and were published between 2013 and 2015.
-We computed frequencies for each word by combining text from titles and abstracts for the background and training articles.
-A log-odds score is computed for each word using the following formula: $lo(w)=\log_2\frac{•f_{tr}(word)+\delta}{f_{bg}(word)+\delta}$, where  is a prior probability ($\delta = 10^{-5}$, by default), $f_{bg}(word)$ and $f_{tr}(word)$ are the frequencies of a $word$ in the background and training datasets respectively.
-Word frequencies are computed by counting the number of times a word appears in the pool of titles and abstracts, the counts are normalised by the total number of words in each set.
+For stage 2, in order to compute a log-odds score for bioinformatic
+words, we computed the frequency of words that were not removed by our
+high frequency filter in two different groups of articles:
+bioinformatics-background and bioinformatics-benchmark articles. The
+text from bioinformatics-background articles were drawn from the
+bioinformatics literature, but these were not necessarily associated
+with benchmark studies. For background text we used Pubmed
+(\cite{Sayers2010-vm,McEntyre2001-fl} to select 8,908 articles that
+contained the word “bioinformatics” in the title or abstract and were
+published between 2013 and 2015. We computed frequencies for each word
+by combining text from titles and abstracts for the background and
+training articles. A log-odds score is computed for each word using
+the following formula:
+$lo(w)=\log_2\frac{•f_{tr}(word)+\delta}{f_{bg}(word)+\delta}$, where
+is a prior probability ($\delta = 10^{-5}$, by default),
+$f_{bg}(word)$ and $f_{tr}(word)$ are the frequencies of a $word$ in
+the background and training datasets respectively. Word frequencies
+are computed by counting the number of times a word appears in the
+pool of titles and abstracts, the counts are normalised by the total
+number of words in each set.
 
-Thirdly, we also collected a group of candidate benchmark articles by mining Pubmed for articles that are likely to be benchmarks of bioinformatic software, these may match the terms: “((bioinformatics) AND (algorithms OR programs OR software)) AND (accuracy OR assessment OR benchmark OR comparison OR performance) AND (speed OR time)”.
-Further terms used in this search were progressively added as relevant enriched terms were identified in later iterations.
-The final query is given in \textbf{supplementary materials}.
+Thirdly, we also collected a group of candidate benchmark articles by
+mining Pubmed for articles that are likely to be benchmarks of
+bioinformatic software, these may match the terms: “((bioinformatics)
+AND (algorithms OR programs OR software)) AND (accuracy OR assessment
+OR benchmark OR comparison OR performance) AND (speed OR
+time)”. Further terms used in this search were progressively added as
+relevant enriched terms were identified in later iterations. The final
+query is given in \textbf{supplementary materials}.
 
-A score is computed for each candidate article by summing the log-odds scores for the words in title and abstract, i.e.
-$s(a)=\sum_i^Nlo(w_i)$.
-The high scoring candidate articles are then manually evaluated against the ALB-criteria.
-Accuracy and speed ranks are extracted from the articles that meet the criteria, and these are also added to the set of training articles.
-The evaluated candidate articles that do not meet the ALB-criteria are incorporated into the set of background articles.
-This process is iterated a number of times and has resulted in the identification of \textbf{{\color{black}43}} benchmark articles, that contain \textbf{{\color{black}102}} different benchmarks, together these rank \textbf{{\color{black}243}} distinct software packages.
+A score is computed for each candidate article by summing the log-odds
+scores for the words in title and abstract,
+i.e. $s(a)=\sum_i^Nlo(w_i)$. The high scoring candidate articles are
+then manually evaluated against the ALB-criteria. Accuracy and speed
+ranks are extracted from the articles that meet the criteria, and
+these are also added to the set of training articles. The evaluated
+candidate articles that do not meet the ALB-criteria are incorporated
+into the set of background articles. This process is iterated a number
+of times and has resulted in the identification of
+\textbf{{\color{red}\numBenchmarkPubs}} benchmark articles, that
+contain \textbf{{\color{red}\numBenchmarks}} different benchmarks, together these
+rank \textbf{{\color{red}\numTools~}} distinct software packages.
 
-There is a potential for bias to have been introduced into this dataset.
-Some possible forms of bias include converging on a niche group of benchmark studies due to the literature mining technique that we have used.
-A further possibility is that benchmark studies themselves are biased, either including very high performing or very low performing software tools.
-To address each of these concerns we have attempted to be as comprehensive as possible in terms of benchmark inclusion, as well as include comprehensive benchmarks.
-By which we mean studies that include all available software tools that address a biological problem.
+There is a potential for bias to have been introduced into this
+dataset. Some possible forms of bias include converging on a niche
+group of benchmark studies due to the literature mining technique that
+we have used. A further possibility is that benchmark studies
+themselves are biased, either including very high performing or very
+low performing software tools. To address each of these concerns we
+have attempted to be as comprehensive as possible in terms of
+benchmark inclusion, as well as include comprehensive benchmarks. By
+which we mean studies that include all available software tools that
+address a biological problem.
 
-\textbf{Data extraction and processing:} for each article that met the ALB-criteria and contained data on both the accuracy and speed from their tests we extracted ranks for each method.
-Many articles contained multiple benchmarks, in these cases we selected a range of these, the provenance of which is stored with the accuracy metric and raw speed and accuracy rank data for each method.
-In line with rank-based statistics, the cases where methods were tied are resolved by using a midpoint rank (e.g.
-if method 3 and 4 are tied, the rank 3.5 is used) \cite{Mann1947-re}.
-Each rank extraction was independently verified by at least one other co-author to ensure both the provenance of the data could be established and that the ranks were correct.
-The ranks for each benchmark were then normalised to lie between 0 and 1 using the formula $1-\frac{r-1}{n-1}$ where ‘r’ is a method’s rank and ‘n’ is the number of methods in the benchmark.
-For methods that were benchmarked multiple times with multiple metrics (e.g.
-BWA is evaluated in 6 different articles \cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd}) a mean normalised rank is used to summarise the performance.
+\textbf{Data extraction and processing:} for each article that met the
+ALB-criteria and contained data on both the accuracy and speed from
+their tests we extracted ranks for each tool. Many articles contained
+multiple benchmarks, in these cases we selected a range of these, the
+provenance of which is stored with the accuracy metric and raw speed
+and accuracy rank data for each tool. In line with rank-based
+statistics, the cases where tools were tied are resolved by using a
+midpoint rank (e.g. if tool 3 and 4 are tied, the rank 3.5 is used)
+\cite{Mann1947-re}. Each rank extraction was independently verified by
+at least one other co-author to ensure both the provenance of the data
+could be established and that the ranks were correct. The ranks for
+each benchmark were then normalised to lie between 0 and 1 using the
+formula $1-\frac{r-1}{n-1}$ where ‘r’ is a tool’s rank and ‘n’ is the
+number of tools in the benchmark. For tools that were benchmarked
+multiple times with multiple metrics (e.g. BWA is evaluated in 6
+different articles
+\cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
+a mean normalised rank is used to summarise the performance.
 
-For each method we identified the corresponding publications in GoogleScholar, the total number of citations was recorded, the corresponding authors were also identified and if these had public GoogleScholar profiles we extracted their H-index and calculated a M-index ($\frac{H-index}{y}$) where ‘$y$’ is the number of years since their first publication.
-For the journals that each method is published in we extracted the “journal impact factor” (JIF) and the H5-index from Thompson-Reuters and GoogleScholar Metrics databases respectively.
-The year of publication was also recorded for each method.
-A “relative age” and “relative citations” was also computed for each method.
-For each benchmark, software was ranked by year of first publication (or number of citations), ranks were assigned and then normalised as described above.
-Methods ranked in multiple evaluations were then assigned a mean value for “relative age” and “relative citations”.
+For each tool we identified the corresponding publications in
+GoogleScholar, the total number of citations was recorded, the
+corresponding authors were also identified and if these had public
+GoogleScholar profiles we extracted their H-index and calculated a
+M-index ($\frac{H-index}{y}$) where ‘$y$’ is the number of years since
+their first publication. For the journals that each tool is published
+in we extracted the “journal impact factor” (JIF) and the H5-index
+from Thompson-Reuters and GoogleScholar Metrics databases
+respectively. The year of publication was also recorded for each
+tool. A “relative age” and “relative citations” was also computed for
+each tool. For each benchmark, software was ranked by year of first
+publication (or number of citations), ranks were assigned and then
+normalised as described above. Tools ranked in multiple evaluations
+were then assigned a mean value for “relative age” and “relative
+citations”.
 
-\textbf{Statistical analysis:} For each method we have up to 10 statistics (1.
-corresponding author’s H-index, 2.
-corresponding author’s M-index, 3.
-journal H5 index, 4.
-journal impact factor, 5.
-normalised accuracy rank, 6.
-normalised speed rank, 7.
-number of citations, 8.
-relative age, 9.
-relative number of citations, 10.
-year first published).
-These have been evaluated in a pairwise fashion to produce Figure~\ref{fig:allfactors} A\&B, the R code for these is given in the supplement.
+\textbf{Statistical analysis:} For each tool we have up to 10
+statistics (1. corresponding author’s H-index, 2. corresponding
+author’s M-index, 3. journal H5 index, 4. journal impact factor,
+5. normalised accuracy rank, 6. normalised speed rank, 7. number of
+citations, 8. relative age, 9. relative number of citations, 10. year
+first published). These have been evaluated in a pairwise fashion to
+produce Figure~\ref{fig:allfactors} A\&B, the R code for these is
+given in the supplement.
 
-The linear models that we used to test for relationships between speed, accuracy and the above measures are:
+The linear models that we used to test for relationships between
+speed, accuracy and the above measures are:
 
@@
 
-For each benchmark of three or more methods, we extracted the published accuracy and speed ranks.
-In order to identify if there is an enrichment of certain accuracy and speed pairings we constructed a permutation test.
-The individual accuracy and speed ranks were reassigned to methods in a random fashion and each new accuracy and speed rank pairing was recorded.
-For each benchmark this procedure was repeated 10,000 times.
-These permuted rankings were normalised and compared to the real rankings to produce the ‘x’ points in Figure~\ref{fig:allfactors}B and the heatmap and histograms in Figure~\ref{fig:speedaccuracy}.
-The heatmap in Figure~\ref{fig:speedaccuracy} is based upon Z-scores ($Z=\frac{x-\bar{x}}{\sigma}$).
-For each cell in a $10\times 10$ grid a Z-score is computed to illustrate the abundance or lack of methods in a cell relative to the permuted data.
+For each benchmark of three or more tools, we extracted the published
+accuracy and speed ranks. In order to identify if there is an
+enrichment of certain accuracy and speed pairings we constructed a
+permutation test. The individual accuracy and speed ranks were
+reassigned to tools in a random fashion and each new accuracy and
+speed rank pairing was recorded. For each benchmark this procedure was
+repeated 10,000 times. These permuted rankings were normalised and
+compared to the real rankings to produce the ‘x’ points in
+Figure~\ref{fig:allfactors}B and the heatmap and histograms in
+Figure~\ref{fig:speedaccuracy}. The heatmap in
+Figure~\ref{fig:speedaccuracy} is based upon Z-scores
+($Z=\frac{x-\bar{x}}{\sigma}$). For each cell in a $10\times 10$ grid
+a Z-score is computed to illustrate the abundance or lack of tools in
+a cell relative to the permuted data.
 
@@
 discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia
-Palmer and Jason Tylianakis. Michael A. Black, Murray Cox, Raquel
+Palmer and Jason Tylianakis. Murray Cox, Raquel
 Norel, Alexandros Stamatakis, Jens Stoye,Tandy Warnow, provided
@@
 administered by the Royal Society of New Zealand. 
+{\bf AND ETC....}
 

### 2021-02-24 manuscript/supplementary.tex
Updating MS.
Major update: larger dataset, revised and simplified code, new figures.

@@
 
+%\Authors{Paul P.
+Gardner\textsuperscript{1,2}*, James Paterson\textsuperscript{1,2}, Fatemeh Ashari Ghomi\textsuperscript{1,2}, Sinan Uğur Umu\textsuperscript{1,2}, Stephanie McGimpsey\textsuperscript{1,2}} % Authors
+%\affiliation{\textsuperscript{1}\textit{School of Biological Sciences, University of Canterbury, Christchurch, New Zealand.}} % Author affiliation
+%\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre and the Bio-Protection Research Centre, University of Canterbury, Christchurch, New Zealand.}}
+%\affiliation{*\textbf{Corresponding author}: paul.gardner@canterbury.ac.nz} % Corresponding author
+
 \Authors{Paul P.
-Gardner\textsuperscript{1,2}*, James Paterson\textsuperscript{1,2}, Fatemeh Ashari Ghomi\textsuperscript{1,2}, Sinan Uğur Umu\textsuperscript{1,2}, Stephanie McGimpsey\textsuperscript{1,2}} % Authors
-\affiliation{\textsuperscript{1}\textit{School of Biological Sciences, University of Canterbury, Christchurch, New Zealand.}} % Author affiliation
-\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre and the Bio-Protection Research Centre, University of Canterbury, Christchurch, New Zealand.}}
-\affiliation{*\textbf{Corresponding author}: paul.gardner@canterbury.ac.nz} % Corresponding author
+Gardner\textsuperscript{1,2}*, James M.
+Paterson\textsuperscript{?}, Fatemeh Ashari-Ghomi\textsuperscript{?}, Sinan U.
+Umu\textsuperscript{?}, Stephanie McGimpsey\textsuperscript{3,4}, Aleksandra Pawlik\textsuperscript{5},
+Alex Gavryushkin\textsuperscript{6}, Michael A Black\textsuperscript{1}
+} % Authors
+
+\affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
+\affiliation{\textsuperscript{3}\textit{McConomy School of Dance, Derry, Ireland.}}
+\affiliation{\textsuperscript{4}\textit{Research for Good, Limavady, Ireland.}}
+\affiliation{\textsuperscript{5}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
+\affiliation{\textsuperscript{6}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+\affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author
 
@@
 \centering
-\includegraphics[width=0.75\textwidth]{supplementary-figures-small.pdf}
+\includegraphics[width=0.95\textwidth]{supplementary-figures-small.pdf}
 \caption{The distributions for the measures we expected to be
@@
 normalised accuracy rank for each benchmarked method; Speed -- the
-mean normalised speed rank for each benchmarked method; Cites -- the number of
-citations to the most cited manuscript describing a method, data from GoogleScholar;
-JIF -- the Journal Impact Factor to the highest impact journal that has published
-a manuscript describing a method, data from 2014 Thompson-Reuters Citation Reports;
+mean normalised speed rank for each benchmarked method;
 JH5 -- the Journal H5 index to the highest impact journal that has published
-a manuscript describing a method, data from GoogleScholar 2015 Metrics;
+a manuscript describing a method, data from GoogleScholar 2020 Metrics;
+Cites -- the number of citations to the most cited manuscript describing a method, data from GoogleScholar;
 H-index -- the H-index for the highest profile corresponding author from the
@@
 highest profile corresponding author from the manuscripts describing a method,
-data from GoogleScholar User Profiles;}
+data from GoogleScholar User Profiles;
+Relative ages for different tools, for each benchmark tools were ranked based upon publication dates, these ranks were normalised to lie between 0 and 1.
+}
 \label{fig:metricDistributions}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.95\textwidth]{supplementary-figures-pairs.pdf}
+\caption{}
+\label{fig:metricPairs}
 \end{figure}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.5\textwidth]{smoothScatter-speed-vs-accuracy.pdf}
-\caption{A smoothed color density representation of a scatterplot of
-  normalised speed ranks and normalised accuracy ranks. Dark red
-  regions are indicative of of a high density of points, blue shades
-  indicate the reverse. The expected inverse relationship between
-  speed and accuracy is indicated with a dashed black line, points
-  above this line could be considered comparatively fast and accurate,
-  points below are the reverse. A locally weighted regression (lowess)
-  curve is shown in red.}
-\label{fig:smoothScatterAccSpeed}
-\end{figure}
+%% \begin{figure}[H]
+%% \centering
+%% \includegraphics[width=0.5\textwidth]{smoothScatter-speed-vs-accuracy.pdf}
+%% \caption{A smoothed color density representation of a scatterplot of
+%%   normalised speed ranks and normalised accuracy ranks. Dark red
+%%   regions are indicative of of a high density of points, blue shades
+%%   indicate the reverse. The expected inverse relationship between
+%%   speed and accuracy is indicated with a dashed black line, points
+%%   above this line could be considered comparatively fast and accurate,
+%%   points below are the reverse. A locally weighted regression (lowess)
+%%   curve is shown in red.}
+%% \label{fig:smoothScatterAccSpeed}
+%% \end{figure}
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.85\textwidth]{smoothScatters.pdf}
-\caption{A smoothed color density representation of a scatterplot of
-  journal impact factors (JIF) on the left and number of citations on
-  the right vs normalised accuracy ranks. As above, dark red regions
-  are indicative of of a high density of points, blue shades indicate
-  the reverse. The JIF for journals that are major publishers of
-  bioinformatic software are indicated on the left. }
-\label{fig:smoothScatters}
-\end{figure}
+%% \begin{figure}[H]
+%% \centering
+%% \includegraphics[width=0.85\textwidth]{smoothScatters.pdf}
+%% \caption{A smoothed color density representation of a scatterplot of
+%%   journal impact factors (JIF) on the left and number of citations on
+%%   the right vs normalised accuracy ranks. As above, dark red regions
+%%   are indicative of of a high density of points, blue shades indicate
+%%   the reverse. The JIF for journals that are major publishers of
+%%   bioinformatic software are indicated on the left. }
+%% \label{fig:smoothScatters}
+%% \end{figure}
 
@@
 \centering
-\includegraphics[width=0.95\textwidth]{relAge-SpeedVsAccuracy-heatmap-edited.pdf}
-\caption{{\bf A.} Violin plots for the relative age distribution for software tools in each of the 5 2x2 cells indicated in B.
-The five boxes correspond to the four extreme corners of the speed vs accuracy spectrum (i.e.
-slow and inaccurate, slow and accurate, fast and inaccurate, fast and accurate) and the central box (medium speed and medium accuracy).
-{\bf B.} A heatmap indicating the relative age of software in a range of relative accuracy and speed rankings.
-Blue colours indicate an abundance of older software tools in an accuracy and speed category, while light colours indicate younger software in an accuracy and speed category.
-}
+%\includegraphics[width=0.95\textwidth]{relAge-SpeedVsAccuracy-heatmap-edited.pdf}
+\includegraphics[width=0.45\textwidth]{relAge-speedAcc.pdf}
+\includegraphics[width=0.45\textwidth]{relAge-SpeedVsAccuracy-heatmap.pdf}
+\caption{{\bf Left.} Violin plots for the relative age distribution
+  for software tools in each of the 5 2x2 cells indicated in B. The
+  five boxes correspond to the four extreme corners of the speed vs
+  accuracy spectrum (i.e. slow and inaccurate, slow and accurate, fast
+  and inaccurate, fast and accurate) and the central box (medium speed
+  and medium accuracy). {\bf Right.} A heatmap indicating the relative age
+  of software in a range of relative accuracy and speed rankings. Blue
+  colours indicate an abundance of older software tools in an accuracy
+  and speed category, while light colours indicate younger software in
+  an accuracy and speed category. }
 \label{fig:ageplot}

### 2021-02-24 bin/prettyPlot.R

@@ -3,11 +3,6 @@
 #Rscript ../bin/prettyPlot.R
 #R CMD BATCH ../bin/prettyPlot.R
 
-#check which of these are still in use:
-#if (!requireNamespace("BiocManager", quietly = TRUE))
-                                        #    install.packages("BiocManager")
-#BiocManager::install("multtest")
-#install.packages(c("MASS", "RColorBrewer", "fields", "gplots", "gplots", "hash", "vioplot"))
 
 library(MASS)  # in case it is not already loaded 
 library(RColorBrewer)
@@ -15,111 +10,85 @@ library(fields)
 library(gplots)
 #library(metap)
 #library("CombinePValue")
-library(gplots)
 library("hash")
 library("vioplot")
 
 k <- 11
 my.cols <- rev(brewer.pal(k, "RdYlBu"))
-d <-read.table("meanRankSpeedData.tsv", header=T)
-dr<-read.table( "rawRankSpeedData.tsv", header=T)
+d2005 <-read.table("meanRankSpeedData.tsv",      header=T)
+dr2005<-read.table( "rawRankSpeedData2005-2020.tsv", header=T)
 
+#d2015 <-read.table("meanRankSpeedData2015.tsv",      header=T)
+#dr2015<-read.table( "rawRankSpeedData2016-2020.tsv", header=T)
 
 
 
 ######################################################################
 #TESTING IF SPEED inv. ACCURACY IS TRUE:
-reg1  <- lm(d$accuracyRank ~ d$speedRank)
-#reg1c <- coefficients(reg1)
+####Regression:
+reg1  <- lm(d2005$accuracyRank ~ d2005$speedRank)
 summary(reg1)
 
-regM <- lm(accuracyRank ~ speedRank + IF + H5 + cites + hindex + mindex + relAge + relCites, data=d)
+regM <- lm(accuracyRank ~ speedRank + H5 + citations + hindex + mindex + relAge + relCites + version + commits + contributors, data=d2005, na.action=na.roughfix)
 summary(regM)
 
-
-xx<-cbind(d$accuracyRank, d$speedRank, d$IF, d$H5, d$cites, d$hindex, d$mindex, d$relAge, d$relCites,2016-d$yearPublished+1)
-colnames(xx)<-c("accuracy","speed","JIF","JH5","cites","hindex","mindex","relAge","relCites","age")
+####PCA:
[…]

### 2021-02-24 bin/tsv2data.pl

@@ -2,98 +2,208 @@
 
 use warnings;
 use strict;
+use Data::Dumper;
 
-my %methodInfo; 
-my @methodInfoKeys = qw(yearPublished IF H5 cites hindex mindex); 
-# head -n 1 speed-vs-accuracy-toolInfo2005-2015.tsv | tr "\t" "\n" | nl
-#      1	method
-#      2	yearPublished
-#      3	journal
-#      4	impactFactor
-#      5	Journal H5-index
-#      6	totalCitations
-#      7	Corresponding author: H-index
-#      8	Corresponding author: M-index
-#      9	fullCite
-#Read tool info:
-open(IN0, "< speed-vs-accuracy-toolInfo2005-2015.tsv");  #Does\ bioinformatic\ software\ trade\ speed\ for\ accuracy-\ -\ methodInfo.tsv
-while(my $in = <IN0>){
-    next if $in =~ /^tool/; 
-    $in =~ s/[^[:ascii:]]//g; #strip off unicode chars
-    chomp($in);
-    $in =~  s/\r//g;
-    my @in = split(/\t/, $in); 
+###################################
+#Parse tool info (e.g. age, impact, etc.):
+my @methodInfoKeys = qw(yearPublished H5 citations hindex mindex version commits contributors); #IF 
+my $methodInfoPtr = extractMethodInfo("speed-vs-accuracy-toolInfo2005-2020.tsv", "journalInfo2020.tsv"); #, \%methodInfo); 
+my %methodInfo = %{$methodInfoPtr};
     
-    if(defined($in[0]) && length($in[0]) > 1){
-	
-	if(defined($in[1])){
-	    my @yrs = split(/;\s*/, $in[1]);
-	    $methodInfo{$in[0]}{'yearPublished'}='NA';
-	    $methodInfo{$in[0]}{'yearPublished'}=minA(@yrs) if ($in[1] !~ 'NA'); #using the first publication
-	    #printf "[$in[0]]\tyrsArray[@yrs]\tyearStr[$in[1]]\tminA[%d]\n", minA(@yrs);
-	}
-	
-	if(defined($in[3])){
-	    my @ifs = split(/;\s*/, $in[3]);	    
-	    $methodInfo{$in[0]}{'IF'}='NA';
-	    $methodInfo{$in[0]}{'IF'}=maxA(@ifs) if ($in[3] !~ 'NA'); #using the highest impact factor
-	}
[…]

### 2021-02-24 manuscript/manuscript-speed-accuracy.tex
Updated README.md, manuscript and figures.

@@
 \usepackage{amsmath}
+\usepackage{stmaryrd}
+\usepackage{todonotes}
+\usepackage{maori}
 \graphicspath{{../figures/}}
@@
 Gardner\textsuperscript{1,2}*, James M.
-Paterson\textsuperscript{?}, Fatemeh Ashari-Ghomi\textsuperscript{?}, Sinan U.
-Umu\textsuperscript{?}, Stephanie McGimpsey\textsuperscript{3,4}, Aleksandra Pawlik\textsuperscript{5},
-Alex Gavryushkin\textsuperscript{6}, Michael A Black\textsuperscript{1}
+Paterson\textsuperscript{3}, Stephanie McGimpsey\textsuperscript{4}, Fatemeh Ashari-Ghomi\textsuperscript{5}, Sinan U.
+Umu\textsuperscript{6}, Aleksandra Pawlik\textsuperscript{7},
+Alex Gavryushkin\textsuperscript{8}, Michael A Black\textsuperscript{1}
 } % Authors
@@
 \affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
-\affiliation{\textsuperscript{3}\textit{McConomy School of Dance, Derry, Ireland.}}
-\affiliation{\textsuperscript{4}\textit{Research for Good, Limavady, Ireland.}}
-\affiliation{\textsuperscript{5}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
-\affiliation{\textsuperscript{6}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+\affiliation{\textsuperscript{3}\textit{Department of Civil and Natural Resources Engineering, University of Canterbury, Christchurch, New Zealand.}}
+\affiliation{\textsuperscript{4}\textit{Parasites and Microbes, Wellcome Sanger Institute, Wellcome Genome Campus, Hinxton,
+8 Cambridgeshire, CB10 1RQ, UK.}}
+\affiliation{\textsuperscript{5}\textit{Research Group for Genomic Epidemiology, National Food Institute, Technical University of Denmark, Kongens Lyngby, Denmark.}}
+\affiliation{\textsuperscript{6}\textit{Department of Research, Cancer Registry of Norway, Oslo, Norway.}} % Author affiliation
+\affiliation{\textsuperscript{7}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
+\affiliation{\textsuperscript{8}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+
 \affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author
@@
   data. In the face of rapidly increasing volumes of data, heuristic
-  methods that trade software speed for mathematical completeness must
+  methods that trade software speed for accuracy must
   be employed.  We are interested in testing for trade-offs between
-  speed and accuracy and for other factors are indicative of accurate
+  speed and accuracy and for other factors that are indicative of accurate
   software.
@@
 We have extracted accuracy and speed ranks from independent benchmarks of
-computational biology software of different software tools, and evaluate the
+computational biology software of different software tools, and evaluated the
 factors that are likely to influence accuracy e.g. speed, author
@@
 tools. Our results do show that accurate bioinformatic software tools
-are generally the product many gradual improvements, often from
-multiple developers, based upon github records. In addition, we find
-that bioinformatics has an excess of slow and inaccurate software
-tools across many sub-disciplines. Meanwhile, there are few tools that
+are generally the product of many gradual improvements, often from
+multiple developers, based upon matching GitHub records with benchmark results. 
+In addition, we find
+that the field of bioinformatics has an excess of slow and inaccurate software
+tools, this is consistent across many sub-disciplines. Meanwhile, there are few tools that
 are middle-of-road in terms of accuracy and speed trade-offs. We
 hypothesise that a form of publication-bias unduly influences the
-publication and development of bioinformatic software tools. In other
-words, at present software that is not highly ranked on speed and not
-highly ranked on accuracy is difficult to publish due to author,
+publication and development of bioinformatic software. In other
+words, software that is intermediate in terms of both speed and accuracy may be difficult to publish due to author,
 editor and reviewer practices. This leaves an unfortunate gap in the
-literature upon which future software refinements cannot be
-constructed.  }
+literature as the ideal tools may fall in the gap e.g. high accuracy is generally not useful if years of CPU time is required, 
+while high speed is not useful if the results are very inaccurate.  }
 
@@
 operating system and other factors may play a greater role in a software
-tool's popularity. There have been several notable incidences where
+tool's popularity. Indeed, there have been several notable incidences where
 inaccurate software has caused considerable harm
-\cite{leveson1993investigation}.
+\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
 %Fergusson covid predictions?
@@
 Progress in the biological sciences is increasingly limited by the
-ability to analyse increasing volumes of data, therefore the
+ability to analyse large volumes of data, therefore the
 dependence of biologists on software is also increasing
-\cite{Marx2013-zi}. There is an increasing use of technological
+\cite{Marx2013-zi}. There is an increasing reliance on technological
 solutions for automating biological data generation
@@
 computational efficiency of data processing and analysis software is
-of great importance to decrease the energy and time costs of research
-\cite{Gombiner2011-md}. Furthermore, even small error rates can have a
-major impact on the number of false inferences as datasets become
+of great importance to decrease the energy, climate impact, and time costs of research
+\cite{Gombiner2011-md}. Furthermore, even small error rates can have
+major impacts on the number of false inferences as datasets become
 larger \cite{Storey2003-cv}.
@@
 diversity of options from which practical solutions can be selected,
-many possible options creates a dilemma for users. In the absence of
+having many possible options creates a dilemma for users. In the absence of
 any recent gold-standard benchmarks, how should scientific software be
@@
 accuracy is the degree to which predictions or measurements reflect
-the biological truths based on expert-derived curated datasets. In some
+the biological truths based on expert-derived curated datasets (see Methods for the  mathematical definition used here). In some
 fields biological accuracy is very difficult to ascertain, for
-example, in phylogenetics it is nearly impossible to know the
+example, in phylogenetics it is nearly impossible in most cases to know the
 true ancestral relationships between organisms. In situations like this,
@@
 A number of possible predictors of software quality are used by the
-community of computational biology software users. Some accessible,
+community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. Some accessible,
 quantifiable and frequently used proxies for identifying high quality
 software include: \textbf{1. Recency:} recently published software
-tools may have built upon the results of past work or be an update to
-an existing software. Therefore, these could be more accurate and
+tools may have built upon the results of past work, or be an update to
+an existing software. Therefore these may be more accurate and
 faster. \textbf{2. Wide adoption:} a software tool may be widely used
@@
 frequently trade accuracy for speed. For example, heuristic
-software such as the popular homology search tool, BLAST, compromise
+software such as the popular homology search tool, BLAST, compromises
 the mathematical guarantee of optimal solutions for more speed
@@
 \cite{Fourment2008-vl}, and the level of hardware optimisation
-\cite{Farrar2007-ky,Dematte2010-ph}; however, the implementation is
-likely to have more of an impact (e.g. brute-force approaches versus
+\cite{Farrar2007-ky,Dematte2010-ph};
+however, the implementation generally has more of an impact (e.g. brute-force approaches versus
 rapid and sensitive pre-filtering
 \cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).
+%\todo{This paper
+%https://science.sciencemag.org/content/368/6495/eaam9744.full
+%shows a thousands-fold increase in speed when parallelisation
+%is implemented with things like cache, architecture, etc.
+%mind VS naive parallelisation as in Python, R, etc.}
 
@@
 user-friendly, word-of-mouth and ``used in a similar analysis''
-\cite{Loman2015-bw}.
-%This information is not as readily
-%quantifiable as the above measures. However,
-However, citation metrics may be a
+\cite{Loman2015-bw}. However, citation metrics may be a
 useful proxy for the above.
@@
 
-With the wide adoption of Github ($\approx 47\%$ of the
+With the wide adoption of GitHub ($\approx 47\%$ of the
 \textbf{{\color{red}\numTools}} tools included in this study could be
-linked to a Github repository), quantifiable data on software
+linked to a GitHub repository), quantifiable data on software
 development such as the number of contributors to code, number of code
@@
 each tool, we have collected measures that may be predictive
-of accuracy, and may be subjectively employed by the researcher
+of accuracy, and may be subjectively employed by the research
 community as a proxy for software quality. These include relative
@@
 \textbf{{\color{red}\numTools~}} distinct software tools. This software has
-been developed for solving a broad cross-section computational biology
+been developed for solving broad cross-section computational biology
 tasks.
-%% These include software for homology search
-%% \cite{Freyhult2007-et}, genome sequence analysis (e.g. read mapping or
-%% sequence assembly)
-%% \cite{Junemann2014-mb,Tran2014-pe,Zhang2011-nd,Abbas2014-gu,Bao2011-lv,Caboche2014-lj,Kleftogiannis2013-wi,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Yang2013-aj,Holtgrewe2011-fd,Rackham2015-ag,Huang2015-wu},
-%% multiple sequence alignment
-%% \cite{Thompson2011-rf,Nuin2006-nk,Pais2014-sr,Pervez2014-zp,Liu2010-rp},
-%% cell tracking \cite{Maska2014-ak}, transcriptome analysis
-%% \cite{Li2012-wr,Lu2013-fs,Liu2014-kz,Kumar2016-xz}, RNA interaction
-%% prediction \cite{Pain2015-gr}, protein interactions
-%% \cite{Tikk2010-qd}, protein structure prediction
-%% \cite{Kolodny2005-ry,Wallner2005-qi}, epistasis \cite{Shang2011-vy},
-%% metagenomic analysis \cite{Lindgreen2016-tt,Bazinet2012-wf},
-%% repetitive sequence prediction \cite{Saha2008-kd}, proteomics
-%% \cite{Lange2008-pt,Yang2009-oc} and phylogenetics
-%% \cite{Liu2011-pz,Yang2011-dv,Oscamou2008-md,Bayzid2013-hc,Liu2009-lx}.
-Each
-software tool was benchmarked in at least one of
+Each software tool was benchmarked in at least one of
 \textbf{{\color{red}\numBenchmarkPubs~}} publications that satisfy the Boulesteix
@@
 For each of the publications describing these tools, we have (where
-possible) the H5-index published by
-\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
-  Scholar Metrics}. We have collected the maximum H-index and
+possible) collected the journal's H5-index (\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
+  Scholar Metrics}), the maximum H-index and
 corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
@@
 associated with a tool has been cited using Google Scholar (data
-collected over a 6 month period in late 2020). Note that the citation metrics
-are not static and will change over time.
+collected over a 6 month period in late 2020). Note that citation metrics
+are not static and will change over time. In addition, where possible we also extract the version number, the number
+of commits and number of contributors from GitHub repositories.
 
@@
 pairwise combination of the mean normalised accuracy and speed ranks,
-the year published, mean relative age (compared to software in the
+with the year published, mean relative age (compared to software in the
 same benchmarks), journal H5 metrics, the total number of citations,
@@
 benchmarks) and the maximum H- and corresponding M-indices for the corresponding
-authors. Where possible we also extract the version number, the number
-of commits and number of contributors from Github repositories.  The
-results are presented in Figure~\ref{fig:allfactors}A. We find
+authors, version number, and numbers of commits and contributors.
+The results are presented in Figure~\ref{fig:allfactors}A. We find
 significant associations between most of the citation-based metrics
@@
 
-\begin{figure*}[h]
+\begin{figure*}[ht]
 \includegraphics[width=\textwidth]{figure1.pdf}
@@
   indicate a positive correlation, blue colours indicate a negative
-  correlation. Correlations with a P-value less than 0.05 are
-  indicated with a ‘X’. The dashed rectangular area is illustrated in
+  correlation. Correlations with an uncorrected P-value less than 0.05 are
+  indicated with a `$\vartimes$'. The dashed rectangular area is illustrated in
   more detail in \textbf{B}, the relationship between speed and
@@
   for each benchmark and recorded Spearman’s rho for the significant
-  correlations (P<0.05). These values are marked with “x”s in the
+  correlations with an uncorrected P-value less than 0.05.
+  These values are marked with blues `$\vartimes$'s in the
   barplot. Significant correlations are marked with red asterisks.}
@@
 Data on the
-number of updates to software tools from Github such as the number of versions,
+number of updates to software tools from GitHub such as the number of versions,
 commits and contributors \textbf{{\color{red}was significantly correlated with 
-accuracy (respective Spearman’s rhos = 0.13, 0.22, 0.21, respective P-values = 0.0027, 0.00079, 0.0013). 
+accuracy (respective Spearman’s rhos = 0.13, 0.22, 0.21, respective uncorrected P-values = 0.0027, 0.00079, 0.0013). 
 These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A).
@@
 expected distributions generated by 1,000 randomized ranks. The results
-of which are shown in Figure~\ref{fig:speedaccuracy}. We identified \textbf{{\color{red}18}}
+of this are shown in Figure~\ref{fig:speedaccuracy}. We identified \textbf{{\color{red}18}}
 bins where there was a significant excess or dearth of tools. For
@@
 P-value=0.0063}}). We find that the amount of software classed as “fast
-and accurate” and “fast and inaccurate” are at approximately the
-expected proportions based upon the permutation test. The number of
+and accurate”, “fast and inaccurate” and "slow and accurate" are at approximately the
+expected proportions based upon the permutation test. 
+While there are very few tools placed in the middle of the ranks on accuracy and speed.
+The number of
 significant results is in excess, and is not due to multiple testing, as
@@
 
-\begin{figure*}
+\begin{figure*}[ht]
 %\includegraphics[scale=0.55]{figure2.pdf}
-\includegraphics[width=\textwidth]{figure2.pdf}
+\centering
+\includegraphics[width=0.75\textwidth]{figure2.pdf}
 \caption{A heatmap indicating the relative paucity or abundance of
@@
   fast/accurate. Z-scores with a P-value less than 0.05 are indicated
-  with a ‘X’.}
+  with a ‘$\vartimes$’.}
 \label{fig:speedaccuracy}
@@
 finding of a strong relationship between the number of commits to
-github (i.e. software updates) and accuracy, highlights the benefits of
+GitHub (i.e. software updates) and accuracy, highlights the benefits of
 long-term development.
@@
 Our study finds little evidence to support that impact-based metrics
-are poorly related to software quality, this is unfortunate as these
+have any relationship with software quality, this is unfortunate as these
 are frequently cited reasons for selecting software tools
-\cite{Loman2015-bw}. This implies that the recorded high citation
+\cite{Loman2015-bw}. This implies that the high citation
 rates for bioinformatic software
 \cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
-reflection of user-friendliness and possibly the Matthew Effect
-\cite{Lariviere2010-kx,Merton1968-cb} than accuracy.
+reflection of other factors such as user-friendliness or the Matthew Effect
+\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.
 
@@
 probability that a study being published is influenced by the results
-it contains \cite{sterling1959publication}. The community of
+it contains \cite{sterling1995publication}. The community of
 developers, reviewers and editors may be unwilling to publish software
@@
 
-{\bf STUDY LIMITATIONS...}
+{\bf DESCRIBE STUDY LIMITATIONS... 
+\begin{itemize}
+    \item Difficult to test publication biases
+    \item Heterogeneous measures of accuracy between benchmarks  
+    \item possible lack of independence between different benchmarks 
+    \item Non-uniform inclusion of tools in different benchmarks -- some tools ranked many times (e.g. velvet and soap assemblers ranked 26 different times)
+    \item How to deal with inconsistently reported versioning of software? 
+    \item Metrics and tools performance are all moving targets, only a snap shot is used here  
+\end{itemize}
+}
 
@@
 The current
-practises of publishers, editors, reviewers and authors of software
+practices of publishers, editors, reviewers and authors of software
 tools therefore deprive our community of tools for building effective
@@
 \end{enumerate}
-For stage 1, we identified a list of high frequency (e.g. f(word) $>$
-1/10,000) words by pooling the content of two control texts
+For stage 1, we identified a list of high frequency (e.g. $f$(word) $>$
+$1/10,000$) words by pooling the content of two control texts
 \cite{Carroll1865-hk,Tolkien1937-ke}.
@@
 the following formula:
-$lo(w)=\log_2\frac{•f_{tr}(word)+\delta}{f_{bg}(word)+\delta}$, where
+%\todo{Word is denoted $w$ on the left and $word$ on the right.}
+$lo(word)=\log_2\frac{f_{tr}(word)+\delta}{f_{bg}(word)+\delta}$, where
+$\delta$
 is a prior probability ($\delta = 10^{-5}$, by default),
@@
 candidate articles that do not meet the ALB-criteria are incorporated
-into the set of background articles. This process is iterated a number
-of times and has resulted in the identification of
+into the set of background articles. This process is iterated and has resulted in the identification of
 \textbf{{\color{red}\numBenchmarkPubs}} benchmark articles, that
@@
 each benchmark were then normalised to lie between 0 and 1 using the
-formula $1-\frac{r-1}{n-1}$ where ‘r’ is a tool’s rank and ‘n’ is the
+formula $1-\frac{r-1}{n-1}$ where ‘$r$’ is a tool’s rank and ‘$n$’ is the
 number of tools in the benchmark. For tools that were benchmarked
@@
 \cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
-a mean normalised rank is used to summarise the performance.
+a mean normalised rank is used to summarise the accuracy and speed performance. 
+Or, more formally:
+
+\begin{equation*}
+\begin{split}
+accuracy =& \sum_{i=1..N} 1-\frac{r^{accuracy}_i-1}{n_i-1}, \\
+speed    =& \sum_{i=1..N} 1-\frac{r^{speed   }_i-1}{n_i-1}
+\end{split}
+\end{equation*}
 
@@
 their first publication. For the journals that each tool is published
-in we extracted the “journal impact factor” (JIF) and the H5-index
-from Thompson-Reuters and GoogleScholar Metrics databases
-respectively. The year of publication was also recorded for each
+in we extracted the H5-index
+from GoogleScholar Metrics. The year of publication was also recorded for each
 tool. A “relative age” and “relative citations” was also computed for
@@
 
-\textbf{Statistical analysis:} For each tool we have up to 10
-statistics (1. corresponding author’s H-index, 2. corresponding
-author’s M-index, 3. journal H5 index, 4. journal impact factor,
-5. normalised accuracy rank, 6. normalised speed rank, 7. number of
-citations, 8. relative age, 9. relative number of citations, 10. year
-first published). These have been evaluated in a pairwise fashion to
+The papers describing tools were checked for information on version numbers and links to GitHub. 
+Google was also employed to identify GitHub repositories.
+When a repository was matched with a tool, the number of "commits" and number of "contributors" was collected, when details of version numbers were provided, these were also harvested.
+Version numbers are inconsistently used between groups, and may begin at either 0 or 1.
+To counter this issue we have added '1' to all versions less than '1', e.g.
+version 0.31 become 1.31.
+In addition, multiple point releases may be used e.g.
+'version 5.2.6', these have been mapped to the nearest decimal value '5.26'.
+
+\textbf{Statistical analysis:} For each tool we have manually collected up to 12 different
+statistics from GoogleScholar, GitHub and directly from literature describing tools (
+1. corresponding author’s H-index, 
+2. corresponding author’s M-index, 
+3. journal H5 index, 
+4. normalised accuracy rank, 
+5. normalised speed rank, 
+6. number of citations, 
+7. relative age, 
+8. relative number of citations, 
+9. year first published, 
+10. version
+11. number of commits to GitHub,
+12. number of contributors to GitHub). 
+These have been evaluated in a pairwise fashion to
 produce Figure~\ref{fig:allfactors} A\&B, the R code for these is
-given in the supplement.
+given in the GitHub repository (linked below).
 
-The linear models that we used to test for relationships between
-speed, accuracy and the above measures are:
-
-\begin{equation*}
-\begin{split}
-accuracy=& c_0+c_1\times speed+c_2\times JIF+c_3\times H5+\\
-& c_4\times citations+c_5\times Hindex+\\
-& c_6\times Mindex+c_7\times relativeAge+\\
-& c_8\times relativeCitations
-\end{split}
-\end{equation*}
-
-\begin{equation*}
-\begin{split}
-speed=& c_0+c_{1}\times accuracy+c_{2}\times JIF+c_{3}\times H5+\\
-& c_{4}\times citations+c_{5}\times Hindex+\\
-& c_{6}\times Mindex+c_{7}\times relativeAge+\\
-& c_{8}\times relativeCitations
-\end{split}
-\end{equation*}
+%The linear models that we used to test for relationships between
+%speed, accuracy and the above measures are:
+%
+%\begin{equation*}
+%\begin{split}
+%accuracy=& c_0+c_1\times speed+c_2\times JIF+c_3\times H5+\\
+%& c_4\times citations+c_5\times Hindex+\\
+%& c_6\times Mindex+c_7\times relativeAge+\\
+%& c_8\times relativeCitations
+%\end{split}
+%\end{equation*}
+%
+%\begin{equation*}
+%\begin{split}
+%speed=& c_0+c_{1}\times accuracy+c_{2}\times JIF+c_{3}\times H5+\\
+%& c_{4}\times citations+c_{5}\times Hindex+\\
+%& c_{6}\times Mindex+c_{7}\times relativeAge+\\
+%& c_{8}\times relativeCitations
+%\end{split}
+%\end{equation*}
 
@@
 repeated 10,000 times. These permuted rankings were normalised and
-compared to the real rankings to produce the ‘x’ points in
+compared to the real rankings to produce the ‘$\vartimes$’ points in
 Figure~\ref{fig:allfactors}B and the heatmap and histograms in
@@
 \fussy
-\url{https://github.com/UCanCompBio/speed-vs-accuracy-meta-analysis}
+\url{https://github.com/Gardner-BinfLab/speed-vs-accuracy-meta-analysis}
 \sloppy
@@
 PPG is supported by a Rutherford Discovery Fellowship,
-administered by the Royal Society of New Zealand. 
+administered by the Royal Society Te Ap\=arangi, 
+PPG, AG and MB acknowledge support from a Data Science Programmes grant
+(UOAX1932).
+
 {\bf AND ETC....}

### 2021-02-24 manuscript/supplementary.tex
Updated README.md, manuscript and figures.

@@
 
-\PaperTitle{Supplementary results for: A meta-analysis of computational biology benchmarks reveals publication bias affects on speed and accuracy} % Article title
-
-%\Authors{Paul P.
-Gardner\textsuperscript{1,2}*, James Paterson\textsuperscript{1,2}, Fatemeh Ashari Ghomi\textsuperscript{1,2}, Sinan Uğur Umu\textsuperscript{1,2}, Stephanie McGimpsey\textsuperscript{1,2}} % Authors
-%\affiliation{\textsuperscript{1}\textit{School of Biological Sciences, University of Canterbury, Christchurch, New Zealand.}} % Author affiliation
-%\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre and the Bio-Protection Research Centre, University of Canterbury, Christchurch, New Zealand.}}
-%\affiliation{*\textbf{Corresponding author}: paul.gardner@canterbury.ac.nz} % Corresponding author
+\PaperTitle{Supplementary results for: A meta-analysis of bioinformatics software benchmarks reveals that publication-bias unduly influences software accuracy} % Article title
 
@@
 \includegraphics[width=0.45\textwidth]{cites-SpeedVsAccuracy-heatmap.pdf}
-\caption{Heatmaps of normalised speed ranks and normalised accuracy
+\caption{{\bf Citation metrics and software tool speeds/accuracies.} Heatmaps of normalised speed ranks and normalised accuracy
   ranks, both $x$ and $y$ dimensions are discretised into a $10 \times
-  10$ matrix. The shading indicates median H-indices for corresponding
-  authors (darker shade indicates a higher H) in the figure on the
-  left. The shading indicates median citations (on a log scale), for
-  software tools in the figure on the right. }
+  10$ matrix. The shading indicates different median citation-based
+  feature scores (redder shade indicates a higher value).  The shading
+  indicate median H-indices ({\bf left}) and $log_{10}(citations)$
+  ({\bf right}). }
 \label{fig:heatmaps}
@@
 \includegraphics[width=0.45\textwidth]{relAge-SpeedVsAccuracy-heatmap.pdf}
-\caption{{\bf Left.} Violin plots for the relative age distribution
+\caption{{\bf Relative software tool ages and speeds/accuracies.} {\bf Left.} Violin plots for the relative age distribution
   for software tools in each of the 5 2x2 cells indicated in B. The
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.45\textwidth]{commits-SpeedVsAccuracy-heatmap.pdf}
+\includegraphics[width=0.45\textwidth]{contributors-SpeedVsAccuracy-heatmap.pdf}
+\caption{{\bf Version control use and software tool speeds/accuracies.}
+  Heatmaps of normalised speed ranks and normalised accuracy
+  ranks, both $x$ and $y$ dimensions are discretised into a $10 \times
+  10$ matrix. The shading indicates different median github-derived
+  feature scores (redder shade indicates a higher value).  The shading
+  indicate the number of commits ({\bf left}) and the number of code contributors
+  ({\bf right}).
+ }
+\label{fig:heatmapsCommits}
+\end{figure}
+
+The complete list of benchmarks used for this study
+\cite{pmid32183840,
+pmid32138645,
+pmid31948481,
+pmid31874603,
+pmid31984131,
+pmid31639029,
+pmid31465436,
+pmid31324872,
+pmid31159850,
+pmid31136576,
+pmid31080946,
+pmid31015787,
+pmid30936559,
+pmid30717772,
+pmid30658573,
+pmid29568413,
+pmid28934964,
+pmid28808243,
+pmid28739658,
+pmid28569140,
+pmid28052134,
+pmid27256311,
+pmid26862001,
+pmid26778510,
+pmid26628557,
+pmid26220471,
+pmid25777524,
+pmid25760244,
+pmid25574120,
+pmid25521762,
+pmid25511303,
+pmid25198770,
+pmid24839440,
+pmid24708189,
+pmid24602402,
+pmid24526711,
+pmid24086547,
+pmid23842808,
+pmid23758764,
+pmid23593445,
+pmid23393030,
+pmid22574964,
+pmid22506536,
+pmid22492192,
+pmid22287634,
+pmid22152123,
+pmid22172045,
+pmid22132132,
+pmid21856737,
+pmid21615913,
+pmid21525877,
+pmid21483869,
+pmid21423806,
+pmid21113338,
+pmid20617200,
+pmid20047664,
+pmid19179695,
+pmid19126200,
+pmid19046431,
+pmid18793413,
+pmid18287116,
+pmid17151342,
+pmid17062146,
+pmid15840834,
+pmid15701525,
+ng2013estimating}.
+
 %----------------------------------------------------------------------------------------
@@
 \bibliographystyle{unsrt}
-\bibliography{paulall}
+\bibliography{supp-references}
 

### 2021-02-24 bin/prettyPlot.R

@@ -253,6 +253,7 @@ medRelAges <- 0*h2dNorm$counts
 medHindex  <- 0*h2dNorm$counts
 medCites   <- 0*h2dNorm$counts
 medCommits <- 0*h2dNorm$counts
+medContribs<- 0*h2dNorm$counts
 breaks <- seq(0,1, length=gridRes+1)
 
 for(i in 1:gridRes){
@@ -263,6 +264,7 @@ for(i in 1:gridRes){
 	    medHindex[i,j]  <- median(d2005$hindex[ breaks[i] <= d2005$accuracyRank & d2005$accuracyRank <= breaks[i+1] & breaks[j] <= d2005$speedRank & d2005$speedRank <= breaks[j+1]],na.rm=TRUE)
 	    medCites[i,j]   <- median(  log10(d2005$citations[  breaks[i] <= d2005$accuracyRank & d2005$accuracyRank <= breaks[i+1] & breaks[j] <= d2005$speedRank & d2005$speedRank <= breaks[j+1]]+1),na.rm=TRUE)
 	    medCommits[i,j]   <- median(  log10(d2005$commits[  breaks[i] <= d2005$accuracyRank & d2005$accuracyRank <= breaks[i+1] & breaks[j] <= d2005$speedRank & d2005$speedRank <= breaks[j+1]]+1),na.rm=TRUE)
+	    medContribs[i,j]   <- median(  log10(d2005$contributors[  breaks[i] <= d2005$accuracyRank & d2005$accuracyRank <= breaks[i+1] & breaks[j] <= d2005$speedRank & d2005$speedRank <= breaks[j+1]]+1),na.rm=TRUE)
       }
 }
 
@@ -363,7 +365,7 @@ dev.off()
 colScale <- 9
 pdf(file=    "../figures/relAge-SpeedVsAccuracy-heatmap.pdf", width = 10,  height = 10)
 par(mar = c(8,4,4,8) + .1) #c(bottom, left, top, right). default: c(5, 4, 4, 2) + 0.1
[…]

### 2021-02-25 manuscript/manuscript-speed-accuracy.tex
Added limitations discussion to manuscript, fixed figures, removed highlighting of values.

@@
 editor and reviewer practices. This leaves an unfortunate gap in the
-literature as the ideal tools may fall in the gap e.g. high accuracy is generally not useful if years of CPU time is required, 
-while high speed is not useful if the results are very inaccurate.  }
+literature as the ideal tools may fall in the gap e.g. high accuracy tools are not always useful if years of CPU time are required, 
+while high speed is not useful if the results are also inaccurate.  }
 
@@
 software-based solutions often exists
-\cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut,Wikipedia_contributors2015-vj,Wikipedia_contributors2015-hr}. While
+\cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut}. While
 this is a good problem to have, and points to a
@@
 With the wide adoption of GitHub ($\approx 47\%$ of the
-\textbf{{\color{red}\numTools}} tools included in this study could be
-linked to a GitHub repository), quantifiable data on software
+{\color{black}\numTools} tools included in this study could be
+linked to a GitHub repository), and consequently quantifiable data on software
 development such as the number of contributors to code, number of code
@@
 of computational biology software, and manually extracted accuracy and
-speed rankings for \textbf{{\color{red}\numTools~}} unique software tools. For
+speed rankings for {\color{black}\numTools~} unique software tools. For
 each tool, we have collected measures that may be predictive
@@
 We have collected relative accuracy and speed ranks for
-\textbf{{\color{red}\numTools~}} distinct software tools. This software has
+{\color{black}\numTools~} distinct software tools. This software has
 been developed for solving broad cross-section computational biology
@@
 Each software tool was benchmarked in at least one of
-\textbf{{\color{red}\numBenchmarkPubs~}} publications that satisfy the Boulesteix
+{\color{black}\numBenchmarkPubs~} publications that satisfy the Boulesteix
 criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria
@@
 
-For each of the publications describing these tools, we have (where
-possible) collected the journal's H5-index (\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
-  Scholar Metrics}), the maximum H-index and
-corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
-authors for each tool, and the number of times the publication(s)
-associated with a tool has been cited using Google Scholar (data
-collected over a 6 month period in late 2020). Note that citation metrics
-are not static and will change over time. In addition, where possible we also extract the version number, the number
-of commits and number of contributors from GitHub repositories.
-
-We have computed the Spearman’s correlation coefficient for each
-pairwise combination of the mean normalised accuracy and speed ranks,
-with the year published, mean relative age (compared to software in the
-same benchmarks), journal H5 metrics, the total number of citations,
-the relative number of citations (compared to software in the same
-benchmarks) and the maximum H- and corresponding M-indices for the corresponding
-authors, version number, and numbers of commits and contributors.
-The results are presented in Figure~\ref{fig:allfactors}A. We find
-significant associations between most of the citation-based metrics
-(journal H5, citations, relative citations, H-index and
-M-index). There is also a negative correlation between the year of
-publication, the relative age and many of the citation-based metrics.
-
-\begin{figure*}[ht]
+\begin{figure*}[htb!]
 \includegraphics[width=\textwidth]{figure1.pdf}
@@
 
+For each of the publications describing these tools, we have (where
+possible) collected the journal's H5-index (\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
+  Scholar Metrics}), the maximum H-index and
+corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
+authors for each tool, and the number of times the publication(s)
+associated with a tool has been cited using Google Scholar (data
+collected over a 6 month period in late 2020). Note that citation metrics
+are not static and will change over time. In addition, where possible we also extract the version number, the number
+of commits and number of contributors from GitHub repositories.
+
+We have computed the Spearman’s correlation coefficient for each
+pairwise combination of the mean normalised accuracy and speed ranks,
+with the year published, mean relative age (compared to software in the
+same benchmarks), journal H5 metrics, the total number of citations,
+the relative number of citations (compared to software in the same
+benchmarks) and the maximum H- and corresponding M-indices for the corresponding
+authors, version number, and numbers of commits and contributors.
+The results are presented in Figure~\ref{fig:allfactors}A. We find
+significant associations between most of the citation-based metrics
+(journal H5, citations, relative citations, H-index and
+M-index). There is also a negative correlation between the year of
+publication, the relative age and many of the citation-based metrics.
+
 Data on the
 number of updates to software tools from GitHub such as the number of versions,
-commits and contributors \textbf{{\color{red}was significantly correlated with 
-accuracy (respective Spearman’s rhos = 0.13, 0.22, 0.21, respective uncorrected P-values = 0.0027, 0.00079, 0.0013). 
+commits and contributors was significantly correlated with 
+accuracy (respective Spearman’s rhos = {\color{black}0.13, 0.22, 0.21,} respective uncorrected P-values = {\color{black}0.0027, 0.00079, 0.0013}). 
 These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A).
-}}
 We also found that author reputation metrics, journal impacts and the age
@@
 accuracy or speed (see Figure~\ref{fig:allfactors}). The strongest
-association was between accuracy and \textbf{{\color{red}the relative
-    number of citations}} (Spearman’s rho = \textbf{{\color{red}0.092,
-    P-value = 0.047}}). But the effect size is only just over the
-significance threshold and may be explained by multiple-testing.
+association was between accuracy and the relative
+number of citations (Spearman’s rho = {\color{black}0.092,
+P-value = 0.047}). But the effect size is only just over the
+significance threshold and may be the result of multiple-testing.
 
@@
 expected distributions generated by 1,000 randomized ranks. The results
-of this are shown in Figure~\ref{fig:speedaccuracy}. We identified \textbf{{\color{red}18}}
+of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}18}
 bins where there was a significant excess or dearth of tools. For
-example, there was an excess of “slow and inaccurate” software (\textbf{{\color{red}Z=2.50,
-P-value=0.0063}}). We find that the amount of software classed as “fast
+example, there was an excess of “slow and inaccurate” software ({\color{black}Z=2.50,
+P-value=0.0063}). We find that the amount of software classed as “fast
 and accurate”, “fast and inaccurate” and "slow and accurate" are at approximately the
@@
 significant results is in excess, and is not due to multiple testing, as
-the probability of finding \textbf{{\color{red}18}} of 100 tests significant by chance is
-low (P-value = \textbf{{\color{red}$2.2\times 10^{-6}$}}, exact binomial test).
+the probability of finding {\color{black}18} of 100 tests significant by chance is
+low (P-value = {\color{black}$2.2\times 10^{-6}$}, exact binomial test).
 
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[htb!]
 %\includegraphics[scale=0.55]{figure2.pdf}
@@
 corresponding to the four central deciles are
-highlighted \textbf{{\color{red}(Z = -1.9, -2.5, -2.9 and -1.1, P-values = 0.03, 0.006,
-0.002 and 0.14}}, respectively, reading from top to bottom, left to
+highlighted {\color{black}(Z = -1.9, -2.5, -2.9 and -1.1, P-values = 0.03, 0.006,
+0.002 and 0.14}, respectively, reading from top to bottom, left to
 right). We also tested the relative age of the software tools in
@@
 accuracy plot. We found that the “slow and inaccurate” tools ($N=27$) were
-generally published earlier than ``fast and accurate'' tools ($N=23$) \textbf{{\color{red}(W = 341, P=0.005,
-one-tailed Wilcoxon test)}} (Figure S7).
+generally published earlier than ``fast and accurate'' tools ($N=23$) {\color{black}(W = 341, P=0.005,
+one-tailed Wilcoxon test)} (Figure S7).
 
@@
 We have gathered data on the relative speeds and accuracies of
-\textbf{{\color{red}\numTools~}} bioinformatic tools from
-\textbf{{\color{red}\numBenchmarkPubs}} benchmarks published between
-2005 and 2020. {\bf Our results suggest there are major benefits for
-  long term support for software \cite{siepel2019challenges}.} The
-finding of a strong relationship between the number of commits to
+{\color{black}\numTools~} bioinformatic tools from
+{\color{black}\numBenchmarkPubs} benchmarks published between
+2005 and 2020. Our results provide significant support for the suggestion that there are major benefits to
+the long-term support of software development \cite{siepel2019challenges}. The
+finding of a strong relationship between the number of commits and code contributors to
 GitHub (i.e. software updates) and accuracy, highlights the benefits of
@@
 software is generally published earlier than fast and accurate tools
-\textbf{{\color{red}(P=0.007, one-tailed Wilcoxon test)}}, therefore
+{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
 comparisons were not required to publish the slow and
@@
 
-{\bf DESCRIBE STUDY LIMITATIONS... 
-\begin{itemize}
-    \item Difficult to test publication biases
-    \item Heterogeneous measures of accuracy between benchmarks  
-    \item possible lack of independence between different benchmarks 
-    \item Non-uniform inclusion of tools in different benchmarks -- some tools ranked many times (e.g. velvet and soap assemblers ranked 26 different times)
-    \item How to deal with inconsistently reported versioning of software? 
-    \item Metrics and tools performance are all moving targets, only a snap shot is used here  
-\end{itemize}
-}
+While we have taken pains to mitigate many issues with our analysis, nevertheless 
+some limitations remain. For example, it has proven difficult to verify if the gap in medium accuracy 
+and medium speed software is genuinely the result of publication bias, or due to additional factors that we have 
+not taken in to account. In addition, all of the features we have used here are moving targets, e.g. as software 
+tools are refined, their relative accuracies and speeds will change, the citation metrics, ages, and version control 
+derived measures also change over time. We merely report a snapshot of values from 2020.  The benchmarks themselves may 
+also introduce biases into the study, for example there are issues with a potential lack of independence between benchmarks 
+(e.g. shared datasets, metrics and tools), there are heterogeneous measures of accuracy and speed and often unclear 
+processes for including different tools.
+
+%{\bf DESCRIBE STUDY LIMITATIONS... 
+%\begin{itemize}
+%    \item Difficult to test publication biases
+%    \item Heterogeneous measures of accuracy between benchmarks  
+%    \item possible lack of independence between different benchmarks 
+%    \item Non-uniform inclusion of tools in different benchmarks -- some tools ranked many times (e.g. velvet and soap assemblers ranked 26 different times)
+%    \item How to deal with inconsistently reported versioning of software? 
+%    \item Metrics and tools performance are all moving targets, only a snap shot is used here  
+%\end{itemize}
+%}
 
@@
 that certain approaches are not practical research avenues
-\cite{Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
+\cite{fanelli2012negative}. % Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
 The current
 practices of publishers, editors, reviewers and authors of software
-tools therefore deprive our community of tools for building effective
+tools therefore may be depriving our community of tools for building effective
 and productive workflows.
 
-The most reliable way to select software tools is through neutral
+The most reliable way to identify accurate software tools is through neutral
 software benchmarks \cite{Boulesteix2013-vb}. We are hopeful that
@@
 $\delta$
-is a prior probability ($\delta = 10^{-5}$, by default),
+is a pseudo-count added for each word ($\delta = 10^{-5}$, by default),
 $f_{bg}(word)$ and $f_{tr}(word)$ are the frequencies of a $word$ in
@@
 into the set of background articles. This process is iterated and has resulted in the identification of
-\textbf{{\color{red}\numBenchmarkPubs}} benchmark articles, that
-contain \textbf{{\color{red}\numBenchmarks}} different benchmarks, together these
-rank \textbf{{\color{red}\numTools~}} distinct software packages.
+{\color{black}\numBenchmarkPubs} benchmark articles, that
+contain {\color{black}\numBenchmarks} different benchmarks, together these
+rank {\color{black}\numTools~} distinct software packages.
 
@@
 Raw datasets, software and documents are available under a CC-BY license:\\
-\fussy
-\url{https://docs.google.com/spreadsheets/d/14xIY2PHNvxmV9MQLpbzSfFkuy1RlzDHbBOCZLJKcGu8/edit?usp=sharing}\\
-and here:\\
-\fussy
-\url{https://dx.doi.org/10.6084/m9.figshare.4299320.v1}\\
-\sloppy
-Additional documentation, code, figures and raw data is available here:\\
+%\fussy
+%\url{https://docs.google.com/spreadsheets/d/14xIY2PHNvxmV9MQLpbzSfFkuy1RlzDHbBOCZLJKcGu8/edit?usp=sharing}\\
+%and here:\\
+%\fussy
+%\url{https://dx.doi.org/10.6084/m9.figshare.4299320.v1}\\
+%\sloppy
+%Additional documentation, code, figures and raw data is available here:\\
 \fussy
@@
 
-{\bf AND ETC....}
+%{\bf AND ETC....}
 

### 2021-02-28 bin/prettyPlot.R
Updated Fig1. added multiple testing correction and wheat-coloured violin plots.

@@ -62,7 +62,7 @@ summary(pca)
 #All vs All Spearman Heatmap
 
 dNames <- c("H5", "relAge",    "yearPublished", "citations", "relCites",   "mindex",  "hindex",  "accuracyRank",  "speedRank", "version", "commits", "contributors" )
-pNames <- c("JH5", "Rel. age", "Year",          "Citations", "Rel. cites", "M-index", "H-index", "Accuracy",      "Speed", "version", "commits", "contributors")
+pNames <- c("JH5", "Rel. age", "Year",          "Citations", "Rel. cites", "M-index", "H-index", "Accuracy",      "Speed",     "version", "commits", "contributors")
 pvalMatrix<-matrix(1, length(dNames), length(dNames))
 rhoMatrix <-matrix(0, length(dNames), length(dNames))
 sigMatrix <-matrix("",length(dNames), length(dNames))
@@ -75,12 +75,14 @@ colnames(sigMatrix)     <-pNames
 rownames(sigMatrix)     <-pNames
 sigCount     <- 0
 sigCount2015 <- 0
-for(i in 1:length(dNames)){
-      for(j in 1:length(dNames)){
+N <- length(dNames)
+for(i in 1:N){
+      for(j in 1:N){
 	    spear<-cor.test(d2005[,dNames[i] == colnames(d2005)], d2005[,dNames[j] == colnames(d2005)], method = "spearman", exact = T) 
-	    pvalMatrix[i,j] <- spear$p.value 
+	    pvalMatrix[i,j] <- p.adjust(spear$p.value,"BH", n=N*(N-1)/2-N)
 	    rhoMatrix[i,j]  <- spear$estimate
-	    if(spear$p.value < 0.05){
+	    #print(paste(c("P:", round(spear$p.value,3), "P.adj:", round(p.adjust(spear$p.value,"BH", n=N*(N-1)/2-N),3)), sep=""))
+	    if(p.adjust(spear$p.value,"BH", n=N*(N-1)/2-N) <= 0.05){ #spear$p.value < 0.05){
 		sigMatrix[i,j]  <- "X"
                 sigCount <- sigCount + 1
 	    }
@@ -179,10 +181,14 @@ slowInacc <- vector(mode = "numeric", length = 4*numPerms)
 slowAcc   <- vector(mode = "numeric", length = 4*numPerms)
[…]

### 2021-03-12 manuscript/manuscript-speed-accuracy.tex
Major revision for manuscript and figures after comments from Mik.

@@
 
-\PaperTitle{A meta-analysis of bioinformatics software benchmarks reveals that publication-bias unduly influences software accuracy} % Article title
+\PaperTitle{A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy} 
+%Analysis of sixty software benchmarks shows the benefits of good software practice 
+%Assessing the state of the bioinformatic software landscape with a meta-analysis of benchmarks
+%Sustained software development, not citation metrics, is indicative of accurate bioinformatic software 
+%
 
@@
 %\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
-\Abstract{ Computational biology has provided widely used and powerful
+\Abstract{Computational biology provides widely used and powerful
   software tools for testing and making inferences about biological
   data. In the face of rapidly increasing volumes of data, heuristic
-  methods that trade software speed for accuracy must
-  be employed.  We are interested in testing for trade-offs between
-  speed and accuracy and for other factors that are indicative of accurate
+  methods that trade software speed for accuracy may
+  be employed.  We are have studied these trade-offs using the results of a large number of independent software benchmarks, 
+  and evaluated whether external factors are indicative of accurate
   software.
@@
 We have extracted accuracy and speed ranks from independent benchmarks of
-computational biology software of different software tools, and evaluated the
-factors that are likely to influence accuracy e.g. speed, author
-reputation, journal impact, recency or developer efforts.
+different bioinformatic software tools, and evaluated whether the
+speed, author reputation, journal impact, recency and developer efforts are indicative of accuracy.
 
-We found that software speed, author reputation, journal impact, the
+We found that software speed, author reputation, journal impact, 
 number of citations and age are all unreliable predictors of software
 accuracy. This is unfortunate because citations, author and journal
-reputation are frequently used reasons for selecting software
-tools. Our results do show that accurate bioinformatic software tools
-are generally the product of many gradual improvements, often from
-multiple developers, based upon matching GitHub records with benchmark results. 
-In addition, we find
-that the field of bioinformatics has an excess of slow and inaccurate software
-tools, this is consistent across many sub-disciplines. Meanwhile, there are few tools that
+reputation are frequently cited reasons for selecting software
+tools. However, GitHub-derived records and high version numbers show that 
+the accurate bioinformatic software tools
+are generally the product of many improvements over time, often from
+multiple developers.
+
+We also find
+that the field of bioinformatics has a large excess of slow and inaccurate software
+tools, and this is consistent across many sub-disciplines. Meanwhile, there are few tools that
 are middle-of-road in terms of accuracy and speed trade-offs. We
-hypothesise that a form of publication-bias unduly influences the
+hypothesise that a form of publication-bias influences the
 publication and development of bioinformatic software. In other
-words, software that is intermediate in terms of both speed and accuracy may be difficult to publish due to author,
-editor and reviewer practices. This leaves an unfortunate gap in the
-literature as the ideal tools may fall in the gap e.g. high accuracy tools are not always useful if years of CPU time are required, 
+words, software that is intermediate in terms of both speed and accuracy may be difficult to publish - possibly due to author,
+editor and reviewer practices. This leaves an unfortunate hole in the
+literature as the ideal tools may fall into this gap. For example, high accuracy tools are not always useful if years of CPU time are required, 
 while high speed is not useful if the results are also inaccurate.  }
@@
 the most cited publications in the entire scientific corpus
-\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. This
-highly-cited software includes implementations of methods for sequence
-alignment and homology inference
-\cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
-phylogenetic analysis
-\cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei},
-statistical analysis of survival patterns in biomedicine
-\cite{Kaplan1958-ju,Cox1972-nu}, biomolecular structure analysis
-\cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj},
-visualization and data collection
-\cite{Kraulis1991-lt,Berman2000-to}. However, the popularity of a
+\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. These
+highly-cited software tools include implementations of methods for sequence
+alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
+phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, 
+biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, and
+visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}. 
+However, the popularity of a
 software tool does not necessarily mean that it is accurate or
 computationally efficient, instead usability, ease of installation,
-operating system and other factors may play a greater role in a software
+operating system support or other indirect factors may play a greater role in a software
 tool's popularity. Indeed, there have been several notable incidences where
-inaccurate software has caused considerable harm
+convenient, yet inaccurate software has caused considerable harm
 \cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
@@
 species tracking), therefore the biological sciences have become
-increasingly dependent upon computational software for processing
+increasingly dependent upon software tools for processing
 large quantities of data \cite{Marx2013-zi}. As a consequence, the
@@
 of great importance to decrease the energy, climate impact, and time costs of research
-\cite{Gombiner2011-md}. Furthermore, even small error rates can have
-major impacts on the number of false inferences as datasets become
-larger \cite{Storey2003-cv}.
+\cite{Gombiner2011-md}. Furthermore,  as datasets become
+larger even small error rates can have
+major impacts on the number of false inferences \cite{Storey2003-cv}.
 
-The gold-standard for determining accuracy is for independent
-researchers to conduct benchmarks, which can serve a useful role in
+The gold-standard for determining accuracy is for researchers independent
+of individual tool development to conduct benchmarking studies, these benchmarks can serve a useful role in
 reducing the over-optimistic reporting of software accuracy
 \cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
-\cite{Norel2011-cq,Buchka:2021}. Benchmark studies typically use a number of
-positive and negative control datasets, predictions from different software tools can then be
-partitioned into true or false groups and a variety of metrics can be
-used to evaluate the performance differences
+\cite{Norel2011-cq,Buchka:2021}. Benchmarking typically involves the use a number of
+positive and negative control datasets, so that predictions from different software tools can be
+partitioned into true or false groups, allowing a variety of metrics to be
+used to evaluate performance
 \cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
@@
 %\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
-The aim of these benchmarks is to independently identify tools that
+The aim of these benchmarks is to robustly identify tools that
 make acceptable compromises in terms of balancing speed with
@@
 any recent gold-standard benchmarks, how should scientific software be
-selected? In the following we presume that ``biological
-accuracy'' is the most desirable feature for a software tool. Biological
+selected? In the following we presume that the ``biological
+accuracy'' of predictions is the most desirable feature for a software tool. Biological
 accuracy is the degree to which predictions or measurements reflect
-the biological truths based on expert-derived curated datasets (see Methods for the  mathematical definition used here). In some
-fields biological accuracy is very difficult to ascertain, for
-example, in phylogenetics it is nearly impossible in most cases to know the
-true ancestral relationships between organisms. In situations like this,
-researchers may use simulated or high-confidence datasets.
+the biological truths based on expert-derived curated datasets (see Methods for the  mathematical definition used here). 
+%In some
+%fields biological accuracy is difficult to ascertain, for
+%example, in phylogenetics it is nearly impossible in most cases to know the
+%true ancestral relationships between organisms. In situations like this,
+%researchers may use simulated or high-confidence datasets to assess
+%software performance.
 
 A number of possible predictors of software quality are used by the
-community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. Some accessible,
-quantifiable and frequently used proxies for identifying high quality
+community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
+Some accessible, quantifiable and frequently used proxies for identifying high quality
 software include: \textbf{1. Recency:} recently published software
 tools may have built upon the results of past work, or be an update to
-an existing software. Therefore these may be more accurate and
+an existing tool. Therefore these may be more accurate and
 faster. \textbf{2. Wide adoption:} a software tool may be widely used
 because it is fast and accurate, or because it is well-supported and
-user-friendly. In fact, “large user base”, “word-of-mouth”,
-“wide-adoption”, "personal recommendation," and "recommendation from a
-close colleague," are frequent responses to surveys of “how do
-scientists select software?”
+user-friendly. In fact,``large user base'', ``word-of-mouth'',
+``wide-adoption'', ``personal recommendation'', and ``recommendation from a
+close colleague'', were frequent responses to surveys of ``how do
+scientists select software?''
 \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
@@
 consequence, an argument could be made that well respected and
-high-profile authors will produce better software
+high-profile authors may write better software
 \cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
@@
 \cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
-interpret this fact as slower software is likely to be more
+interpret this fact as implying that slower software is likely to be more
 accurate. But speed may also be influenced by the programming language
@@
 \cite{Farrar2007-ky,Dematte2010-ph};
-however, the implementation generally has more of an impact (e.g. brute-force approaches versus
+However, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
 rapid and sensitive pre-filtering
-\cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).
+\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
 %\todo{This paper
@@
 
-Other factors that are less quantifiable that influence whether a
-software tool is selected include: whether the documentation is good,
-user-friendly, word-of-mouth and ``used in a similar analysis''
-\cite{Loman2015-bw}. However, citation metrics may be a
-useful proxy for the above.
+%Other factors that are less quantifiable that influence whether a
+%%software tool is selected include: whether the documentation is good,
+%user-friendly, word-of-mouth and ``used in a similar analysis''
+%\cite{Loman2015-bw}. However, citation metrics may be a
+%useful proxy for the above.
 %This may also explain the reason why
@@
 
-With the wide adoption of GitHub ($\approx 47\%$ of the
+With the wide adoption of GitHub ($47\%$ of the
 {\color{black}\numTools} tools included in this study could be
 linked to a GitHub repository), and consequently quantifiable data on software
-development such as the number of contributors to code, number of code
-changes and versions is also available
+development time and intensity indicators, such as the number of contributors to code, number of code
+changes and versions is available for these
 \cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.
@@
 {\color{black}\numTools~} distinct software tools. This software has
-been developed for solving broad cross-section computational biology
+been developed for solving a broad cross-section of computational biology
 tasks.
@@
 \caption{\textbf{A.} A heatmap indicating the relationships between
-  proposed predictors of software quality. Spearman’s rho is used to
-  infer correlations between metrics such as the H- and M-indices of
-  corresponding authors, number of citations, journal impact factors
-  and H5 indices, the year and relative age of software and the mean
-  relative rankings of software speed and accuracy. Red colours
+  different features of bioinformatic software tools. Spearman’s rho is used to
+  infer correlations between metrics such as citations based metrics, the year and relative age of publication, 
+  version number, GitHub derived activity measures, and 
+  the mean
+  relative speed and accuracy rankings. Red colours
   indicate a positive correlation, blue colours indicate a negative
-  correlation. Correlations with an uncorrected P-value less than 0.05 are
-  indicated with a `$\vartimes$'. The dashed rectangular area is illustrated in
+  correlation. Correlations with an P-value less than 0.05 (corrected for multiple-testing using the Benjami-Hochberg method) are
+  indicated with a `$\vartimes$' symbol. The correlations with accuracy are illustrated in
   more detail in \textbf{B}, the relationship between speed and
@@
   %hierarchical clustering.
-  \textbf{B.} A barplot illustrating the correlation, as measured by
-  Spearman’s rho, between normalised accuracy ranks and software tool
-  features that may be predictive of accuracy. In order to give an
-  appreciation of the difference between the observed effect-sizes and
-  significant effect sizes, we generated 1000 permuted accuracy ranks
-  for each benchmark and recorded Spearman’s rho for the significant
-  correlations with an uncorrected P-value less than 0.05.
-  These values are marked with blues `$\vartimes$'s in the
-  barplot. Significant correlations are marked with red asterisks.}
+  \textbf{B.} Violin plots of Spearman's correlations between permuted accuracy ranks and 
+    different software tool features. 
+    The unpermuted correlations are indicated with a red asterix.
+    For each benchmark, 1,000 permuted sets of accuracy and speed ranks were generated, and the ranks were normalised to lie between 0 and 1 (see Methods for details).
+  Circled asterixs are significant (empirical P-value $< 0.05$, corrected for multiple-testing using the Benjami-Hochberg method).}
 \label{fig:allfactors}
@@
 authors, version number, and numbers of commits and contributors.
-The results are presented in Figure~\ref{fig:allfactors}A. We find
+The results are presented in Figure~\ref{fig:allfactors}A\&B. We find
 significant associations between most of the citation-based metrics
@@
 commits and contributors was significantly correlated with 
-accuracy (respective Spearman’s rhos = {\color{black}0.13, 0.22, 0.21,} respective uncorrected P-values = {\color{black}0.0027, 0.00079, 0.0013}). 
-These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A).
-We also found that author reputation metrics, journal impacts and the age
+software accuracy (respective Spearman’s rhos = {\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg corrected P-values = {\color{black}0.029, 0.042, 0.063}), Figure~\ref{fig:allfactors}B.
+The significance if these features was further confirmed with a permutation test (Figure~\ref{fig:allfactors}B).
+These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A \& Supplementary Figure ????).
+We also found that reputation metrics such as citations, author and journal H-indices, and the age
 of tools were generally \textbf{not} correlated with either tool
-accuracy or speed (see Figure~\ref{fig:allfactors}). The strongest
-association was between accuracy and the relative
-number of citations (Spearman’s rho = {\color{black}0.092,
-P-value = 0.047}). But the effect size is only just over the
-significance threshold and may be the result of multiple-testing.
+accuracy or speed (Figure~\ref{fig:allfactors}A\&B). 
+%The strongest
+%association was between accuracy and the relative
+%number of citations (Spearman’s rho = {\color{black}0.092,
+%P-value = 0.047}). But the effect size is only just over the
+%significance threshold and may be the result of multiple-testing.
 
@@
 from each benchmark were randomly permuted, generating 1,000
-randomized speed and accuracy ranks. In the cells of a $10\times10$
+randomized speed and accuracy ranks. In the cells of a $3\times3$
 grid spanning the normalised speed and accuracy ranks we computed a
@@
 expected distributions generated by 1,000 randomized ranks. The results
-of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}18}
-bins where there was a significant excess or dearth of tools. For
-example, there was an excess of “slow and inaccurate” software ({\color{black}Z=2.50,
-P-value=0.0063}). We find that the amount of software classed as “fast
-and accurate”, “fast and inaccurate” and "slow and accurate" are at approximately the
-expected proportions based upon the permutation test. 
-While there are very few tools placed in the middle of the ranks on accuracy and speed.
-The number of
-significant results is in excess, and is not due to multiple testing, as
-the probability of finding {\color{black}18} of 100 tests significant by chance is
-low (P-value = {\color{black}$2.2\times 10^{-6}$}, exact binomial test).
+of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}4}
+of 9 bins where there was a significant excess or dearth of tools. For
+example, there was an excess of ``slow and inaccurate'' software ({\color{black}Z=3.39,
+P-value=$3.5\times 10^{-4}$}), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively).
+We find that only the ``fast and inaccurate'' extreme class is at approximately the
+expected proportions based upon the permutation test (Figure 2B).
 
@@
 \includegraphics[width=0.75\textwidth]{figure2.pdf}
-\caption{A heatmap indicating the relative paucity or abundance of
-  software in the range of possible accuracy and speed rankings. Red
+\caption{\textbf{A.} A heatmap indicating the relative paucity or abundance of
+  software in the range of possible accuracy and speed rankings. Redder
   colours indicate an abundance of software tools in an accuracy and
-  speed category, while blue colours indicate scarcity of software in
+  speed category, while bluer colours indicate scarcity of software in
   an accuracy and speed category. The abundance is quantified using a
-  Z-score computation for each bin, this is derived from 10,000 random
-  permutations of the speed and accuracy ranks from each
+  Z-score computation for each bin, this is derived from 1,000 random
+  permutations of speed and accuracy ranks from each
   benchmark. Mean normalised ranks of accuracy and speed have been
-  binned into 100 classes (a $10\times10$ grid) that range from
-  comparatively slow/inaccurate to comparatively
-  fast/accurate. Z-scores with a P-value less than 0.05 are indicated
-  with a ‘$\vartimes$’.}
+  binned into 9 classes (a $3\times 3$ grid) that range from
+  comparatively slow and inaccurate to comparatively
+  fast and accurate. Z-scores with a P-value less than 0.05 are indicated
+  with a ‘$\vartimes$’. \textbf{B.} The z-score distributions from the permutation tests (indicated with the wheat coloured violin plots) compared to 
+  the z-score for the observed values for each of the corner and middle square of the heatmap.}
 \label{fig:speedaccuracy}
@@
 
-There is a major reduction in the number of software tools that are
+The largest difference between the observed and expected software ranks is the reduction in the number of software tools that are
 classed as intermediate in terms of both speed and accuracy based on
-permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The cells
-corresponding to the four central deciles are
-highlighted {\color{black}(Z = -1.9, -2.5, -2.9 and -1.1, P-values = 0.03, 0.006,
-0.002 and 0.14}, respectively, reading from top to bottom, left to
-right). We also tested the relative age of the software tools in
-the indicated corners and central regions of the speed vs
-accuracy plot. We found that the “slow and inaccurate” tools ($N=27$) were
-generally published earlier than ``fast and accurate'' tools ($N=23$) {\color{black}(W = 341, P=0.005,
-one-tailed Wilcoxon test)} (Figure S7).
+permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The middle cell
+of Figure 2A and left-most violin plot of Figure 2B highlight this extreme, (Z=-6.38,
+P-value=$9.0\times 10^{-11}$).
 
@@
 GitHub (i.e. software updates) and accuracy, highlights the benefits of
-long-term development.
+long-term or at least intensive development.
 
 Our study finds little evidence to support that impact-based metrics
-have any relationship with software quality, this is unfortunate as these
+have any relationship with software quality, which is unfortunate, as these
 are frequently cited reasons for selecting software tools
-\cite{Loman2015-bw}. This implies that the high citation
+\cite{Loman2015-bw}. This implies that high citation
 rates for bioinformatic software
@@
 high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
-In addition, we find an under-representation of software that has
+In addition, there is an large under-representation of software that has
 intermediate levels of both accuracy and speed. A possible explanation
@@
 that is not highly ranked on speed or accuracy. If correct, this may
-have unfortunate consequences.  We also found that slow and inaccurate
-software is generally published earlier than fast and accurate tools
-{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
-comparisons were not required to publish the slow and
-inaccurate tools.
+have unfortunate consequences as these tools may never-the-less have further uses.
+%We also found that slow and inaccurate
+%software is generally published earlier than fast and accurate tools
+%{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
+%comparisons with existing software were not required to publish the slow and
+%inaccurate tools.
 
@@
 and medium speed software is genuinely the result of publication bias, or due to additional factors that we have 
-not taken in to account. In addition, all of the features we have used here are moving targets, e.g. as software 
+not taken in to account. In addition, all of the features we have used here are moving targets. For example, as software 
 tools are refined, their relative accuracies and speeds will change, the citation metrics, ages, and version control 
-derived measures also change over time. We merely report a snapshot of values from 2020.  The benchmarks themselves may 
-also introduce biases into the study, for example there are issues with a potential lack of independence between benchmarks 
-(e.g. shared datasets, metrics and tools), there are heterogeneous measures of accuracy and speed and often unclear 
+derived measures also change over time. Here we report a snapshot of values from 2020.  The benchmarks themselves may 
+also introduce biases into the study. For example, there are issues with a potential lack of independence between benchmarks 
+(e.g., shared datasets, metrics and tools), there are heterogeneous measures of accuracy and speed and often unclear 
 processes for including different tools.
@@
 We propose that the full spectrum of software tool accuracies and
-speeds serve a useful purpose to the research community. Like negative
-results, if honestly reported, illustrate to the research community
+speeds serves a useful purpose to the research community. Like negative
+results, if honestly reported this information, illustrates to the research community
 that certain approaches are not practical research avenues
@@
 \textbf{Criteria for inclusion:} We are interested in using
-computational biology benchmarks that satisfy Anne-Laure Boulesteix’s
+computational biology benchmarks that satisfy Boulesteix’s
 (ALB) three criteria for a “neutral comparison study”
@@
 
-Literature mining: We identified an initial list of 10 benchmark
+\textbf{Literature mining:} We identified an initial list of 10 benchmark
 articles that satisfy the ALB-criteria. These were identified based
 upon previous knowledge of published articles and were supplemented
-with several literature searches (e.g. ‘“benchmark” AND “cputime”’ was
+with several literature searches (e.g., [``benchmark'' AND ``cputime''] was
 used to query both GoogleScholar and Pubmed
@@
 For our machine-learning-based literature screening, we computed a
-score ($s(a)$) for each article that tells us the likelihood that it
+score, $s(a)$, for each article that tells us the likelihood that it
 is a benchmark. In brief, our approaches uses 3 stages:
@@
 by combining text from titles and abstracts for the background and
-training articles. A log-odds score is computed for each word using
+training articles. A log-odds score was computed for each word using
 the following formula:
 %\todo{Word is denoted $w$ on the left and $word$ on the right.}
-$lo(word)=\log_2\frac{f_{tr}(word)+\delta}{f_{bg}(word)+\delta}$, where
+
+\[lo(word)=\log_2\frac{f_{tr}(word)+\delta}{f_{bg}(word)+\delta}\]
+
+Where
 $\delta$
-is a pseudo-count added for each word ($\delta = 10^{-5}$, by default),
-$f_{bg}(word)$ and $f_{tr}(word)$ are the frequencies of a $word$ in
+was a pseudo-count added for each word ($\delta = 10^{-5}$, by default),
+$f_{bg}(word)$ and $f_{tr}(word)$ were the frequencies of a $word$ in
 the background and training datasets respectively. Word frequencies
-are computed by counting the number of times a word appears in the
-pool of titles and abstracts, the counts are normalised by the total
+were computed by counting the number of times a word appears in the
+pool of titles and abstracts, the counts were normalised by the total
 number of words in each set.
@@
 Thirdly, we also collected a group of candidate benchmark articles by
-mining Pubmed for articles that are likely to be benchmarks of
-bioinformatic software, these may match the terms: “((bioinformatics)
+mining Pubmed for articles that were likely to be benchmarks of
+bioinformatic software, these match the terms: “((bioinformatics)
 AND (algorithms OR programs OR software)) AND (accuracy OR assessment
@@
 then manually evaluated against the ALB-criteria. Accuracy and speed
-ranks are extracted from the articles that meet the criteria, and
-these are also added to the set of training articles. The evaluated
-candidate articles that do not meet the ALB-criteria are incorporated
-into the set of background articles. This process is iterated and has resulted in the identification of
-{\color{black}\numBenchmarkPubs} benchmark articles, that
-contain {\color{black}\numBenchmarks} different benchmarks, together these
-rank {\color{black}\numTools~} distinct software packages.
+ranks were extracted from the articles that met the criteria, and
+these were added to the set of training articles. The evaluated
+candidate articles that did not meet the ALB-criteria were incorporated
+into the set of background articles. This process was iterated and resulted in the identification of
+{\color{black}\numBenchmarkPubs} benchmark articles,
+containing {\color{black}\numBenchmarks} different benchmarks. Together these
+ranked {\color{black}\numTools~} distinct software packages.
 
@@
 have attempted to be as comprehensive as possible in terms of
-benchmark inclusion, as well as include comprehensive benchmarks. By
-which we mean studies that include all available software tools that
-address a biological problem.
+benchmark inclusion, as well as including comprehensive benchmarks (i.e.,
+studies that include all available software tools that
+address a specific biological problem).
 
@@
 ALB-criteria and contained data on both the accuracy and speed from
-their tests we extracted ranks for each tool. Many articles contained
+their tests, we extracted ranks for each tool. Many articles contained
 multiple benchmarks, in these cases we selected a range of these, the
@@
 and accuracy rank data for each tool. In line with rank-based
-statistics, the cases where tools were tied are resolved by using a
-midpoint rank (e.g. if tool 3 and 4 are tied, the rank 3.5 is used)
+statistics, the cases where tools were tied were resolved by using a
+midpoint rank (e.g., if tool 3 and 4 are tied, the rank 3.5 was used)
 \cite{Mann1947-re}. Each rank extraction was independently verified by
@@
 number of tools in the benchmark. For tools that were benchmarked
-multiple times with multiple metrics (e.g. BWA is evaluated in 6
+multiple times with multiple metrics (e.g., BWA was evaluated in 6
 different articles
 \cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
-a mean normalised rank is used to summarise the accuracy and speed performance. 
+a mean normalised rank was used to summarise the accuracy and speed performance. 
 Or, more formally:
@@
 For each tool we identified the corresponding publications in
-GoogleScholar, the total number of citations was recorded, the
-corresponding authors were also identified and if these had public
-GoogleScholar profiles we extracted their H-index and calculated a
+GoogleScholar; the total number of citations was recorded, the
+corresponding authors were also identified, and if they had public
+GoogleScholar profiles, we extracted their H-index and calculated a
 M-index ($\frac{H-index}{y}$) where ‘$y$’ is the number of years since
-their first publication. For the journals that each tool is published
-in we extracted the H5-index
-from GoogleScholar Metrics. The year of publication was also recorded for each
-tool. A “relative age” and “relative citations” was also computed for
+their first publication. The journal quality was estimated using the H5-index
+from GoogleScholar Metrics.
+
+The year of publication was also recorded for each
+tool. “Relative age” and “relative citations” were also computed for
 each tool. For each benchmark, software was ranked by year of first
@@
 Google was also employed to identify GitHub repositories.
-When a repository was matched with a tool, the number of "commits" and number of "contributors" was collected, when details of version numbers were provided, these were also harvested.
+When a repository was matched with a tool, the number of ``commits'' and number of ``contributors'' was collected, when details of version numbers were provided, these were also harvested.
 Version numbers are inconsistently used between groups, and may begin at either 0 or 1.
-To counter this issue we have added '1' to all versions less than '1', e.g.
-version 0.31 become 1.31.
+To counter this issue we have added `1' to all versions less than `1', for example, version 0.31 become 1.31.
 In addition, multiple point releases may be used e.g.
-'version 5.2.6', these have been mapped to the nearest decimal value '5.26'.
+`version 5.2.6', these have been mapped to the nearest decimal value `5.26'.
 
-\textbf{Statistical analysis:} For each tool we have manually collected up to 12 different
-statistics from GoogleScholar, GitHub and directly from literature describing tools (
-1. corresponding author’s H-index, 
+\textbf{Statistical analysis:} For each tool we manually collected up to 12 different
+statistics from GoogleScholar, GitHub and directly from literature describing tools 
+(1. corresponding author’s H-index, 
 2. corresponding author’s M-index, 
@@
 12. number of contributors to GitHub). 
-These have been evaluated in a pairwise fashion to
-produce Figure~\ref{fig:allfactors} A\&B, the R code for these is
-given in the GitHub repository (linked below).
+These were evaluated in a pairwise fashion to
+produce Figure~\ref{fig:allfactors} A\&B, the R code used to generate these is
+given in a GitHub repository (linked below).
 
@@
 For each benchmark of three or more tools, we extracted the published
-accuracy and speed ranks. In order to identify if there is an
+accuracy and speed ranks. In order to identify whether there was an
 enrichment of certain accuracy and speed pairings we constructed a
@@
 speed rank pairing was recorded. For each benchmark this procedure was
-repeated 10,000 times. These permuted rankings were normalised and
+repeated 1,000 times. These permuted rankings were normalised and
 compared to the real rankings to produce the ‘$\vartimes$’ points in
@@
 Figure~\ref{fig:speedaccuracy} is based upon Z-scores
-($Z=\frac{x-\bar{x}}{\sigma}$). For each cell in a $10\times 10$ grid
-a Z-score is computed to illustrate the abundance or lack of tools in
+($Z=\frac{x-\bar{x}}{s}$). For each cell in a $3\times 3$ grid
+a Z-score (and corresponding P-value is computed, either with the `pnorm' distribution function in R (Figure 2A) or empirically (Figure 2B)) is computed to illustrate the abundance or lack of tools in
 a cell relative to the permuted data.
@@
 
+This work was largely conducted on the traditional territory of K\=ai Tahu.
+
 %{\bf AND ETC....}

### 2021-03-12 manuscript/supplementary.tex
Major revision for manuscript and figures after comments from Mik.

@@
 
-\PaperTitle{Supplementary results for: A meta-analysis of bioinformatics software benchmarks reveals that publication-bias unduly influences software accuracy} % Article title
+\PaperTitle{Supplementary results for: A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy} % Article title
 
@@
   that matched a hand-selected list of keywords associated with
-  benchmarks were then scored and ranked with a “sum of bits”
+  benchmarks: \texttt{((bioinformatics OR (computational AND biology)) AND (algorithmic OR algorithms OR biotechnologies OR computational OR kernel OR methods OR procedure OR programs OR software OR technologies)) AND (accuracy OR analysis OR assessment OR benchmark OR benchmarking OR biases OR comparing OR comparison OR comparisons OR comparative OR comprehensive OR effectiveness OR estimation OR evaluation OR metrics OR efficiency OR performance OR perspective OR quality OR rated OR robust OR strengths OR suitable OR suitability OR superior OR survey OR weaknesses OR correctness OR correct OR evaluate OR competing OR competition) AND (complexity OR cputime OR runtime OR walltime OR duration OR elapsed OR fast OR faster OR perform OR performance OR slow OR slower OR speed OR time OR (computational AND resources))}.
+These
+  were then scored and ranked with a “sum of bits”
   score. High ranking articles were then inspected, those that met our
-  criteria were added to the training set, those that didn’t were
+  criteria were added to the training set, those that did not were
   added to the background set of articles.}
@@
 \begin{figure}[H] \centering
-\includegraphics[width=1.05\textwidth]{wordScores.pdf} \caption{The 40
+\includegraphics[width=0.99\textwidth]{wordScores.pdf} \caption{The 40
   highest and lowest scoring words that are associated with
@@
 
-\begin{figure}[H]
+\begin{figure}[htb!]
 \centering
@@
 \end{figure}
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[width=0.95\textwidth]{supplementary-distributions-permuted.pdf}
+\caption{The distributions for the permuted accuracy and speed values, where accuracy and speed ranks have been normalised to lie between 0 and 1.
+}
+\label{fig:metricDistributionsPerm}
+\end{figure}
+
+\clearpage
 
@@
 \caption{{\bf Citation metrics and software tool speeds/accuracies.} Heatmaps of normalised speed ranks and normalised accuracy
-  ranks, both $x$ and $y$ dimensions are discretised into a $10 \times
-  10$ matrix. The shading indicates different median citation-based
+  ranks, both $x$ and $y$ dimensions are discretised into a $3 \times
+  3$ matrix. The shading indicates different median citation-based
   feature scores (redder shade indicates a higher value).  The shading
@@
   Heatmaps of normalised speed ranks and normalised accuracy
-  ranks, both $x$ and $y$ dimensions are discretised into a $10 \times
-  10$ matrix. The shading indicates different median github-derived
+  ranks, both $x$ and $y$ dimensions are discretised into a $3 \times
+  3$ matrix. The shading indicates different median github-derived
   feature scores (redder shade indicates a higher value).  The shading

### 2021-03-12 bin/prettyPlot.R

@@ -12,6 +12,7 @@ library(gplots)
 #library("CombinePValue")
 library("hash")
 library("vioplot")
+library("randomForest")
 
 k <- 11
 my.cols <- rev(brewer.pal(k, "RdYlBu"))
@@ -165,7 +166,8 @@ dev.off()
 accPerms <-read.table("meanRankAccuracyPerms.tsv", header=T)
 spdPerms <-read.table(   "meanRankSpeedPerms.tsv", header=T)
 
-gridRes  <- 10; #10 x 10 grid
+#Figure 2
+gridRes  <- 3; #10 x 10 or 5 x 5 or 3 x 3 grid
 numPerms <- max(accPerms$permutation)+1
 
 #library(gplots)
@@ -174,12 +176,18 @@ h2dNorm     <- hist2d(d2005$accuracyRank,d2005$speedRank, show=FALSE, same.scale
 sumX  <- 0*h2dNorm$counts
 sumXX <- 0*h2dNorm$counts
 
-
-fastInacc <- vector(mode = "numeric", length = 4*numPerms)
-fastAcc   <- vector(mode = "numeric", length = 4*numPerms)
-slowInacc <- vector(mode = "numeric", length = 4*numPerms)
-slowAcc   <- vector(mode = "numeric", length = 4*numPerms)
-midBlock  <- vector(mode = "numeric", length = 4*numPerms)
+if(gridRes == 10){
+     multiplier <- 4
+}else if(gridRes == 5){
+     multiplier <- 1    
+}else if(gridRes == 3){
+     multiplier <- 1
+}    
+fastInacc <- vector(mode = "numeric", length = multiplier*numPerms)
+fastAcc   <- vector(mode = "numeric", length = multiplier*numPerms)
+slowInacc <- vector(mode = "numeric", length = multiplier*numPerms)
+slowAcc   <- vector(mode = "numeric", length = multiplier*numPerms)
+midBlock  <- vector(mode = "numeric", length = multiplier*numPerms)
 
 dNames <- c("mindex",  "hindex",  "relAge",  "H5", "speedRank", "citations", "relCites",   "yearPublished", "version", "commits", "contributors")
[…]

### 2021-03-12 bin/tsv2data.pl

@@ -349,7 +349,7 @@ sub permutationsRelativeRanks {
 	    }
 	    
 	    if(defined($methodInfo->{$meth}{'citations'} )){
-		$citations{$meth} = $methodInfo->{$meth}{'citations'}        if (isNumeric($methodInfo->{$meth}{'citations'})); 
+		$citations{$meth} = $methodInfo->{$meth}{'citations'} if (isNumeric($methodInfo->{$meth}{'citations'})); 
 	    }
 	    
 	    #normaliseH returns a hash

### 2021-03-18 manuscript/supplementary.tex
Updated manuscript title.

@@
 
-\PaperTitle{Supplementary results for: A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy} % Article title
+\PaperTitle{Supplementary results for: Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software} % Article title
 

### 2021-08-06 manuscript/manuscript-speed-accuracy.tex

@@
 
-\PaperTitle{A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy} 
-%Analysis of sixty software benchmarks shows the benefits of good software practice 
-%Assessing the state of the bioinformatic software landscape with a meta-analysis of benchmarks
-%Sustained software development, not citation metrics, is indicative of accurate bioinformatic software 
+\PaperTitle{Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software}
+
+%1 A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy
+%2 Analysis of sixty software benchmarks shows the benefits of good software practice 
+%3 Assessing the state of the bioinformatic software landscape with a meta-analysis of benchmarks
+%4 Sustained software development, not citation metrics, is indicative of accurate bioinformatic software 
 %
@@
 administered by the Royal Society Te Ap\=arangi, 
-PPG, AG and MB acknowledge support from a Data Science Programmes grant
+PPG, AG and MAB acknowledge support from a Data Science Programmes grant
 (UOAX1932).

### 2021-08-08 manuscript/manuscript-speed-accuracy.tex
Updated README. Re-compiled manuscript & supplementary materials.

@@
 \graphicspath{{../figures/}}
+\usepackage{xr}
+\externaldocument{supplementary}
+
 \setlength{\abovecaptionskip}{15pt plus 3pt minus 2pt}
@@
 
-Data on the
-number of updates to software tools from GitHub such as the number of versions,
-commits and contributors was significantly correlated with 
-software accuracy (respective Spearman’s rhos = {\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg corrected P-values = {\color{black}0.029, 0.042, 0.063}), Figure~\ref{fig:allfactors}B.
-The significance if these features was further confirmed with a permutation test (Figure~\ref{fig:allfactors}B).
-These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A \& Supplementary Figure ????).
-We also found that reputation metrics such as citations, author and journal H-indices, and the age
-of tools were generally \textbf{not} correlated with either tool
-accuracy or speed (Figure~\ref{fig:allfactors}A\&B). 
+Data on the number of updates to software tools from GitHub such as
+the number of versions, commits and contributors was significantly
+correlated with software accuracy (respective Spearman’s rhos =
+{\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg
+corrected P-values = {\color{black}0.029, 0.042, 0.063}),
+Figure~\ref{fig:allfactors}B.  The significance of these features was
+further confirmed with a permutation test
+(Figure~\ref{fig:allfactors}B).  These features were not correlated
+with speed however (see Figure~\ref{fig:allfactors}A \& Supplementary
+Figures~\ref{fig:metricPairs} and~\ref{fig:S2}).  We also found that
+reputation metrics such as citations, author and journal H-indices,
+and the age of tools were generally \textbf{not} correlated with
+either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).
 %The strongest

### 2021-08-12 bin/prettyPlot.R
Adding additional github features per referee reqs.

@@ -159,6 +159,50 @@ lines(c(-100,100),c(0,0))
 dev.off()
 
 
+#####################
+#Reviewer 2 requested plots:
+#plot each software package (potentially sized by the number of studies on which a software tool was included)
+pdf(file=    "../figures/numberBenchmarksPerToolBarplot.pdf", width = 100,  height = 5)
+op<-par(mfrow=c(1,1),cex=1.0,las=2, mar = c(12,4,4,4) + .1)
+bp<-barplot(d2005$numTests,names=d2005$method, ylab="Number of benchmarks",ylim=c(0,26),main="Number of benchmarks per software tool")
+dev.off()
+
+#Presumably the data are not fully complete (i.e., some scholar profiles couldn't be identified, etc). A table within either the methods or results breaking out the completeness of the records would be helpful.
+
+naFeatureCounts<-matrix(0, length(dNames), 2)
+for(i in 1:length(dNames)){      
+      naFeatureCounts[i,2] <- sum(is.na(d2005[,dNames[i] == colnames(d2005)]))
+      naFeatureCounts[i,1] <- length(d2005[,dNames[i] == colnames(d2005)]) - naFeatureCounts[i,2]      
+}
+rownames(naFeatureCounts) <- pNames
+colnames(naFeatureCounts) <- c("Known", "NA")
+
+ixF <- sort(naFeatureCounts[,1], index.return=T, decreasing=T)$ix
+
+pdf(file=    "../figures/numberRealValueFeaturesBarplot.pdf", width = 7,  height = 5)
+op<-par(mfrow=c(1,1),cex=1.0,las=2, mar = c(6,4,4,4) + .1)
+barplot(t(naFeatureCounts[ixF,]), col=c("plum3", "plum1"), xlab = "", ylab = "Count", ylim=c(0,500), xlim = c(0,length(dNames)+6), width = 1, main="Data completeness")
+legend("bottomright", 
+ legend = c("Real val.", "NA"),
+ fill = c("plum3", "plum1"),
+ title = "Data")
+for(i in 1:length(dNames)){      
+      text( i*(length(dNames)+2)/length(dNames) , 30, naFeatureCounts[ixF[i],1], srt=90 )
+}
+dev.off()
+
+#Reviewer 3 requested plots:
[…]

### 2021-08-12 bin/tsv2data.pl

@@ -6,7 +6,7 @@ use Data::Dumper;
 
 ###################################
 #Parse tool info (e.g. age, impact, etc.):
-my @methodInfoKeys = qw(yearPublished H5 citations hindex mindex version commits contributors); #IF 
+my @methodInfoKeys = qw(yearPublished H5 citations hindex mindex version commits contributors issues issuesFracOpen pullrequests forks); #IF 
 my $methodInfoPtr = extractMethodInfo("speed-vs-accuracy-toolInfo2005-2020.tsv", "journalInfo2020.tsv"); #, \%methodInfo); 
 my %methodInfo = %{$methodInfoPtr};
     

### 2021-08-12 bin/prettyPlot.R
Updated perl parser & R vis code to cater for additional features, issues, pull requests, forks etc.

@@ -30,12 +30,12 @@ dr2005<-read.table( "rawRankSpeedData2005-2020.tsv", header=T)
 reg1  <- lm(d2005$accuracyRank ~ d2005$speedRank)
 summary(reg1)
 
-regM <- lm(accuracyRank ~ speedRank + H5 + citations + hindex + mindex + relAge + relCites + version + commits + contributors, data=d2005, na.action=na.roughfix)
+regM <- lm(accuracyRank ~ speedRank + H5 + citations + hindex + mindex + relAge + relCites + version + commits + contributors + issues + issuesFracOpen + pullrequests + forks, data=d2005, na.action=na.roughfix)
 summary(regM)
 
 ####PCA:
-xx<-cbind(d2005$accuracyRank, d2005$speedRank, d2005$H5, d2005$citations, d2005$hindex, d2005$mindex, d2005$relAge, d2005$relCites,d2005$yearPublished, d2005$version, d2005$commits, d2005$contributors)
-colnames(xx)<-c("accuracy","speed","JH5","citations","hindex","mindex","relAge","relCites","age", "version", "commits", "contributors")
+xx<-cbind(d2005$accuracyRank, d2005$speedRank, d2005$H5, d2005$citations, d2005$hindex, d2005$mindex, d2005$relAge, d2005$relCites,d2005$yearPublished, d2005$version, d2005$commits, d2005$contributors, d2005$issues, d2005$issuesFracOpen, d2005$pullrequests, d2005$forks)
+colnames(xx)<-c("accuracy","speed","JH5","citations","hindex","mindex","relAge","relCites","age", "version", "commits", "contributors", "issues", "issuesFracOpen", "pullrequests", "forks")
 pca<-prcomp(na.omit(xx), center = TRUE,scale=TRUE)
 pca
 summary(pca)
@@ -62,8 +62,8 @@ summary(pca)
 #FIGURE 1 A & B
 #All vs All Spearman Heatmap
 
-dNames <- c("H5", "relAge",    "yearPublished", "citations", "relCites",   "mindex",  "hindex",  "accuracyRank",  "speedRank", "version", "commits", "contributors" )
-pNames <- c("JH5", "Rel. age", "Year",          "Citations", "Rel. cites", "M-index", "H-index", "Accuracy",      "Speed",     "version", "commits", "contributors")
[…]

### 2021-08-12 bin/tsv2data.pl

@@ -38,7 +38,7 @@ printResults("meanRankSpeedData.tsv", "meanRankAccuracyPerms.tsv", "meanRankSpee
 
 ###################################
 #Make all the figures, do some statistics: 
-system("R CMD BATCH --no-save ../bin/prettyPlot.R");
+#system("R CMD BATCH --no-save ../bin/prettyPlot.R");
 
 exit(0);
 
@@ -68,22 +68,26 @@ sub extractMethodInfo {
     
     
 # head -n 1 speed-vs-accuracy-toolInfo2005-2020.tsv | tr "\t" "\n" | nl
-#      1	tool
-#      2	yearPublished
-#      3	journal
-#      4	impactFactor(2017)
-#      5	Journal H5-index(2017)
-#      6	totalCitations(2017)
-#      7	totalCitations(2020)
-#      8	H-index: (Corresponding author)(2017)
-#      9	M-index (Corresponding author)(2017)
-#     10	H (2020)
-#     11	M (2020)
-#     12	Versions
-#     13	Commits (Github)
-#     14	Contributers (Github)
-#     15	Github repo
-#     16	fullCite
+    #  1	tool
+    #  2	yearPublished
+    #  3	journal
+    #  4	impactFactor(2017)
+    #  5	Journal H5-index(2017)
+    #  6	totalCitations(2017)
+    #  7	totalCitations(2020)
+    #  8	H-index: (Corresponding author)(2017)
+    #  9	M-index (Corresponding author)(2017)
+    # 10	H (2020)
+    # 11	M (2020)
+    # 12	Versions
+    # 13	Commits (Github)
+    # 14	Contributers (Github)
+    # 15	Issues Open
+    # 16	Issues Closed
+    # 17	Pull requests
+    # 18	Forks
+    # 19	Github repo
+    # 20	fullCite
 
 
     #For each tool collect:
@@ -188,6 +192,38 @@ sub extractMethodInfo {
 		$methodInfo{$in[0]}{'contributors'}='NA';
 		$methodInfo{$in[0]}{'contributors'}= $in[13] if (isNumeric($in[13])); 
 	    }
+
+	    #Issues in github
+	    if(defined($in[14]) && defined($in[15])){
+		$methodInfo{$in[0]}{'issues'}='NA';
[…]

### 2021-08-13 manuscript/supplementary.tex
Updated README. Re-compiled manuscript & supplementary materials.
Working additional features into plots and analyses.

@@
 \includegraphics[width=0.99\textwidth]{wordScores.pdf} \caption{The 40
-  highest and lowest scoring words that are associated with
+  highest ({\bf right}) and lowest ({\bf left}) scoring words that are associated with
   bioinformatic benchmark articles from the training benchmark articles,
@@
 
-\section*{Data mining} % The \section*{} command stops section numbering
+\section*{Data visualisations} % The \section*{} command stops section numbering
 
@@
 \includegraphics[width=0.95\textwidth]{supplementary-figures-small.pdf}
-\caption{The distributions for the measures we expected to be
+\caption{The distributions for the measures that have been proposed to be
 predictive of software quality used in this study. These
-are, reading from left to right, top to bottom: Accuracy -- the mean
-normalised accuracy rank for each benchmarked method; Speed -- the
-mean normalised speed rank for each benchmarked method;
-JH5 -- the Journal H5 index to the highest impact journal that has published
-a manuscript describing a method, data from GoogleScholar 2020 Metrics;
+are, reading from left to right, top to bottom:
+Accuracy -- the mean normalised accuracy rank for each benchmarked method;
+Speed -- the mean normalised speed rank for each benchmarked method;
+JH5 -- the Journal H5 index to the highest impact journal that has published a manuscript describing a method, data from GoogleScholar 2020 Metrics;
 Cites -- the number of citations to the most cited manuscript describing a method, data from GoogleScholar;
-H-index -- the H-index for the highest profile corresponding author from the
-manuscripts describing a method, data from GoogleScholar User Profiles;
-M-index -- the M-index (H-index/(\#years since first publication)) for the
-highest profile corresponding author from the manuscripts describing a method,
-data from GoogleScholar User Profiles;
-Relative ages for different tools, for each benchmark tools were ranked based upon publication dates, these ranks were normalised to lie between 0 and 1.
+H-index -- the H-index for the highest profile corresponding author from the manuscripts describing a method, data from GoogleScholar User Profiles;
+M-index -- the M-index (H-index/(\#years since first publication)) for the highest profile corresponding author from the manuscripts describing a method, data from GoogleScholar User Profiles;
+Relative ages -- for different tools, for each benchmark tools were ranked based upon publication dates, these ranks were normalised to lie between 0 and 1.
+Relative \#citations -- within each benchmark, what is the relative age of each tool, this corrects for older research fields (e.g.
+sequence alignment) vs newer research fields (e.g.
+single-cell sequencing);
+Year published -- the first year each software tool was published;
+Version -- the most resent version number for each tool;
+Commits -- the number of commits to Github;
+Contributors -- the number of contributors to a Github repository;
+\#Issues -- the number of open and closed issues for each Github repository;
+\%Open issues -- the pecentage of issues that remain open for each Github repository;
+\#Pull requests -- the number of pull requests to a Github repository;
+\#Forks -- the number of times a Github repository has been forked;
 }
@@
 \includegraphics[width=0.95\textwidth]{supplementary-distributions-permuted.pdf}
-\caption{The distributions for the permuted accuracy and speed values, where accuracy and speed ranks have been normalised to lie between 0 and 1.
+\caption{The distributions for the permuted accuracy ({\bf left}) and speed ({\bf right}) values, where accuracy and speed ranks have been normalised to lie between 0 and 1.
 }
@@
 \includegraphics[width=0.95\textwidth]{supplementary-figures-pairs.pdf}
-\caption{}
+\caption{Dot plots for each feature relative to every other feature
+  (accuracy, speed, JH5, citations, H-index, M-index, relAge,
+  relCites, yearPublished, version, commits, contributors,
+  ).
+
+}
 \label{fig:metricPairs}
@@
 \includegraphics[width=0.45\textwidth]{spearmanBarplotSpeed.pdf}
-\caption{The correlation between method accuracy (on the left) and
-  method speed (on the right) and measures we expected to be
+\caption{The correlation between method accuracy ({\bf left}) and
+  method speed ({\bf right}) and measures we expected to be
   predictive of software quality. E.g. author reputation measures
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.45\textwidth]{hindex-SpeedVsAccuracy-heatmap.pdf}
-\includegraphics[width=0.45\textwidth]{cites-SpeedVsAccuracy-heatmap.pdf}
-\caption{{\bf Citation metrics and software tool speeds/accuracies.} Heatmaps of normalised speed ranks and normalised accuracy
-  ranks, both $x$ and $y$ dimensions are discretised into a $3 \times
-  3$ matrix. The shading indicates different median citation-based
-  feature scores (redder shade indicates a higher value).  The shading
-  indicate median H-indices ({\bf left}) and $log_{10}(citations)$
-  ({\bf right}). }
-\label{fig:heatmaps}
-\end{figure}
+%% \begin{figure}[H]
+%% \centering
+%% \includegraphics[width=0.45\textwidth]{hindex-SpeedVsAccuracy-heatmap.pdf}
+%% \includegraphics[width=0.45\textwidth]{cites-SpeedVsAccuracy-heatmap.pdf}
+%% \caption{{\bf Citation metrics and software tool speeds/accuracies.} Heatmaps of normalised speed ranks and normalised accuracy
+%%   ranks, both $x$ and $y$ dimensions are discretised into a $3 \times
+%%   3$ matrix. The shading indicates different median citation-based
+%%   feature scores (redder shade indicates a higher value).  The shading
+%%   indicate median H-indices ({\bf left}) and $log_{10}(citations)$
+%%   ({\bf right}). }
+%% \label{fig:heatmaps}
+%% \end{figure}
 
@@
 
+%% \begin{figure}[H]
+%% \centering
+%% \includegraphics[width=0.45\textwidth]{commits-SpeedVsAccuracy-heatmap.pdf}
+%% \includegraphics[width=0.45\textwidth]{contributors-SpeedVsAccuracy-heatmap.pdf}
+%% \caption{{\bf Version control use and software tool speeds/accuracies.}
+%%   Heatmaps of normalised speed ranks and normalised accuracy
+%%   ranks, both $x$ and $y$ dimensions are discretised into a $3 \times
+%%   3$ matrix. The shading indicates different median github-derived
+%%   feature scores (redder shade indicates a higher value).  The shading
+%%   indicate the number of commits ({\bf left}) and the number of code contributors
+%%   ({\bf right}).
+%%  }
+%% \label{fig:heatmapsCommits}
+%% \end{figure}
+
+\clearpage
+\newpage
+
 \begin{figure}[H]
 \centering
-\includegraphics[width=0.45\textwidth]{commits-SpeedVsAccuracy-heatmap.pdf}
-\includegraphics[width=0.45\textwidth]{contributors-SpeedVsAccuracy-heatmap.pdf}
-\caption{{\bf Version control use and software tool speeds/accuracies.}
-  Heatmaps of normalised speed ranks and normalised accuracy
-  ranks, both $x$ and $y$ dimensions are discretised into a $3 \times
-  3$ matrix. The shading indicates different median github-derived
-  feature scores (redder shade indicates a higher value).  The shading
-  indicate the number of commits ({\bf left}) and the number of code contributors
-  ({\bf right}).
+\includegraphics[width=0.99\textwidth]{numberBenchmarksPerToolBarplot.pdf}\\
+\includegraphics[width=0.99\textwidth]{numberRealValueFeaturesBarplot.pdf}
+\caption{ {\bf Top:} The number of benchmarks each software tool has appeared in. {\bf Bottom:}
+  Plot of the amount missing/known data for each of the software features used in this analysis. 
  }
-\label{fig:heatmapsCommits}
+\label{fig:numBenchmarks}
 \end{figure}
+
+\clearpage
+\newpage
 

### 2021-08-13 bin/prettyPlot.R

@@ -42,10 +42,9 @@ summary(pca)
 
 ####Random Forest:
 # library(randomForest)
-
 # ind <- sample(2,nrow(d2005),replace=TRUE,prob=c(0.7,0.3))
-# trainData <- d2005[ind==1,c(2,3,6:14)]
-# testData  <- d2005[ind==2,c(2,3,6:14)]
+# trainData <- d2005[ind==1,c(2,3,6:19)]
+# testData  <- d2005[ind==2,c(2,3,6:19)]
 # acc.rf <- randomForest(accuracyRank ~ ., data=trainData,ntree=1000,proximity=TRUE, na.action=na.roughfix)
 # round(importance(acc.rf,type=1), 2)
 # par(mfrow=c(1,1))
@@ -55,7 +54,6 @@ summary(pca)
 # abline(reg1,col="red")
 # corTest<-cor.test( testData$accuracyRank,predTest)
 # text(0.05,0.4,paste("R2: ", round(corTest$estimate,2), "\np-value: ", round(corTest$p.value,2), "\nN: ", length(predTest)), col="red", pos=4)
-
 # varImpPlot(acc.rf,type=1)
 
 ######################################################################
@@ -91,7 +89,7 @@ for(i in 1:N){
 }
 
 #FIGURE 1A
-pdf(file=    "../figures/spearmanHeatmap.pdf", width = 7,  height = 6)
+pdf(file=    "../figures/spearmanHeatmap.pdf", width = 9,  height = 8)
 par(mar = c(8,4,4,4) + .1) #c(bottom, left, top, right). default: c(5, 4, 4, 2) + 0.1
 heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black", col=rev(redblue(40)), density.info="none", trace="none", dendrogram=c("column"), symm=F,symkey=T,symbreaks=T, scale="none", key.title = "", srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), breaks=(-20:20)/20,
 margins = c(8, 8), cexRow=1.5, cexCol=1.5,font=2)
@@ -120,6 +118,8 @@ forksA          <-cor.test(d2005$accuracyRank, as.numeric(d2005$forks),         
[…]

### 2021-08-16 manuscript/manuscript-speed-accuracy.tex
Updating/adding figures, supplementary materials and minor manuscript edits.

@@
-\documentclass[fleqn,10pt]{SelfArx} % Document font size and equations flushed left
+\documentclass{bmcart}
+%\documentclass[fleqn,10pt]{SelfArx} % Document font size and equations flushed left
 
+\usepackage[utf8]{inputenc} %unicode support
 \usepackage{float}
@@
 \usepackage[hyphens]{url}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{xcolor}
@@
 
-\newcommand{\beginsupplement}{%
-        \setcounter{table}{0}
-        \renewcommand{\thetable}{S\arabic{table}}%
-        \setcounter{figure}{0}
-        \renewcommand{\thefigure}{S\arabic{figure}}%
-     }
+%% \definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
+%% \definecolor{color2}{RGB}{200,200,200} % Color of the boxes behind the abstract and headings
+%% \definecolor{color3}{RGB}{5,5,5} % Color of the boxes behind the abstract and headings
 
-\definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
-\definecolor{color2}{RGB}{200,200,200} % Color of the boxes behind the abstract and headings
-\definecolor{color3}{RGB}{5,5,5} % Color of the boxes behind the abstract and headings
+%\JournalInfo{.} % Journal information ``Journal, Vol. XXI, No. 1, 1-5, 2015''
+%\Archive{ } % Additional notes (e.g. copyright, DOI, review/research article)
 
-\JournalInfo{.} % Journal information ``Journal, Vol. XXI, No. 1, 1-5, 2015''
-\Archive{ } % Additional notes (e.g. copyright, DOI, review/research article)
-
+%%% Put your definitions there:
+\startlocaldefs
 \def\numTools{499}
@@
 %cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l
+\endlocaldefs
 
-\PaperTitle{Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software}
+%%% Begin BMC format...
+\begin{document}
 
-%1 A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy
-%2 Analysis of sixty software benchmarks shows the benefits of good software practice 
-%3 Assessing the state of the bioinformatic software landscape with a meta-analysis of benchmarks
-%4 Sustained software development, not citation metrics, is indicative of accurate bioinformatic software 
-%
+%%% Start of article front matter
+\begin{frontmatter}
 
-\Authors{Paul P.
+\begin{fmbox}
+\dochead{Research}
+
+%\PaperTitle{Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software}
+\title{Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software}
+
+%% \Authors{Paul P.
 Gardner\textsuperscript{1,2}*, James M.
@@
 Umu\textsuperscript{6}, Aleksandra Pawlik\textsuperscript{7},
-Alex Gavryushkin\textsuperscript{8}, Michael A Black\textsuperscript{1}
-} % Authors
+%% Alex Gavryushkin\textsuperscript{8}, Michael A Black\textsuperscript{1}
+%% } % Authors
 
-\affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
-\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
-\affiliation{\textsuperscript{3}\textit{Department of Civil and Natural Resources Engineering, University of Canterbury, Christchurch, New Zealand.}}
-\affiliation{\textsuperscript{4}\textit{Parasites and Microbes, Wellcome Sanger Institute, Wellcome Genome Campus, Hinxton,
-8 Cambridgeshire, CB10 1RQ, UK.}}
-\affiliation{\textsuperscript{5}\textit{Research Group for Genomic Epidemiology, National Food Institute, Technical University of Denmark, Kongens Lyngby, Denmark.}}
-\affiliation{\textsuperscript{6}\textit{Department of Research, Cancer Registry of Norway, Oslo, Norway.}} % Author affiliation
-\affiliation{\textsuperscript{7}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
-\affiliation{\textsuperscript{8}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+%james.paterson@canterbury.ac.nz, srm88@cam.ac.uk, fatgho@food.dtu.dk, sinan.ugur.umu@kreftregisteret.no, PawlikA@landcareresearch.co.nz, alex@biods.org, mik.black@otago.ac.nz
 
-\affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author
+%Authors:
+\author[
+  addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
+  corref={aff1},                       % id of corresponding address, if any
+  email={paul.gardner@otago.ac.nz}   % email address
+]{\inits{P.P.}\fnm{Paul P.} \snm{Gardner}}
+\author[
+  addressref={aff3},
+  email={james.paterson@canterbury.ac.nz}
+]{\inits{J.M.}\fnm{James M.} \snm{Paterson}}
+\author[
+  addressref={aff4},
+  email={srm88@cam.ac.uk}
+]{\inits{S.}\fnm{Stephanie} \snm{McGimpsey}}
+\author[
+  addressref={aff5},
+  email={fatgho@food.dtu.dk}
+]{\inits{F.}\fnm{Fatemeh} \snm{Ashari-Ghomi}}
+\author[
+  addressref={aff6},
+  email={sinan.ugur.umu@kreftregisteret.no}
+]{\inits{S.U.}\fnm{Sinan U.} \snm{Umu}}
+\author[
+  addressref={aff7},
+  email={PawlikA@landcareresearch.co.nz}
+]{\inits{A.}\fnm{Aleksandra} \snm{Pawlik}}
+\author[
+  addressref={aff8},
+  email={alex@biods.org}
+]{\inits{A.}\fnm{Alex} \snm{Gavryushkin}}
+\author[
+  addressref={aff1},
+  email={mik.black@otago.ac.nz}
+]{\inits{M.A.}\fnm{Michael A.} \snm{Black}}
 
-\Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
-\newcommand{\keywordname}{Keywords} % Defines the keywords heading name
+%Affiliations:
+%% \affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+%% \affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
+%% \affiliation{\textsuperscript{3}\textit{Department of Civil and Natural Resources Engineering, University of Canterbury, Christchurch, New Zealand.}}
+%% \affiliation{\textsuperscript{4}\textit{Parasites and Microbes, Wellcome Sanger Institute, Wellcome Genome Campus, Hinxton,
+%% 8 Cambridgeshire, CB10 1RQ, UK.}}
+%% \affiliation{\textsuperscript{5}\textit{Research Group for Genomic Epidemiology, National Food Institute, Technical University of Denmark, Kongens Lyngby, Denmark.}}
+%% \affiliation{\textsuperscript{6}\textit{Department of Research, Cancer Registry of Norway, Oslo, Norway.}} % Author affiliation
+%% \affiliation{\textsuperscript{7}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
+%% \affiliation{\textsuperscript{8}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation
+
+%% \affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author
+
+%% \Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
+%% \newcommand{\keywordname}{Keywords} % Defines the keywords heading name
+
+\address[id=aff1]{%                       % unique id
+  \orgdiv{Department of Biochemistry,},   % department, if any
+  \orgname{University of Otago},          % university, etc
+  \city{Dunedin},                         % city
+  \cny{New Zealand}                       % country
+}
+\address[id=aff2]{%
+  \orgdiv{Biomolecular Interaction Centre},
+  \orgname{University of Canterbury},
+  %\street{},
+  %\postcode{}
+  \city{Christchurch},
+  \cny{New Zealand}
+}
+\address[id=aff3]{%
+  \orgdiv{Department of Civil and Natural Resources Engineering},
+  \orgname{University of Canterbury},
+  \city{Christchurch},
+  \cny{New Zealand}
+}
+\address[id=aff4]{%
+  \orgdiv{Parasites and Microbes},
+  \orgname{Wellcome Sanger Institute},
+  \city{Hinxton},
+  \cny{UK}
+}
+\address[id=aff5]{%
+  \orgdiv{Research Group for Genomic Epidemiology},
+  \orgname{Technical University of Denmark},
+  \city{Kongens Lyngby},
+  \cny{Denmark}
+}
+\address[id=aff6]{%
+  \orgdiv{Department of Research},
+  \orgname{Cancer Registry of Norway},
+  \city{Oslo},
+  \cny{Norway}
+}
+\address[id=aff7]{%
+%  \orgdiv{Department of Research},
+  \orgname{Manaaki Whenua - Landcare Research},
+  \city{Lincoln},
+  \cny{New Zealand}
+}
+\address[id=aff8]{%                       
+  \orgdiv{Department of Computer Science},   
+  \orgname{University of Otago},          
+  \city{Dunedin},                         
+  \cny{New Zealand}                       
+}
+
+\end{fmbox}% comment this for two column layout
 
@@
 
-%\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
-\Abstract{Computational biology provides widely used and powerful
-  software tools for testing and making inferences about biological
-  data. In the face of rapidly increasing volumes of data, heuristic
-  methods that trade software speed for accuracy may
-  be employed.  We are have studied these trade-offs using the results of a large number of independent software benchmarks, 
-  and evaluated whether external factors are indicative of accurate
+%\Abstract{
+\begin{abstractbox}
+  \begin{abstract} % abstract
+    \parttitle{Background:} Computational biology provides widely used and
+    powerful software tools for testing and making inferences about
+    biological data. In the face of rapidly increasing volumes of
+    data, heuristic methods that trade software speed for accuracy may
+    be employed.  We are have studied these trade-offs using the
+    results of a large number of independent software benchmarks, and
+    evaluated whether external factors are indicative of accurate
     software.
 
-We have extracted accuracy and speed ranks from independent benchmarks of
-different bioinformatic software tools, and evaluated whether the
-speed, author reputation, journal impact, recency and developer efforts are indicative of accuracy.
+    \parttitle{Method:} We have extracted accuracy and speed ranks from
+    independent benchmarks of different bioinformatic software tools,
+    and evaluated whether the speed, author reputation, journal
+    impact, recency and developer efforts are indicative of accuracy.
 
-We found that software speed, author reputation, journal impact, 
-number of citations and age are all unreliable predictors of software
-accuracy. This is unfortunate because citations, author and journal
-reputation are frequently cited reasons for selecting software
-tools. However, GitHub-derived records and high version numbers show that 
-the accurate bioinformatic software tools
-are generally the product of many improvements over time, often from
-multiple developers.
+    \parttitle{Results:} We found that software speed, author reputation,
+    journal impact, number of citations and age are all unreliable
+    predictors of software accuracy. This is unfortunate because
+    citations, author and journal reputation are frequently cited
+    reasons for selecting software tools. However, GitHub-derived
+    records and high version numbers show that the accurate
+    bioinformatic software tools are generally the product of many
+    improvements over time, often from multiple developers.
 
-We also find
-that the field of bioinformatics has a large excess of slow and inaccurate software
-tools, and this is consistent across many sub-disciplines. Meanwhile, there are few tools that
-are middle-of-road in terms of accuracy and speed trade-offs. We
-hypothesise that a form of publication-bias influences the
-publication and development of bioinformatic software. In other
-words, software that is intermediate in terms of both speed and accuracy may be difficult to publish - possibly due to author,
-editor and reviewer practices. This leaves an unfortunate hole in the
-literature as the ideal tools may fall into this gap. For example, high accuracy tools are not always useful if years of CPU time are required, 
-while high speed is not useful if the results are also inaccurate.  }
+    We also find that the field of bioinformatics has a
+    large excess of slow and inaccurate software tools, and this is
+    consistent across many sub-disciplines. Meanwhile, there are few
+    tools that are middle-of-road in terms of accuracy and speed
+    trade-offs. We hypothesise that a form of publication-bias
+    influences the publication and development of bioinformatic
+    software. In other words, software that is intermediate in terms
+    of both speed and accuracy may be difficult to publish - possibly
+    due to author, editor and reviewer practices. This leaves an
+    unfortunate hole in the literature as the ideal tools may fall
+    into this gap. For example, high accuracy tools are not always
+    useful if years of CPU time are required, while high speed is not
+    useful if the results are also inaccurate.
+%}
+\end{abstract}
 
-\begin{document}
+%% \begin{keyword}
+%% \kwd{sample}
+%% \kwd{article}
+%% \kwd{author}
+%% \end{keyword}
 
-\flushbottom % Makes all text pages the same height
-\maketitle % Print the title and abstract box
+\end{abstractbox}
+\end{frontmatter}
+
+%\begin{document}
+%\flushbottom % Makes all text pages the same height
+%\maketitle % Print the title and abstract box
 %\tableofcontents % Print the contents section
-
-\thispagestyle{empty} % Removes page numbering from the first page
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-%\onecolumn
-%\beginsupplement
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-%%REFERENCES::
-%reference Stefan Buchka$^{1}$, Rory Wilson$^2$, Alexander Hapfelmeier$^3$,\\  Paul Gardner$^4$,  Anne-Laure Boulesteix PAPER!!!
-
-%% A comprehensive analysis of the usability and archival stability of omics computational tools and resources
-%% https://www.biorxiv.org/content/10.1101/452532v1
-
-%% Tools and best practices for data processing in allelic expression analysis
-%% https://genomebiology.biomedcentral.com/articles/10.1186/s13059-015-0762-6
-
-%% Towards reproducible, transparent, and systematic benchmarking of omics computational tools
-%% https://osf.io/p8yd9
-
-%% Essential guidelines for computational method benchmarking
-%% https://arxiv.org/abs/1812.00661
+%\thispagestyle{empty} % Removes page numbering from the first page
 
@@
 software tool does not necessarily mean that it is accurate or
-computationally efficient, instead usability, ease of installation,
+computationally efficient{\color{red};} instead usability, ease of installation,
 operating system support or other indirect factors may play a greater role in a software
@@
 The gold-standard for determining accuracy is for researchers independent
-of individual tool development to conduct benchmarking studies, these benchmarks can serve a useful role in
+of individual tool development to conduct benchmarking studies{\color{red};} these benchmarks can serve a useful role in
 reducing the over-optimistic reporting of software accuracy
 \cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
-\cite{Norel2011-cq,Buchka:2021}. Benchmarking typically involves the use a number of
+\cite{Norel2011-cq,Buchka:2021a}. Benchmarking typically involves the use a number of
 positive and negative control datasets, so that predictions from different software tools can be
@@
 \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
-impact:} high profile journals are run by editors and reviewers who
+impact:} {\color{red}many believe that} high profile journals are run by editors and reviewers who
 carefully select and curate the best manuscripts. Therefore, high
@@
 \cite{Farrar2007-ky,Dematte2010-ph};
-However, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
+{\color{red}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
 rapid and sensitive pre-filtering
@@
 changes and versions is available for these
-\cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.
+\cite{ray2014large,Dozmorov:2018,Mangul:2019}.
 
@@
 \begin{figure*}[htb!]
-\includegraphics[width=\textwidth]{figure1.pdf}
+\includegraphics[width=0.9\textwidth]{figure1.pdf}
 \caption{\textbf{A.} A heatmap indicating the relationships between
@@
 For each of the publications describing these tools, we have (where
-possible) collected the journal's H5-index (\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
-Scholar Metrics}), the maximum H-index and
+possible) collected the journal's H5-index (
+%\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{
+Google Scholar Metrics
+%}
+), the maximum H-index and
 corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
@@
 (Figure~\ref{fig:allfactors}B).  These features were not correlated
-with speed however (see Figure~\ref{fig:allfactors}A \& Supplementary
-Figures~\ref{fig:metricPairs} and~\ref{fig:S2}).  We also found that
+with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Supplementary
+Figures~S5 and~S6}%\ref{fig:S2}
+).  We also found that
 reputation metrics such as citations, author and journal H-indices,
@@
 either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).
+
 %The strongest
@@
 \centering
-\includegraphics[width=0.75\textwidth]{figure2.pdf}
+\includegraphics[width=0.95\textwidth]{figure2.pdf}
 \caption{\textbf{A.} A heatmap indicating the relative paucity or abundance of
@@
 that is not highly ranked on speed or accuracy. If correct, this may
-have unfortunate consequences as these tools may never-the-less have further uses.
+have unfortunate consequences as these tools may {\color{red}nevertheless} have further uses.
 %We also found that slow and inaccurate
@@
 computational biology benchmarks that satisfy Boulesteix’s
-(ALB) three criteria for a “neutral comparison study”
+(ALB) three criteria for a ``neutral comparison study''
 \cite{Boulesteix2013-vb}. Firstly, the main focus of the article is
 the comparison and \textbf{not} the introduction of a new
-tool. Secondly, the authors should be reasonably neutral, which means
+tool {{\color{red}as these can be biased} \cite{Buchka:2021a}. Secondly, the authors should be reasonably neutral, which means
 that the authors should not generally have been involved in the
@@
 ALB-criteria and contained data on both the accuracy and speed from
-their tests, we extracted ranks for each tool. Many articles contained
-multiple benchmarks, in these cases we selected a range of these, the
+their tests, we extracted ranks for each tool.
+{\color{red}Until published datasets are made available in consistent, machine-readable formats this step is necessarily a manual process -- ranks were extracted from a mixture of manuscript figures, tables and supplementary materials, each data source is documented in Supplementary Table 1.
+In addition, a variety of accuracy metrics are reported e.g.
+``accuracy'', ``AUROC'', ``F-measure'', ``Gain'', ``MCC'', ``N50'', ``PPV'', ``precision'', ``RMSD'', ``sensitivity'', ``TPR'', ``tree error'', etc.
+Our analysis makes the necessarily pragmatic assumption that highly ranked tools on one accuracy metric will also be highly ranked on other accuracy metrics.}
+Many articles contained
+multiple benchmarks, in these cases we recorded ranks from each of these, the
 provenance of which is stored with the accuracy metric and raw speed
-and accuracy rank data for each tool. In line with rank-based
+and accuracy rank data for each tool {\color{red}(Supplementary Table 1)}. In line with rank-based
 statistics, the cases where tools were tied were resolved by using a
-midpoint rank (e.g., if tool 3 and 4 are tied, the rank 3.5 was used)
+midpoint rank (e.g., if tools ranked 3 and 4 are tied, the rank 3.5 was used)
 \cite{Mann1947-re}. Each rank extraction was independently verified by
@@
 
-\section*{Data availability}
+\begin{backmatter}
+
+\section*{Acknowledgements}
+The authors acknowledge the valued contribution of invaluable
+discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia
+Palmer and Jason Tylianakis. Murray Cox, Raquel
+Norel, Alexandros Stamatakis, Jens Stoye,Tandy Warnow, provided
+valuable feedback on drafts of the manuscript.
+
+This work was largely conducted on the traditional territory of K\=ai Tahu.
+
+\section*{Funding}
+PPG is supported by a Rutherford Discovery Fellowship,
+administered by the Royal Society Te Ap\=arangi, 
+PPG, AG and MAB acknowledge support from a Data Science Programmes grant
+(UOAX1932).
+
+\section*{Availability of data and materials}
+%\section*{Data availability}
 Raw datasets, software and documents are available under a CC-BY license:\\
@@
 
-\section*{Acknowledgements}
-
-The authors acknowledge the valued contribution of invaluable
-discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia
-Palmer and Jason Tylianakis. Murray Cox, Raquel
-Norel, Alexandros Stamatakis, Jens Stoye,Tandy Warnow, provided
-valuable feedback on drafts of the manuscript.
-
-PPG is supported by a Rutherford Discovery Fellowship,
-administered by the Royal Society Te Ap\=arangi, 
-PPG, AG and MAB acknowledge support from a Data Science Programmes grant
-(UOAX1932).
-
-This work was largely conducted on the traditional territory of K\=ai Tahu.
-
-%{\bf AND ETC....}
+\section*{Competing interests}
+The authors declare that they have no competing interests.
 
@@
 %----------------------------------------------------------------------------------------
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
+\bibliographystyle{bmc-mathphys}
 \bibliography{references.bib}
 
+\end{backmatter}
 \end{document}

### 2021-08-16 manuscript/supplementary.tex
Updating/adding figures, supplementary materials and minor manuscript edits.

@@
   (accuracy, speed, JH5, citations, H-index, M-index, relAge,
-  relCites, yearPublished, version, commits, contributors,
-  ).
-
+  relCites, yearPublished, version, commits, contributors, \#Issues,
+\%Open issues, \#Pull requests, and the \#Forks).
 }
@@
 \centering
-%\includegraphics[width=0.95\textwidth]{relAge-SpeedVsAccuracy-heatmap-edited.pdf}
-\includegraphics[width=0.45\textwidth]{relAge-speedAcc.pdf}
-\includegraphics[width=0.45\textwidth]{relAge-SpeedVsAccuracy-heatmap.pdf}
-\caption{{\bf Relative software tool ages and speeds/accuracies.} {\bf Left.} Violin plots for the relative age distribution
-  for software tools in each of the 5 2x2 cells indicated in B. The
-  five boxes correspond to the four extreme corners of the speed vs
-  accuracy spectrum (i.e. slow and inaccurate, slow and accurate, fast
-  and inaccurate, fast and accurate) and the central box (medium speed
-  and medium accuracy). {\bf Right.} A heatmap indicating the relative age
+%\includegraphics[width=0.45\textwidth]{relAge-speedAcc-jitterPlot.pdf}
+%\includegraphics[width=0.45\textwidth]{relAge-SpeedVsAccuracy-heatmap.pdf}
+\includegraphics[width=0.95\textwidth]{relAge.pdf}
+\caption{{\bf Relative software tool ages and speeds/accuracies.}
+{\bf A.} A heatmap indicating the relative age
   of software in a range of relative accuracy and speed rankings. Blue
@@
   and speed category, while light colours indicate younger software in
-  an accuracy and speed category. }
+  an accuracy and speed category.
+  {\bf B.} Jitter-plots for the relative age distribution
+  for software tools in each of the 5 3x3 cells indicated in {\bf B}. The
+  five row correspond to the four extreme corners of the speed vs
+  accuracy spectrum (i.e. slow and inaccurate, slow and accurate, fast
+  and inaccurate, fast and accurate) and the central box (medium speed
+  and medium accuracy). Median, quartile and extreme values are indicated with tick marks for each group.
+  Accurate and fast software tools tend to have been published more recently than inaccurate and slow tools (Wilcox rank sum test, $W=341,~P=5\times 10^{-3}$)
+   }
 \label{fig:ageplot}

### 2021-08-16 bin/prettyPlot.R

@@ -22,8 +22,6 @@ dr2005<-read.table( "rawRankSpeedData2005-2020.tsv", header=T)
 #d2015 <-read.table("meanRankSpeedData2015.tsv",      header=T)
 #dr2015<-read.table( "rawRankSpeedData2016-2020.tsv", header=T)
 
-
-
 ######################################################################
 #TESTING IF SPEED inv. ACCURACY IS TRUE:
 ####Regression:
@@ -33,51 +31,28 @@ summary(reg1)
 regM <- lm(accuracyRank ~ speedRank + H5 + citations + hindex + mindex + relAge + relCites + version + commits + contributors + issues + issuesFracOpen + pullrequests + forks, data=d2005, na.action=na.roughfix)
 summary(regM)
 
-####PCA:
-xx<-cbind(d2005$accuracyRank, d2005$speedRank, d2005$H5, d2005$citations, d2005$hindex, d2005$mindex, d2005$relAge, d2005$relCites,d2005$yearPublished, d2005$version, d2005$commits, d2005$contributors, d2005$issues, d2005$issuesFracOpen, d2005$pullrequests, d2005$forks)
-colnames(xx)<-c("accuracy","speed","JH5","citations","hindex","mindex","relAge","relCites","age", "version", "commits", "contributors", "issues", "issuesFracOpen", "pullrequests", "forks")
-pca<-prcomp(na.omit(xx), center = TRUE,scale=TRUE)
-pca
-summary(pca)
-
-####Random Forest:
-# library(randomForest)
-# ind <- sample(2,nrow(d2005),replace=TRUE,prob=c(0.7,0.3))
-# trainData <- d2005[ind==1,c(2,3,6:19)]
-# testData  <- d2005[ind==2,c(2,3,6:19)]
-# acc.rf <- randomForest(accuracyRank ~ ., data=trainData,ntree=1000,proximity=TRUE, na.action=na.roughfix)
-# round(importance(acc.rf,type=1), 2)
-# par(mfrow=c(1,1))
-# predTest <-predict(acc.rf,testData)
-# plot( testData$accuracyRank,predTest)
-# reg1  <- lm(predTest ~ testData$accuracyRank)
-# abline(reg1,col="red")
-# corTest<-cor.test( testData$accuracyRank,predTest)
[…]

### 2021-08-19 manuscript/manuscript-speed-accuracy.tex
Added Luis to the Acknowledgements, started describing additional features.

@@
 are not static and will change over time. In addition, where possible we also extract the version number, the number
-of commits and number of contributors from GitHub repositories.
+of commits, number of contributors
+{\color{red} total number ``issues'', the proportion of issues that remain open, the number of pull requests, and the number of times the code was forked}
+from public GitHub repositories.
 
@@
 benchmarks) and the maximum H- and corresponding M-indices for the corresponding
-authors, version number, and numbers of commits and contributors.
-The results are presented in Figure~\ref{fig:allfactors}A\&B. We find
+authors, version number,
+{\color{red}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
+The results are presented in Figure~\ref{fig:allfactors}A\&B and S5\&S6. We find
 significant associations between most of the citation-based metrics
@@
 Palmer and Jason Tylianakis. Murray Cox, Raquel
-Norel, Alexandros Stamatakis, Jens Stoye,Tandy Warnow, provided
+Norel, Alexandros Stamatakis, Jens Stoye, Tandy Warnow, and 
+Luis Pedro Coelho provided
 valuable feedback on drafts of the manuscript.

### 2021-08-26 manuscript/supplementary.tex
Added a powercurve analysis.

@@
 \graphicspath{{../figures/}}
+
+\def\numTools{499}
+%cat meanRankSpeedData.tsv | cut -f 4 | grep -v ^method$ | wc -l
+\def\numBenchmarkPubs{69}
+%cat speed-vs-accuracy-toolRanks2005-2020.tsv | cut -f 1 | sort  | uniq | grep -E '[0-9]'  | wc -l
+\def\numBenchmarks{134}
+%cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l
 
@@
 \centering
-\includegraphics[width=0.95\textwidth]{supplementary-distributions-permuted.pdf}
+\includegraphics[width=0.95\textwidth]{supplementary-distributions-permuted.pdf}\\
 \caption{The distributions for the permuted accuracy ({\bf left}) and speed ({\bf right}) values, where accuracy and speed ranks have been normalised to lie between 0 and 1. 
@@
 \includegraphics[width=0.99\textwidth]{numberBenchmarksPerToolBarplot.pdf}\\
-\includegraphics[width=0.99\textwidth]{numberRealValueFeaturesBarplot.pdf}
-\caption{ {\bf Top:} The number of benchmarks each software tool has appeared in. {\bf Bottom:}
+\includegraphics[width=0.99\textwidth]{numberRealValueFeaturesBarplot.pdf}\\
+\includegraphics[width=0.95\textwidth]{powerCurves.pdf}
+
+\caption{ {\bf Top:} The number of benchmarks each software tool has appeared in. {\bf Middle:}
   Plot of the amount missing/known data for each of the software features used in this analysis. 
+{\bf Bottom:} a power curve analysis to determine if the sample sizes for each set of features is sufficient to detect significant correlations between our features of interest.
+For example, for speed vs accuracy $N=\numTools$, which is indicated with a vertical line.
 }

### 2021-08-26 bin/prettyPlot.R

@@ -3,8 +3,7 @@
 #Rscript ../bin/prettyPlot.R
 #R CMD BATCH ../bin/prettyPlot.R
 
-
-library(MASS)  # in case it is not already loaded 
+library(MASS) 
 library(RColorBrewer)
 library(fields)
 library(gplots)
@@ -13,6 +12,7 @@ library(gplots)
 library("hash")
 library("vioplot")
 library("randomForest")
+library("WebPower")
 
 k <- 11
 my.cols <- rev(brewer.pal(k, "RdYlBu"))
@@ -174,53 +174,6 @@ for (i in 1:length(spearmansSP)){
 dev.off()
 
 
-#####################
-#Reviewer/Referee 2 requested plots:
-#plot each software package (potentially sized by the number of studies on which a software tool was included)
-pdf(file=    "../figures/numberBenchmarksPerToolBarplot.pdf", width = 100,  height = 5)
-op<-par(mfrow=c(1,1),cex=1.0,las=2, mar = c(12,4,4,4) + .1)
-bp<-barplot(d2005$numTests,names=d2005$method, ylab="Number of benchmarks",ylim=c(0,26),main="Number of benchmarks per software tool")
-dev.off()
-
-#Presumably the data are not fully complete (i.e., some scholar profiles couldn't be identified, etc). A table within either the methods or results breaking out the completeness of the records would be helpful.
-
-naFeatureCounts<-matrix(0, length(dNamesALL), 2)
-for(i in 1:length(dNamesALL)){      
-      naFeatureCounts[i,2] <- sum(is.na(d2005[,dNamesALL[i] == colnames(d2005)]))
-      naFeatureCounts[i,1] <- length(d2005[,dNamesALL[i] == colnames(d2005)]) - naFeatureCounts[i,2]      
-}
-rownames(naFeatureCounts) <- pNamesALL
-colnames(naFeatureCounts) <- c("Known", "NA")
-
-ixF <- sort(naFeatureCounts[,1], index.return=T, decreasing=T)$ix
-
-pdf(file=    "../figures/numberRealValueFeaturesBarplot.pdf", width = 10,  height = 5)
-op<-par(mfrow=c(1,1),cex=1.0,las=2, mar = c(7,4,4,4) + .1)
[…]

### 2021-08-31 manuscript/manuscript-speed-accuracy.tex
Further manuscript revisions to appease our Reviewer Overlords (who made excellent suggestions).

@@
   \begin{abstract} % abstract
-    \parttitle{Background:} Computational biology provides widely used and
+    \parttitle{Background} Computational biology provides widely used and
     powerful software tools for testing and making inferences about
@@
 
-    \parttitle{Method:} We have extracted accuracy and speed ranks from
+    \parttitle{Method} We have extracted accuracy and speed ranks from
     independent benchmarks of different bioinformatic software tools,
@@
 
-    \parttitle{Results:} We found that software speed, author reputation,
+    \parttitle{Results} We found that software speed, author reputation,
     journal impact, number of citations and age are all unreliable
@@
 
-    We also find that the field of bioinformatics has a
+    \parttitle{Discussion} We also find that the field of bioinformatics has a
     large excess of slow and inaccurate software tools, and this is
@@
 impact journals may be more likely to select manuscripts describing
-good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
+good software \cite{Garfield1955-wf}. \textbf{4. Author/group
 reputation:} the key to any project is the skills of the people
@@
 %mind VS naive parallelisation as in Python, R, etc.}
-
 %Other factors that are less quantifiable that influence whether a
@@
 %projects, as based upon literature surveys \cite{nasir2011critical}.
-
-With the wide adoption of GitHub ($47\%$ of the
-{\color{black}\numTools} tools included in this study could be
-linked to a GitHub repository), and consequently quantifiable data on software
-development time and intensity indicators, such as the number of contributors to code, number of code
-changes and versions is available for these
-\cite{ray2014large,Dozmorov:2018,Mangul:2019}.
+{\color{red}\textbf{6. Effective software versioning:} With the wide adoption of
+public version-control systems like GitHub, quantifiable data on
+software development time and intensity indicators, such as the number
+of contributors to code, number of code changes and versions is
+now available \cite{ray2014large,Dozmorov:2018,Mangul:2019}.}
 
@@
 speed, relative age, the productivity and impact of the corresponding
-authors, journal impact and the number of citations.
+authors, journal impact, number of citations {\color{red}and GitHub activity}.
 
@@
 \caption{\textbf{A.} A heatmap indicating the relationships between
-  different features of bioinformatic software tools. Spearman’s rho is used to
-  infer correlations between metrics such as citations based metrics, the year and relative age of publication, 
-  version number, GitHub derived activity measures, and 
-  the mean
-  relative speed and accuracy rankings. Red colours
-  indicate a positive correlation, blue colours indicate a negative
-  correlation. Correlations with an P-value less than 0.05 (corrected for multiple-testing using the Benjami-Hochberg method) are
-  indicated with a `$\vartimes$' symbol. The correlations with accuracy are illustrated in
-  more detail in \textbf{B}, the relationship between speed and
-  accuracy is shown in more detail in \textbf{Figure 2}.
-  %The
-  %dendrogram was computed using the default ‘heatmap.2 function in R,
-  %which computes a Euclidean distance matrix and a complete-linkage
-  %hierarchical clustering.
-  \textbf{B.} Violin plots of Spearman's correlations between permuted accuracy ranks and 
-    different software tool features. 
-    The unpermuted correlations are indicated with a red asterix.
-    For each benchmark, 1,000 permuted sets of accuracy and speed ranks were generated, and the ranks were normalised to lie between 0 and 1 (see Methods for details).
-  Circled asterixs are significant (empirical P-value $< 0.05$, corrected for multiple-testing using the Benjami-Hochberg method).}
+  different features of bioinformatic software tools. Spearman’s rho
+  is used to infer correlations between metrics such as citations
+  based metrics, the year and relative age of publication, version
+  number, GitHub derived activity measures, and the mean relative
+  speed and accuracy rankings. Red colours indicate a positive
+  correlation, blue colours indicate a negative
+  correlation. Correlations with a P-value less than 0.05 (corrected
+  for multiple-testing using the Benjamini-Hochberg method) are
+  indicated with a `$\vartimes$' symbol. The correlations with
+  accuracy are illustrated in more detail in \textbf{B}, the
+  relationship between speed and accuracy is shown in more detail in
+  \textbf{Figure~\ref{fig:speedaccuracy}}. \textbf{B.} Violin plots of Spearman's
+  correlations for permuted accuracy ranks and different software
+  features.  The unpermuted correlations are indicated with a red
+  asterisk.  For each benchmark, 1,000 permuted sets of accuracy and
+  speed ranks were generated, and the ranks were normalised to lie
+  between 0 and 1 (see Methods for details). Circled asterisks are
+  significant (empirical P-value $< 0.05$, corrected for
+  multiple-testing using the Benjamini-Hochberg method).}
 \label{fig:allfactors}
@@
 {\color{red}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
-The results are presented in Figure~\ref{fig:allfactors}A\&B and S5\&S6. We find
+The results are presented in Figure~\ref{fig:allfactors}A\&B, and Additional file 1: Figures S5\&S6. We find
 significant associations between most of the citation-based metrics
@@
 (Figure~\ref{fig:allfactors}B).  These features were not correlated
-with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Supplementary
-Figures~S5 and~S6}%\ref{fig:S2}
+with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Additional file 1: Figures S5\&S6}%\ref{fig:S2}
 ).  We also found that
@@
 We find that only the ``fast and inaccurate'' extreme class is at approximately the
-expected proportions based upon the permutation test (Figure 2B).
+expected proportions based upon the permutation test (Figure~\ref{fig:speedaccuracy}B).
 
@@
 permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The middle cell
-of Figure 2A and left-most violin plot of Figure 2B highlight this extreme, (Z=-6.38,
+of Figure~\ref{fig:speedaccuracy}A and left-most violin plot of Figure~\ref{fig:speedaccuracy}B highlight this extreme, (Z=-6.38,
 P-value=$9.0\times 10^{-11}$).
@@
 2005 and 2020. Our results provide significant support for the suggestion that there are major benefits to
-the long-term support of software development \cite{siepel2019challenges}. The
+the long-term support of software development \cite{siepel2019challenges}.
+The
 finding of a strong relationship between the number of commits and code contributors to
@@
 \cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.
+%Do the authors have putative mechanism that may describe the the lack of correlation between citations and accuracy?
+     {\color{red} Specifically, software tools published early are
+       more likely to appear in high impact journals due to their
+       perceived novelty and need. Yet without sustained maintenance
+       these may be outperformed by subsequent tools, yet early
+       publications still accrue citations from users, and all
+       subsequent software publications as tools need to be compared
+       in order to publish. Subsequent tools are not perceived to be
+       as novel, hence appear in ``lower'' tier journals, despite
+       being more reliable. Hence, the ``rich'' early publishers get
+       richer in terms of citations. Indeed, citation counts are
+       mainly predictive of age (Figure~\ref{fig:allfactors}A).}
 
 We found the lack of a correlation between software speed and accuracy
-surprising.  The slower software tools are overrepresented at both
-high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
+surprising.  The slower software tools are over-represented at both
+high and low levels of accuracy{\color{red}, with older tools enriched in this group} (Figure~\ref{fig:speedaccuracy} {\color{red}and Additional file 1: Figure S7}).
 In addition, there is an large under-representation of software that has
@@
 have unfortunate consequences as these tools may {\color{red}nevertheless} have further uses.
+
+%Do the authors have putative mechanism that may describe the the lack of correlation between citations and accuracy?
+
 %We also found that slow and inaccurate
@@
 We propose that the full spectrum of software tool accuracies and
-speeds serves a useful purpose to the research community. Like negative
-results, if honestly reported this information, illustrates to the research community
-that certain approaches are not practical research avenues
-\cite{fanelli2012negative}. % Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
-The current
-practices of publishers, editors, reviewers and authors of software
-tools therefore may be depriving our community of tools for building effective
-and productive workflows.
+speeds serves a useful purpose to the research community. Like
+negative results, if honestly reported this information, illustrates
+to the research community that certain approaches are not practical
+research avenues
+\cite{fanelli2012negative}. %Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
+The current novelty-seeking practices of many publishers, editors,
+reviewers and authors of software tools therefore may be depriving our
+community of tools for building effective and productive workflows.
+{\color{red} Indeed, the drive for novelty may be an actively harmful
+  criteria for the software development community, just as it is for reliable
+  and reproducible research
+  \cite{brembs2019reliable}. Novelty-criteria for publication may, in
+  addition, discourage continual, incremental improvements in code
+  post-publication in favour of splashy new tools that are likely to
+  accrue more citations. }
+
+{\color{red} In addition we suggest that further efforts be made to
+  encourage continual updates to software tools. To paraphrase some of
+  the suggestions of Siepel (2019), these efforts may include more
+  secure positions for developers, institutional promotion criteria
+  include software maintenance, lower publication barriers for
+  significant software updates, encourage further funding for software
+  maintenance and improvement - not just new tools
+  \cite{siepel2019challenges}. If these issues were recognised by
+  research managers, funders and reviewers, then perhaps
+  the future bioinformatic software tool landscape will be much improved. }
 
@@
 a machine-learning approach for identifying further candidate articles
-and to identify new search terms to include.
+and to identify new search terms to include. {\color{red}This is outlined in Additional file 1: Figure S1. }
 
@@
 pool of titles and abstracts, the counts were normalised by the total
-number of words in each set.
+number of words in each set. {\color{red}Additional file 1: Figure S2 shows exemplar word scores. }
 
@@
 relevant enriched terms were identified in later iterations. The final
-query is given in \textbf{supplementary materials}.
+query is given in {\color{red} \textbf{Additional file 1}}.
 
@@
 their tests, we extracted ranks for each tool.
-{\color{red}Until published datasets are made available in consistent, machine-readable formats this step is necessarily a manual process -- ranks were extracted from a mixture of manuscript figures, tables and supplementary materials, each data source is documented in Supplementary Table 1.
+{\color{red}Until published datasets are made available in consistent, machine-readable formats this step is necessarily a manual process -- ranks were extracted from a mixture of manuscript figures, tables and supplementary materials, each data source is documented in {\color{red} Additional file 2:} Table S1.
 In addition, a variety of accuracy metrics are reported e.g.
@@
 provenance of which is stored with the accuracy metric and raw speed
-and accuracy rank data for each tool {\color{red}(Supplementary Table 1)}. In line with rank-based
+and accuracy rank data for each tool {\color{red}(Additional file 2: Table S1)}. In line with rank-based
 statistics, the cases where tools were tied were resolved by using a
@@
 ($Z=\frac{x-\bar{x}}{s}$). For each cell in a $3\times 3$ grid
-a Z-score (and corresponding P-value is computed, either with the `pnorm' distribution function in R (Figure 2A) or empirically (Figure 2B)) is computed to illustrate the abundance or lack of tools in
+a Z-score (and corresponding P-value is computed, either with the `pnorm' distribution function in R (Figure~\ref{fig:speedaccuracy}A) or empirically (Figure~\ref{fig:speedaccuracy}B)) is computed to illustrate the abundance or lack of tools in
 a cell relative to the permuted data.
+
+{\color{red} The distributions for each feature and permuted accuracy or speed ranks are shown in {\color{red} Additional file 1: Figures S3\&S4}.
+Scatter-plots for each pair of features is shown in {\color{red} Additional file 1: Figure S5}.
+Plots showing the sample sizes for each tool, and feature are shown in {\color{red} Additional file 1: Figure S8}, illustrates a power analysis to show what effect sizes we are likely to detect for our sample sizes.
+}
 
@@
 
+\section*{Additional information}
+
+\subsection*{Additional file 1}
+
+{\color{red}Supplementary Figures S1-S8 and the neutral software benchmark reference list the accuracy and speed data is derived from [1-66].}
+
+\subsection*{Additional file 2}
+
+{\color{red}Supplementary Tables S1-S7.}
+
 %----------------------------------------------------------------------------------------

### 2021-08-31 manuscript/supplementary.tex
Further manuscript revisions to appease our Reviewer Overlords (who made excellent suggestions).

@@
   Plot of the amount missing/known data for each of the software features used in this analysis. 
-{\bf Bottom:} a power curve analysis to determine if the sample sizes for each set of features is sufficient to detect significant correlations between our features of interest.
+{\bf Bottom:} a power curve analysis shows the sample sizes for each set of features is sufficient to detect correlations $r\ge0.2$ between our features of interest with $power \ge 0.8$.
 For example, for speed vs accuracy $N=\numTools$, which is indicated with a vertical line.

### 2021-08-31 bin/prettyPlot.R

@@ -536,6 +536,7 @@ for( i in 1:length(numSamples)){
      lines(c(numSamples[i],numSamples[i]), c(0, 1))
 }
 legend(700, 0.4, paste("r = ", (0:5)/10), fil=my.cols.p)
+lines(c(0,1000),c(0.8,0.8),lty=2)
 dev.off()
 
 #####################
@@ -573,8 +574,55 @@ dev.off()
 #Reviewer 3 requested plots:
 #"What does the correlation between commits and citations imply?  Given the correlation between commits and accuracy, how might we interpret the lack of the transitive correlation between citations and accuracy?"
 
+######################################################################
 
-pdf(file=    "../figures/selected-scatter-plots.pdf", width = 20,  height = 20)
+
+
+pdf(file=    "../figures/selected-scatter-plots.pdf", width = 10,  height = 15)
+op<-par(mfrow=c(3,2),cex=1.5,las=2)
+plot(d2005$relAge,log10(d2005$citations+1),yaxt = "n",xlab="Relative age", ylab="Number of Citations", pch=20, col="red")
+axis(2,at=c(0,log10(2), 1:4), c(0, 10^(0:4)))
+abline(lm(log10(d2005$citations+1) ~ d2005$relAge, data=d2005, na.action=na.roughfix), lwd=2, col="red" )  #, na.action=na.omit
+cit.yr.cor <- cor.test(d2005$relAge, d2005$citations, method = "spearman")
+text(0.01, 0.25, paste("Spearman's Rho: ", signif(cit.yr.cor$estimate, digits=3), "\nP.value: ", signif(cit.yr.cor$p.value, digits=3)),pos=4)
+#############
+plot(d2005$relAge,log10(d2005$commits+1),yaxt = "n", ylab="Number of Commits", xlab="Relative age", pch=20, col="red")
+axis(2,at=c(0,log10(2), 1:4), c(0, 10^(0:4)))
+abline(lm(log10(d2005$commits+1) ~ d2005$relAge, data=d2005, na.action=na.roughfix), lwd=2, col="red" )  #, na.action=na.omit
+cit.com.cor <- cor.test(d2005$relAge, d2005$commits, method = "spearman")
[…]

### 2021-09-01 manuscript/manuscript-speed-accuracy.tex
Deleting unused figures. Updating README. Renamed supplementey tables.

@@
 of computational biology software, and manually extracted accuracy and
-speed rankings for {\color{black}\numTools~} unique software tools. For
+speed rankings for \numTools~ unique software tools. For
 each tool, we have collected measures that may be predictive
@@
 We have collected relative accuracy and speed ranks for
-{\color{black}\numTools~} distinct software tools. This software has
+\numTools~ distinct software tools. This software has
 been developed for solving a broad cross-section of computational biology
@@
 Each software tool was benchmarked in at least one of
-{\color{black}\numBenchmarkPubs~} publications that satisfy the Boulesteix
+\numBenchmarkPubs~ publications that satisfy the Boulesteix
 criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria
@@
 Data on the number of updates to software tools from GitHub such as
-the number of versions, commits and contributors was significantly
+the version number, {\color{red}and numbers of contributors, commits, forks and issues } was significantly
 correlated with software accuracy (respective Spearman’s rhos =
-{\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg
-corrected P-values = {\color{black}0.029, 0.042, 0.063}),
-Figure~\ref{fig:allfactors}B.  The significance of these features was
+{\color{red}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
+  corrected P-values = $5.4\times 10^{-4}$, $1.1\times 10^{-3}$, $7.9\times 10^{-4}$, $3.4\times 10^{-4}$, $3.2\times 10^{-4}$, Additional file 1: Figures S6}).
+%See arrays "spearmansA" and "spearmansAP" in prettyPlot.R
+The significance of these features was
 further confirmed with a permutation test
 (Figure~\ref{fig:allfactors}B).  These features were not correlated
-with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Additional file 1: Figures S5\&S6}%\ref{fig:S2}
-).  We also found that
+with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Additional file 1: Figures S5\&S6}).
+We also found that
 reputation metrics such as citations, author and journal H-indices,
@@
 expected distributions generated by 1,000 randomized ranks. The results
-of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}4}
+of this are shown in Figure~\ref{fig:speedaccuracy}. We identified 4
 of 9 bins where there was a significant excess or dearth of tools. For
-example, there was an excess of ``slow and inaccurate'' software ({\color{black}Z=3.39,
-P-value=$3.5\times 10^{-4}$}), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively).
+example, there was an excess of ``slow and inaccurate'' software (Z=3.39,
+P-value=$3.5\times 10^{-4}$), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively).
 We find that only the ``fast and inaccurate'' extreme class is at approximately the
@@
 \begin{figure*}[htb!]
-%\includegraphics[scale=0.55]{figure2.pdf}
 \centering
@@
 We have gathered data on the relative speeds and accuracies of
-{\color{black}\numTools~} bioinformatic tools from
-{\color{black}\numBenchmarkPubs} benchmarks published between
+\numTools~ bioinformatic tools from
+\numBenchmarkPubs~ benchmarks published between
 2005 and 2020. Our results provide significant support for the suggestion that there are major benefits to
@@
 \textbf{Criteria for inclusion:} We are interested in using
-computational biology benchmarks that satisfy Boulesteix’s
-(ALB) three criteria for a ``neutral comparison study''
+computational biology benchmarks that satisfy Boulesteix’s (ALB) three
+criteria for a ``neutral comparison study''
 \cite{Boulesteix2013-vb}. Firstly, the main focus of the article is
-the comparison and \textbf{not} the introduction of a new
-tool {{\color{red}as these can be biased} \cite{Buchka:2021a}. Secondly, the authors should be reasonably neutral, which means
-that the authors should not generally have been involved in the
-development of the tools included in the benchmark. Thirdly, the test
-data and evaluation criteria should be sensible. This means that the
-test data should be independent of data that tools have been trained
-upon, and that the evaluation measures appropriately quantify correct
-and incorrect predictions.
+the comparison and \textbf{not} the introduction of a new tool
+{\color{red}as these can be biased} \cite{Buchka:2021a}. Secondly, the
+authors should be reasonably neutral, which means that the authors
+should not generally have been involved in the development of the
+tools included in the benchmark. Thirdly, the test data and evaluation
+criteria should be sensible. This means that the test data should be
+independent of data that tools have been trained upon, and that the
+evaluation measures appropriately quantify correct and incorrect
+predictions. {\color{red} In addition, we excluded benchmarks with too
+  few tools $\le 3$, or those where the results were inaccessible (no
+  supplementary materials or poor figures).}
 
@@
 with benchmark studies. For background text we used Pubmed
-(\cite{Sayers2010-vm,McEntyre2001-fl} to select 8,908 articles that
+\cite{Sayers2010-vm,McEntyre2001-fl} to select 8,908 articles that
 contained the word “bioinformatics” in the title or abstract and were
@@
 into the set of background articles. This process was iterated and resulted in the identification of
-{\color{black}\numBenchmarkPubs} benchmark articles,
-containing {\color{black}\numBenchmarks} different benchmarks. Together these
-ranked {\color{black}\numTools~} distinct software packages.
+\numBenchmarkPubs benchmark articles,
+containing \numBenchmarks~ different benchmarks. Together these
+ranked \numTools~ distinct software packages.
 
@@
 ALB-criteria and contained data on both the accuracy and speed from
-their tests, we extracted ranks for each tool.
-{\color{red}Until published datasets are made available in consistent, machine-readable formats this step is necessarily a manual process -- ranks were extracted from a mixture of manuscript figures, tables and supplementary materials, each data source is documented in {\color{red} Additional file 2:} Table S1.
-In addition, a variety of accuracy metrics are reported e.g.
-``accuracy'', ``AUROC'', ``F-measure'', ``Gain'', ``MCC'', ``N50'', ``PPV'', ``precision'', ``RMSD'', ``sensitivity'', ``TPR'', ``tree error'', etc.
-Our analysis makes the necessarily pragmatic assumption that highly ranked tools on one accuracy metric will also be highly ranked on other accuracy metrics.}
-Many articles contained
-multiple benchmarks, in these cases we recorded ranks from each of these, the
+their tests, we extracted ranks for each tool. {\color{red}Until
+  published datasets are made available in consistent,
+  machine-readable formats this step is necessarily a manual process
+  -- ranks were extracted from a mixture of manuscript figures, tables
+  and supplementary materials, each data source is documented in
+ Additional file 2: Table S1. In addition, a variety of
+  accuracy metrics are reported e.g. ``accuracy'', ``AUROC'',
+  ``F-measure'', ``Gain'', ``MCC'', ``N50'', ``PPV'', ``precision'',
+  ``RMSD'', ``sensitivity'', ``TPR'', ``tree error'', etc. Our
+  analysis makes the necessarily pragmatic assumption that highly
+  ranked tools on one accuracy metric will also be highly ranked on
+  other accuracy metrics.}  Many articles contained multiple
+benchmarks, in these cases we recorded ranks from each of these, the
 provenance of which is stored with the accuracy metric and raw speed
-and accuracy rank data for each tool {\color{red}(Additional file 2: Table S1)}. In line with rank-based
-statistics, the cases where tools were tied were resolved by using a
-midpoint rank (e.g., if tools ranked 3 and 4 are tied, the rank 3.5 was used)
+and accuracy rank data for each tool {\color{red}(Additional file 2:
+  Table S1)}. In line with rank-based statistics, the cases where
+tools were tied were resolved by using a midpoint rank (e.g., if tools
+ranked 3 and 4 are tied, the rank 3.5 was used)
 \cite{Mann1947-re}. Each rank extraction was independently verified by
@@
 each benchmark were then normalised to lie between 0 and 1 using the
-formula $1-\frac{r-1}{n-1}$ where ‘$r$’ is a tool’s rank and ‘$n$’ is the
-number of tools in the benchmark. For tools that were benchmarked
+formula $1-\frac{r-1}{n-1}$ where ‘$r$’ is a tool’s rank and ‘$n$’ is
+the number of tools in the benchmark. For tools that were benchmarked
 multiple times with multiple metrics (e.g., BWA was evaluated in 6
@@
 \cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
-a mean normalised rank was used to summarise the accuracy and speed performance. 
-Or, more formally:
+a mean normalised rank was used to summarise the accuracy and speed
+performance.  Or, more formally:
 
@@
 Norel, Alexandros Stamatakis, Jens Stoye, Tandy Warnow, and 
-Luis Pedro Coelho provided
+{\color{red}Luis Pedro Coelho and three anonymous reviewers} provided
 valuable feedback on drafts of the manuscript.

### 2021-09-01 bin/prettyPlot.R

@@ -56,7 +56,7 @@ for(i in 1:N){
 	    pvalMatrix[i,j] <- p.adjust(spear$p.value,"BH", n=N*(N-1)/2-N)
 	    rhoMatrix[i,j]  <- spear$estimate
 	    #print(paste(c("P:", round(spear$p.value,3), "P.adj:", round(p.adjust(spear$p.value,"BH", n=N*(N-1)/2-N),3)), sep=""))
-	    if(p.adjust(spear$p.value,"BH", n=N*(N-1)/2-N) <= 0.05){ #spear$p.value < 0.05){
+	    if(p.adjust(spear$p.value,"BH", n=N*(N-1)/2-N) <= 0.05){ #spear$p.value < 0.05){#n is N choose 2, less the main diagonal (no double counting)
 		sigMatrix[i,j]  <- "X"
                 sigCount <- sigCount + 1
 	    }

### 2021-09-02 bin/prettyPlot.R
Revised README after an RFTM.

@@ -175,7 +175,6 @@ dev.off()
 
 
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-#FIGURE 1B
 #GENERATE PLOTS FOR THE PERMUTATION TESTS!:
 
 accPerms <-read.table("meanRankAccuracyPerms.tsv", header=T)
@@ -282,6 +281,7 @@ for (i in 1:length(spearmansA)){
 #correct for multiple testing:
 spearmansA.emp.pvals <- p.adjust(spearmansA.emp.pvals,"BH")
 
+#FIGURE 1B
 pdf(file=    "../figures/spearmanBarplot-withPerms-violin.pdf", width = 10,  height = 10)
 op<-par(mfrow=c(1,1),cex=1.5,las=2, mar = c(7,4,4,4) + .1)
 vioplot(rhoAccMatrix,ylim=c(-0.25,0.25), col="wheat",ylab="Spearman's rho", main="Correlation with accuracy")
@@ -352,6 +352,7 @@ heatmap.2(zScores[nrow(zScores):1,],
           col=rev(brewer.pal(n = colScale, name = "RdBu")), density.info="none", trace="none", dendrogram="none", symm=F,symkey=T,symbreaks=T, breaks=seq(-3.5,3.5,length=colScale+1), scale="none", cexRow=1.5, cexCol=1.5, margins = c(2, 2), key.title = "Z", Colv=FALSE, Rowv=FALSE, xlab="Speed",ylab="Accuracy", cex=2.0)
 dev.off()
 
+#FIGURE 2B
 pdf(file=    "../figures/zscores-withPerms-violin.pdf", width = 5,  height = 5)
 op<-par(mfrow=c(1,1),cex=1.0,las=2, mar = c(7,4,1,1) + .1)
 vioplot(
@@ -892,7 +893,7 @@ dev.off()
 # dev.off()
 
 ######################################################################
-#Figure 2: histogram inserts gridRes <- 10
+#histogram inserts gridRes <- 10
 # pdf(file=    "../figures/zscores-SpeedVsAccuracyH.pdf", width = 8,  height = 10)
 # par(mfrow=c(3,3),cex=1.4,las=1)
 # mx <- max(c(slowAcc,fastAcc,slowInacc,fastInacc))

### 2021-09-03 manuscript/manuscript-speed-accuracy.tex
Last wee polish of the manuscript and figures.

@@
 \usepackage{todonotes}
-\usepackage{maori}
-\graphicspath{{../figures/}}
+%\usepackage{maori}
+%\graphicspath{{../figures/}}
 \usepackage{xr}
-\externaldocument{supplementary}
+%\externaldocument{supplementary}
+
+%Māori characters:
+\DeclareUnicodeCharacter{0100}{\=A}
+\DeclareUnicodeCharacter{0101}{\=a}
+\DeclareUnicodeCharacter{0112}{\=E}
+\DeclareUnicodeCharacter{0113}{\=e}
+\DeclareUnicodeCharacter{012A}{\=I}
+\DeclareUnicodeCharacter{012B}{\=\i}
+\DeclareUnicodeCharacter{014C}{\=O}
+\DeclareUnicodeCharacter{014D}{\=o}
+\DeclareUnicodeCharacter{016A}{\=U}
+\DeclareUnicodeCharacter{016B}{\=u}
 
@@
 software tool does not necessarily mean that it is accurate or
-computationally efficient{\color{red};} instead usability, ease of installation,
+computationally efficient{\color{black};} instead usability, ease of installation,
 operating system support or other indirect factors may play a greater role in a software
@@
 The gold-standard for determining accuracy is for researchers independent
-of individual tool development to conduct benchmarking studies{\color{red};} these benchmarks can serve a useful role in
+of individual tool development to conduct benchmarking studies{\color{black};} these benchmarks can serve a useful role in
 reducing the over-optimistic reporting of software accuracy
@@
 \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
-impact:} {\color{red}many believe that} high profile journals are run by editors and reviewers who
+impact:} {\color{black}many believe that} high profile journals are run by editors and reviewers who
 carefully select and curate the best manuscripts. Therefore, high
@@
 \cite{Farrar2007-ky,Dematte2010-ph};
-{\color{red}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
+{\color{black}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
 rapid and sensitive pre-filtering
@@
 %projects, as based upon literature surveys \cite{nasir2011critical}.
-{\color{red}\textbf{6. Effective software versioning:} With the wide adoption of
+{\color{black}\textbf{6. Effective software versioning:} With the wide adoption of
 public version-control systems like GitHub, quantifiable data on
@@
 speed, relative age, the productivity and impact of the corresponding
-authors, journal impact, number of citations {\color{red}and GitHub activity}.
+authors, journal impact, number of citations {\color{black}and GitHub activity}.
 
@@
 of commits, number of contributors
-{\color{red} total number ``issues'', the proportion of issues that remain open, the number of pull requests, and the number of times the code was forked}
+{\color{black} total number ``issues'', the proportion of issues that remain open, the number of pull requests, and the number of times the code was forked}
 from public GitHub repositories.
@@
 authors, version number,
-{\color{red}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
+{\color{black}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
 The results are presented in Figure~\ref{fig:allfactors}A\&B, and Additional file 1: Figures S5\&S6. We find
@@
 Data on the number of updates to software tools from GitHub such as
-the version number, {\color{red}and numbers of contributors, commits, forks and issues } was significantly
+the version number, {\color{black}and numbers of contributors, commits, forks and issues } was significantly
 correlated with software accuracy (respective Spearman’s rhos =
-{\color{red}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
+{\color{black}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
   corrected P-values = $5.4\times 10^{-4}$, $1.1\times 10^{-3}$, $7.9\times 10^{-4}$, $3.4\times 10^{-4}$, $3.2\times 10^{-4}$, Additional file 1: Figures S6}).
@@
 (Figure~\ref{fig:allfactors}B).  These features were not correlated
-with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Additional file 1: Figures S5\&S6}).
+with speed however (see Figure~\ref{fig:allfactors}A {\color{black}\& Additional file 1: Figures S5\&S6}).
 We also found that
@@
 %Do the authors have putative mechanism that may describe the the lack of correlation between citations and accuracy?
-     {\color{red} Specifically, software tools published early are
+     {\color{black} Specifically, software tools published early are
        more likely to appear in high impact journals due to their
@@
 surprising.  The slower software tools are over-represented at both
-high and low levels of accuracy{\color{red}, with older tools enriched in this group} (Figure~\ref{fig:speedaccuracy} {\color{red}and Additional file 1: Figure S7}).
+high and low levels of accuracy{\color{black}, with older tools enriched in this group} (Figure~\ref{fig:speedaccuracy} {\color{black}and Additional file 1: Figure S7}).
 In addition, there is an large under-representation of software that has
@@
 that is not highly ranked on speed or accuracy. If correct, this may
-have unfortunate consequences as these tools may {\color{red}nevertheless} have further uses.
+have unfortunate consequences as these tools may {\color{black}nevertheless} have further uses.
 
@@
 community of tools for building effective and productive workflows.
-{\color{red} Indeed, the drive for novelty may be an actively harmful
+{\color{black} Indeed, the drive for novelty may be an actively harmful
   criteria for the software development community, just as it is for reliable
@@
 
-{\color{red} In addition we suggest that further efforts be made to
+{\color{black} In addition we suggest that further efforts be made to
   encourage continual updates to software tools. To paraphrase some of
@@
 the comparison and \textbf{not} the introduction of a new tool
-{\color{red}as these can be biased} \cite{Buchka:2021a}. Secondly, the
+{\color{black}as these can be biased} \cite{Buchka:2021a}. Secondly, the
 authors should be reasonably neutral, which means that the authors
@@
 evaluation measures appropriately quantify correct and incorrect
-predictions. {\color{red} In addition, we excluded benchmarks with too
+predictions. {\color{black} In addition, we excluded benchmarks with too
   few tools $\le 3$, or those where the results were inaccessible (no
@@
 a machine-learning approach for identifying further candidate articles
-and to identify new search terms to include. {\color{red}This is outlined in Additional file 1: Figure S1. }
+and to identify new search terms to include. {\color{black}This is outlined in Additional file 1: Figure S1. }
 
@@
 pool of titles and abstracts, the counts were normalised by the total
-number of words in each set. {\color{red}Additional file 1: Figure S2 shows exemplar word scores. }
+number of words in each set. {\color{black}Additional file 1: Figure S2 shows exemplar word scores. }
 
@@
 relevant enriched terms were identified in later iterations. The final
-query is given in {\color{red} \textbf{Additional file 1}}.
+query is given in {\color{black} \textbf{Additional file 1}}.
 
@@
 ALB-criteria and contained data on both the accuracy and speed from
-their tests, we extracted ranks for each tool. {\color{red}Until
+their tests, we extracted ranks for each tool. {\color{black}Until
   published datasets are made available in consistent,
@@
 provenance of which is stored with the accuracy metric and raw speed
-and accuracy rank data for each tool {\color{red}(Additional file 2:
+and accuracy rank data for each tool {\color{black}(Additional file 2:
   Table S1)}. In line with rank-based statistics, the cases where
@@
 
-{\color{red} The distributions for each feature and permuted accuracy or speed ranks are shown in {\color{red} Additional file 1: Figures S3\&S4}.
-Scatter-plots for each pair of features is shown in {\color{red} Additional file 1: Figure S5}.
-Plots showing the sample sizes for each tool, and feature are shown in {\color{red} Additional file 1: Figure S8}, illustrates a power analysis to show what effect sizes we are likely to detect for our sample sizes.
+{\color{black} The distributions for each feature and permuted accuracy or speed ranks are shown in {\color{black} Additional file 1: Figures S3\&S4}.
+Scatter-plots for each pair of features is shown in {\color{black} Additional file 1: Figure S5}.
+Plots showing the sample sizes for each tool, and feature are shown in {\color{black} Additional file 1: Figure S8}, illustrates a power analysis to show what effect sizes we are likely to detect for our sample sizes.
 }
@@
 Norel, Alexandros Stamatakis, Jens Stoye, Tandy Warnow, and 
-{\color{red}Luis Pedro Coelho and three anonymous reviewers} provided
+{\color{black}Luis Pedro Coelho and three anonymous reviewers} provided
 valuable feedback on drafts of the manuscript.
@@
 
-{\color{red}Supplementary Figures S1-S8 and the neutral software benchmark reference list the accuracy and speed data is derived from [1-66].}
+{\color{black}Supplementary Figures S1-S8 and the neutral software benchmark reference list the accuracy and speed data is derived from [1-66].}
 
@@
 
-{\color{red}Supplementary Tables S1-S7.}
+{\color{black}Supplementary Tables S1-S7.}
 

### 2021-09-03 bin/prettyPlot.R

@@ -581,17 +581,17 @@ dev.off()
 
 pdf(file=    "../figures/selected-scatter-plots.pdf", width = 10,  height = 15)
 op<-par(mfrow=c(3,2),cex=1.5,las=2)
-plot(d2005$relAge,log10(d2005$citations+1),yaxt = "n",xlab="Relative age", ylab="Number of Citations", pch=20, col="red")
+plot(d2005$yearPublished,log10(d2005$citations+1),yaxt = "n",xlab="Year Published", ylab="Number of Citations", pch=20, col="red", xlim=c(1980,2020))
 axis(2,at=c(0,log10(2), 1:4), c(0, 10^(0:4)))
-abline(lm(log10(d2005$citations+1) ~ d2005$relAge, data=d2005, na.action=na.roughfix), lwd=2, col="red" )  #, na.action=na.omit
-cit.yr.cor <- cor.test(d2005$relAge, d2005$citations, method = "spearman")
-text(0.01, 0.25, paste("Spearman's Rho: ", signif(cit.yr.cor$estimate, digits=3), "\nP.value: ", signif(cit.yr.cor$p.value, digits=3)),pos=4)
+abline(lm(log10(d2005$citations+1) ~ d2005$yearPublished, data=d2005, na.action=na.roughfix), lwd=2, col="red" )  #, na.action=na.omit
+cit.yr.cor <- cor.test(d2005$yearPublished, d2005$citations, method = "spearman")
+text(1980, 0.25, paste("Spearman's Rho: ", signif(cit.yr.cor$estimate, digits=3), "\nP.value: ", signif(cit.yr.cor$p.value, digits=3)),pos=4)
 #############
-plot(d2005$relAge,log10(d2005$commits+1),yaxt = "n", ylab="Number of Commits", xlab="Relative age", pch=20, col="red")
+plot(d2005$yearPublished,log10(d2005$commits+1),yaxt = "n", ylab="Number of Commits", xlab="Year Published", pch=20, col="red", xlim=c(1980,2020))
 axis(2,at=c(0,log10(2), 1:4), c(0, 10^(0:4)))
-abline(lm(log10(d2005$commits+1) ~ d2005$relAge, data=d2005, na.action=na.roughfix), lwd=2, col="red" )  #, na.action=na.omit
-cit.com.cor <- cor.test(d2005$relAge, d2005$commits, method = "spearman")
[…]

### 2022-01-21 manuscript/manuscript-speed-accuracy.tex
Removed a benchmark with a conflicted author, updated figures, manuscript etc.

@@
 %\externaldocument{supplementary}
+\usepackage{url}
+\usepackage{hyperref}
 
@@
 \startlocaldefs
-\def\numTools{499}
+\def\numTools{498}
 %cat meanRankSpeedData.tsv | cut -f 4 | grep -v ^method$ | wc -l
-\def\numBenchmarkPubs{69}
+\def\numBenchmarkPubs{68}
 %cat speed-vs-accuracy-toolRanks2005-2020.tsv | cut -f 1 | sort  | uniq | grep -E '[0-9]'  | wc -l
-\def\numBenchmarks{134}
+\def\numBenchmarks{133}
 %cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l
@@
   \begin{abstract} % abstract
-    \parttitle{Background} Computational biology provides widely used and
-    powerful software tools for testing and making inferences about
-    biological data. In the face of rapidly increasing volumes of
-    data, heuristic methods that trade software speed for accuracy may
-    be employed.  We are have studied these trade-offs using the
-    results of a large number of independent software benchmarks, and
-    evaluated whether external factors are indicative of accurate
+
+    \parttitle{Background:} Computational biology provides software
+    tools for testing and making inferences about biological data. In
+    the face of increasing volumes of data, heuristic methods
+    that trade software speed for accuracy may be employed. We
+    have studied these trade-offs using the results of a large number
+    of independent software benchmarks, and evaluated whether external
+    factors, including speed, author reputation, journal impact,
+    recency and developer efforts, are indicative of accurate
     software.
 
-    \parttitle{Method} We have extracted accuracy and speed ranks from
-    independent benchmarks of different bioinformatic software tools,
-    and evaluated whether the speed, author reputation, journal
-    impact, recency and developer efforts are indicative of accuracy.
+    \parttitle{Results:} We find that software speed, author
+    reputation, journal impact, number of citations and age are
+    unreliable predictors of software accuracy. This is unfortunate
+    because these are frequently
+    cited reasons for selecting software tools. However,
+    GitHub-derived statistics and high version numbers show that accurate
+    bioinformatic software tools are generally the product of many
+    improvements over time. We also
+    find an excess of slow and inaccurate bioinformatic software
+    tools, and this is consistent across many
+    sub-disciplines. Meanwhile, there are few tools that are
+    middle-of-road in terms of accuracy and speed trade-offs.
 
-    \parttitle{Results} We found that software speed, author reputation,
-    journal impact, number of citations and age are all unreliable
-    predictors of software accuracy. This is unfortunate because
-    citations, author and journal reputation are frequently cited
-    reasons for selecting software tools. However, GitHub-derived
-    records and high version numbers show that the accurate
-    bioinformatic software tools are generally the product of many
-    improvements over time, often from multiple developers.
+    \parttitle{Conclusions:} Our findings indicate that accurate
+    bioinformatic software is primarily the product of long-term
+    commitments to software-development. In addition, we hypothesise that
+    bioinformatics software suffers from publication bias. Software
+    that is intermediate in terms of both speed and accuracy may be
+    difficult to publish - possibly due to author, editor and reviewer
+    practises. This leaves an unfortunate hole in the literature, as
+    ideal tools may fall into this gap. High accuracy tools are
+    not always useful if they are slow, while high speed is not useful
+    if the results are also inaccurate.
 
-    \parttitle{Discussion} We also find that the field of bioinformatics has a
-    large excess of slow and inaccurate software tools, and this is
-    consistent across many sub-disciplines. Meanwhile, there are few
-    tools that are middle-of-road in terms of accuracy and speed
-    trade-offs. We hypothesise that a form of publication-bias
-    influences the publication and development of bioinformatic
-    software. In other words, software that is intermediate in terms
-    of both speed and accuracy may be difficult to publish - possibly
-    due to author, editor and reviewer practices. This leaves an
-    unfortunate hole in the literature as the ideal tools may fall
-    into this gap. For example, high accuracy tools are not always
-    useful if years of CPU time are required, while high speed is not
-    useful if the results are also inaccurate.
 %}
@@
 {\color{black}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
-  corrected P-values = $5.4\times 10^{-4}$, $1.1\times 10^{-3}$, $7.9\times 10^{-4}$, $3.4\times 10^{-4}$, $3.2\times 10^{-4}$, Additional file 1: Figures S6}).
+  corrected P-values = $6.7\times 10^{-4}$, $1.1\times 10^{-3}$, $8.4\times 10^{-4}$, $3.4\times 10^{-4}$, $3.1\times 10^{-4}$, Additional file 1: Figures S6}).
 %See arrays "spearmansA" and "spearmansAP" in prettyPlot.R
@@
 of 9 bins where there was a significant excess or dearth of tools. For
-example, there was an excess of ``slow and inaccurate'' software (Z=3.39,
-P-value=$3.5\times 10^{-4}$), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively).
+example, there was an excess of ``slow and inaccurate'' software ({\color{black}Z=3.40,
+P-value=$3.3\times 10^{-4}$}), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively).
 We find that only the ``fast and inaccurate'' extreme class is at approximately the
 expected proportions based upon the permutation test (Figure~\ref{fig:speedaccuracy}B).
+
+%#Z=
+%print(zScores[1,1])
+%#P=
+%pnorm(zScores[1,1], mean = 0, sd = 1, lower.tail = F)
 
@@
 %\section*{Data availability}
-Raw datasets, software and documents are available under a CC-BY license:\\
+Raw datasets, software and documents are available under a CC-BY license at Github \cite{Gardner2022a} and FigShare \cite{Gardner2022b}.
 %\fussy
@@
 %Additional documentation, code, figures and raw data is available here:\\
-\fussy
-\url{https://github.com/Gardner-BinfLab/speed-vs-accuracy-meta-analysis}
+%\fussy
+%Github: \url{https://github.com/Gardner-BinfLab/speed-vs-accuracy-meta-analysis}\\
+%FigShare: \url{https://doi.org/10.6084/m9.figshare.15121818.v2}
+
 \sloppy

### 2022-01-21 bin/prettyPlot.R

@@ -105,6 +105,11 @@ namesA     <-  namesA[ixA]
 dNamesA    <- dNamesA[ixA]
 pNamesA    <- pNamesA[ixA]
 
+#Print correlations and P-values mentioned in the text: "version number, and numbers of contributors, commits, forks and issues"  
+print(spearmansA )
+print(spearmansAP)
+
+
 
 #Figure S4A
 pdf(file=    "../figures/spearmanBarplot.pdf", width = 5,  height = 5)
@@ -281,6 +286,9 @@ for (i in 1:length(spearmansA)){
 #correct for multiple testing:
 spearmansA.emp.pvals <- p.adjust(spearmansA.emp.pvals,"BH")
 
+
+
+
 #FIGURE 1B
 pdf(file=    "../figures/spearmanBarplot-withPerms-violin.pdf", width = 10,  height = 10)
 op<-par(mfrow=c(1,1),cex=1.5,las=2, mar = c(7,4,4,4) + .1)
@@ -337,6 +345,12 @@ for(i in 1:gridRes){
 
 rownames(zScores)<-character(gridRes)
 colnames(zScores)<-character(gridRes)
+
+#For example, there was an excess of ``slow and inaccurate'' software
+#Z=
+print(zScores[1,1])
+#P=
+pnorm(zScores[1,1], mean = 0, sd = 1, lower.tail = F)
 
 binom.test(sigCount,    gridRes^2, p = 0.05)
 

### 2022-01-21 bin/tsv2data.pl

@@ -38,7 +38,7 @@ printResults("meanRankSpeedData.tsv", "meanRankAccuracyPerms.tsv", "meanRankSpee
 
 ###################################
 #Make all the figures, do some statistics: 
-#system("R CMD BATCH --no-save ../bin/prettyPlot.R");
+system("R CMD BATCH --no-save ../bin/prettyPlot.R");
 
 exit(0);
 
